### 2024-03-12 paper/paper.tex
udapte topic segmenter code

\documentclass[PMI,KR]{HSEUniversity}
% Возможные опции: KR или VKR; PI или BI

\title{Dynamic Topic Modeling for Voice Dialogues}
\author{Miakov Timofey Ilich}
\supervisor{Приглашенный преподаватель НИУ ВШЭ-Нижний Новгород}{А.А. Шимко}
\Year{2023}
\Abstract{

}

\bibliography{library.bib}

\begin{document}

\maketitle

\chapter*{Введение}

Юмор - это уникальное и творческое коммуникативное поведение, проявляемое во время социальных взаимодействий. Он создается мультимодальным способом, с использованием слов (текст), жестов (зрение) и просодических сигналов (акустика). Понимание юмора с помощью этих трех модальностей укладывается в рамки мультимодального языка - недавнего направление исследований в области обработки естественного языка (NLP).
Хотя распознавание юмора является признанной областью исследований в НЛП, подход к этой проблеме со стороны мультимодального глубокого обучения недостаточно изученная область.

В данной работе, будет проведено исследование по классификации видео на наличие юмора c использованием датасета UrFunny\cite{urfunny:2019}, который был создан специально для детектирования юмора с помощью разных модальностей. 
Кроме того, исследование будет сосредоточенно на использовании архитектуры Трансформер и его модификациях, которые будут использованны для извлечения признаков из разных модальностей и предсказании классов.

\chapter{UrFunny датасет}
Авторы используемого мной датасета, собрали 1866 видеороликов, а также их стенограммы с портала \href{https://www.ted.com/}{TED}.

TED talks - один из самых разнообразных каналов обмена идеями, как по выступающим, так и по темам. Спикеры из разных слоев общества, этнических групп и культур излагают свои мысли на широко популярном канале. Темы этих презентаций разнообразны: от научных открытий до повседневных событий. Благодаря разнообразию докладчиков и тем выступления на TED охватывают широкий спектр юмора. Таким образом, эта платформа представляет собой уникальный ресурс для изучения динамики юмора в мультимодальной среде.

Видеоролики TED включают в себя ручные расшифровки и маркеры аудитории. Транскрипции отличаются высокой надежностью, что, в свою очередь, позволяет выровнять текст и аудио. Транскрипции также включают надежно аннотированные маркеры поведения аудитории. В частности, маркер “смех” использовался в исследованиях НЛП в качестве показателя юмора. Предыдущие исследования выявили важность, как кульминационного момента, так и контекста для понимания и моделирования юмор. В юмористическом сценарии контекст - это постепенное развитие сюжета, а кульминационный момент - внезапный поворот сюжета, вызывающий смех.

В собранных данных, с помощью маркера смеха были размечены 8257 отрывков, где присутсвует юмор. из расшифровок. Контекст для этих видео извлекался из предыдущих предложений до кульминационной фразы. Используя аналогичный подход, 8257 отрывков, в которых не присутствует юмор, отбираются через случайные промежутки времени, где за последним предложением не сразу следует маркер смеха. 
В наборе данных наблюдается однородное 50\%-ное разделение между положительными и отрицательными примерами.

Приведем немного статистическиз данных о датасете.

Общая продолжительность всего набора данных составляет 90,23 часа. В UR-FUNNY насчитывается в общей сложности 1741 отдельный расказчиков и 417 различных тем. В пятерку наиболее частовстречающихся тем входяттехнологии, наука, культура, глобальные проблемы и дизайн. Средняя продолжительность каждого экземпляра данных составляет 19,67 секунды, при этом контекст в среднем из 14,7 сек. и кульминационный момент в среднем за 4,97 сек. Среднее количество слов в заголовке составляет 16,14, а среднее количество слов в контекстных предложениях - 14,80.

\chapter{Теоретическая часть}

Мультимодальные глубокое обучение — это сфера машинного обучения, целью которого является обучение моделей искусственного интеллекта обрабатывать и находить взаимосвязи между различными типами данных (модальностями) - как правило, изображениями или видео, аудио и текстом. Комбинируя различные модальности, модель глубокого обучения может более универсально воспринимать свою среду, поскольку некоторые сигналы существуют только в определенных модальностях.

Данный раздел предсавит архитектуру Трансформер, который изначально придуман для задач в области НЛП. А также её модификации для обработки текста (RoBERTa), видео (Vision Transformer) и звука (Wave2Vec), которые будут использоваться в практической части работы.

\section{NLP до Трансформера}

Для решения задач в сфере обработки естественного языка до определенного момента широко применялись рекурентный нейронные сети (RNN). Основной архитектурой в таких задачах была модель encoder-decoder\cite{seq2seq:2014} (рис A), в которой каждый из блоков  представлял из себя RNN. 
На вход энкодеру подавались последовательно токены, после чего на выход получался ембеддинг, который в свою очоередь подавался декодеру. Декодер, используя токены изначального предложения, а также переданное представление от кодировщика, генерирует целевую последовательность.

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.21]{img/transformer/enc_dec.png}
    \caption{Архитекутура encoder-decoder}
\end{figure}

Но у такой архитекутры были некоторые проблемы. Во-первых, все предствлание текста сжималось в один вектор, чаще всего это из-за многообразие возможных входных данных и огранниченности вектора терялась важная информация. Во-вторых, декодеру могла понадобиться разная информация от энкодера на разных этапах генерации, но декодер видит только одно представление источника.

Чтобы исправить эти проблемы, представили механизм внимания (attention mechanism\cite{attention:2014}).

\section{Механизм внимания - первый шаг к Трансформеру}

Механизм внимания является частью нейронной сети. На каждом этапе декодирования он решает, какие части входных данных важнее всего. При этом кодировщику не нужно сжимать весь исходный код в один вектор - он создает представления для всех токенов источника.

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.3]{img/transformer/attention_mech1.png}
    \caption{Схема работы механизма внимания}
\end{figure}

\newpage
Алгоритм работы механизма внимания в encoder-decoder модели приведен ниже.

\begin{itemize}
    \item На каждом этапе декодирования механизм внимания получает на вход: состояние декодера $h_t$ и все состояния кодировщика $s_1, s_2, \dots, s_m$ 
    \item Вычисляет коэффициенты внимания (attention scores) для каждого состояния кодеровщика $s_k$ механизм внимания вычисляет свою "релевантность" для этого состояния декодера $h_t$. \\
    Для вычисления можно применить любую функцию, однако, есть несколько популярных и простых вариантов, которые работают довольно хорошо.
    \begin{itemize}
        \item $scores(h_t, s_k) = h_t^T \cdot s_k $
        \item $scores(h_t, s_k) = h_t^T \cdot W \cdot s_k $, где $W$ - матрица (обучаемый параметр)
        \item $scores(h_t, s_k) = w_2^T \cdot \tanh(W \cdot [h_t^T, s_k]) $, где $W, w_2$ - матрицы (обучаемые параметры), а $[h_t^T, s_k]$ означает конкатенацию векторов 
    \end{itemize}

    \item Вычисляет веса внимания (attention weights): распределение вероятностей - softmax, применяемое к показателям внимания;
    \item Вычисляет вывод внимания : взвешенная сумма состояний декодера с весами внимания.
\end{itemize}

\begin{figure}[h]
    \centering
    \includegraphics[scale=1]{img/transformer/self-attention.png}
    \caption{Схема работы self-attention механизма}
\end{figure}

\subsection{Self-attention}

 Self-attention - является одним из ключевых компонентов модели, в которой токены взаимодействуют друг с другом. На каждом шаге в энкодере токены обмениваются информацией между собой, собирает контекст и обновляет предыдущее представление себя (рис 2.3). Такой подход полностью исключает последовательную обработку, как в RNN, что ускоряет работу нейронной сети и дает возможность обрабатывать каждое слово параллельно.

\newpage
\subsection{Query, Key, Value}

Каждый входной токен в self-attention получает три представления, соответствующие ролям, которые он может играть:
\begin{itemize}
    \item query - запрашивает информацию;
    \item key - показывает, какая информация есть у него;
    \item value - дает запрашиваемую информацию;
\end{itemize}

От слова, которое мы сейчас подаем на вход (центральное), требуется query вектор, этим вектором мы запрашиваем у других информацию о текущем слове в данном контексте. От каждого другого слова нам будет нужен вектор key, который показывает, какую информацию данное слово может сказать о смысле центрального слова, а также вектор value, из которого мы и получим нужную нам информацию.

Получив информацию от каждого вектора, мы подаем это на вход softmax функции и получаем итоговый эмбеддинг центрального слова. 
На рисунке нижен представлен полный алгоритм взаимодействия Q, K, V в self-attention механизме.

\newpage
\begin{figure}[h]
    \centering
    \includegraphics[scale=1]{img/transformer/query-key-value.png}
    \caption{Query, Key, Value в self-attention механизма}
\end{figure}

Формула для вычисления выходных данных внимания следующая:

\begin{center}
$Attention(q, k, v) = softmax(\frac{q \cdot k^T}{\sqrt{d_k}}) \cdot v $
\end{center}

\subsection{Masked self-attention}

Для энкодера и декодера используются разные способы получения токенов, если энкодер получает все токены сразу и каждый токен может смотреть на любой другой токен во входном предложении, в декодере, мы генерируем по одному токену за раз: во время генерации мы не знаем, какие токены мы будем генерировать в дальше.
Чтобы запретить декодеру заглядывать вперед, модель использует masked-self attention, который видит токены, только стоящие до него во входной последовательности.

Работа masked self-attention отличается во время обучения и инференса модели. Во время второго декодер не заглядывает вперед - так как, мы не знаем, что будет дальше. Но при обучении используются уже известные нам данные. Поэтому при обучении на вход декодеру подется все целевое предложение целиком - без масок.

На рисунке нижне приведен алгоритм работы masked self-attention механизма.

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.9]{img/transformer/masked self-attention.png}
    \caption{Схема работы masked Self-attention механизма}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[scale=1]{img/transformer/multi-head.png}
    \caption{Схема Multi-head attention}
\end{figure}

\subsection{Multi-Head Attention}

Обычно понимание роли слова в предложении требует понимания того, как оно связано с разными частями предложения. Это важно не только при обработке исходного предложения, но и при генерации целевого. 
Поэтому мы должны позволить модели сосредоточиться на разных вещах: для этого сделали multi-head attention. Вместо одного attention механизма multi-head attention имеет несколько «голов», которые работают независимо.

\begin{center}
$MultiHead(Q, K, V) = Concat(head_{1}, head_{2}, \dots, head_{n})$ \\
$head_{i} = Attention(QW^{i}_{Q}, KW^{i}_{K}, VW^{i}_{V})$
\end{center}

Формально это реализовано в виде нескольких механизмов внимания, результаты которых объединяются. (рис 2.6)

\newpage
\section{Трансформер: Архитектура модель}

Трансформер - это модель, основанная только на механизме винмания и оказавшая огромное влияние на NLP в частности и машинном обучении в целом. Впервые представленная в статье Attention is All You Need\cite{allyouneed:2017} в 2017 году. 
Модель сразу показала лучшие результаты в задаче перевода, в сравнении с encoder-decoder модели.

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.2]{img/transformer/transformer.png}
    \caption{Схема архитектуры Траснформер}
\end{figure}

Трансформер состоит из частей, которое ранне описывались. Encoder состоит из multi-head self-attention механизмов и генерирует представление слов, а decoder из masked self-attention механизма, который уже генерирует новую последовательность. Это происходит в несколько слоев, обычно 6.

Теперь рассмотрими другие составляющие архитекутры:\\

\subsection{Feed forward and Add $\&$ Norm}

Кроме блоков внимания, в Трансформере также присутствуют feed forward слои, которые состоят из линейного слоя и фукнции активации ReLU.

\begin{figure}[h]
    \centering
    \includegraphics[scale=1]{img/transformer/feed_forward.png}
    \caption{Feed-forward слой}
\end{figure}

Вторым важным слоем является LayerNorm, которой независимо нормализует векторное представление каждой последовательности в батче. В Трансформере нормализуются векторное представление каждого токена. Кроме того, LayerNorm имеет обучаемые параметры, $scale$ и $bias$, которые используются после нормализации для масштабирования выходных данных слоя.

\begin{figure}[h]
    \centering
    \includegraphics[scale=1]{img/transformer/layer_norm.png}
    \caption{LayerNorm слой}
\end{figure}

 На рис 2.8 стоит обратить внимание на то, что $\mu_{k}$ и $\sigma_{k}$ оцениваются для каждой последовательности, в отличии от $scale$ и $bias$, которые являются параметрами слоя.

\subsection{Positional encoding}

Поскольку Transformer не знает порядок входных токенов, т.к все токены обрабатываются одновременно, поэтому мы должны явно показать модели положения токенов. Для этого у нас есть два набора вложений: для самихтокенов и для их позиций. Тогда входное представление токена является сумма двух вложений.
Positional encodings можно обучить, но авторы оригинальной статьи обнаружили, что наличие исправленных вложений не ухудшает качество модели. Итоговая формула для позиционные вложения выглядят так:
\begin{center}
$PE_{pos,2i} = \sin(pos/10000^{2i/d_{model}})$, \\
$PE_{pos,2i+1} = \cos(pos/10000^{2i/d_{model}})$, \\
\end{center}

где $pos$ это позиция и $i$ является векторным измерением. Каждое измерение позиционного кодирования соответствует синусоиде, а длины волн образуют геометрическую прогрессию от $2\pi$ до 10000 $\cdot 2\pi$.

\begin{figure}[h]
    \centering
    \includegraphics[scale=1]{img/transformer/pos_encoding.png}
    \caption{Positional encoding}
\end{figure}

\subsection{Residual connection}

Как и в сверточных сетях, в Трансформере применяется residual connections. Смысл residual connection очень прост, добавляет входные данные к выходным данным модели, но в то же время это очень полезная вещь: он борется с проблемой затухания градиента по сети и позволяют использовать много слоев.

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.9]{img/transformer/res_block.png}
    \caption{Residual connection}
\end{figure}

В трансформаторе остаточные соединения используются после каждого блока внимания и полносвязного блока.

\section{Модель TimeSformer}
\subsection{Новый подход к решению задач компьютерного зрения}

Сверточные нейронные сети (CNN) являлись базовой архитектурой для задач распознавания изображений и показывали лучшие результаты во всех бенчмарках. Базовым слоем таких сетей являлись свертки, которые обрабатывали все изображения сразу.

Однако в 2016 году Google опубликовали статью, в которой представили новую нейронную сеть для работы с изображениями - Vision Transformer\cite{vit:2016} (далее ViT). Основным отличием от других нейронных сетей для обработки изображений стало использование архитекутры Transformer вместо сверток.

Ключевая идея, лежащая в основе ViT, заключается в обработке изображения как набор патчей (patches), а затем использовании модели на основе Transformer для обработки этих патчей. Участки сначала сглаживаются в набор векторов, которые затем обрабатываются. Это позволяет модели находить и запоминать долгосрочные зависимости между различными патчами, что потенциально приводит к повышению качества модели по сравнению с традиционными CNN.

Так как основной темой работы, являются видео, основной моделью стал не ViT, а еще более новая модель, разработанная Facebook, которая называется TimeSFormer. 
TimeSformer\cite{timesformer:2021} адаптирует стандартную архитектуру Трансформер и приемы, используемые в ViT, к видео, позволяя изучать пространственно-временные характеристики непосредственно из последовательности исправлений на уровне кадров.

\subsection{Обработка входного изображения}

На вход в модель подается разбитое на N частей изображение. Т.к. Трансформер принимает на вход 1D-вектора,  части изображения с помощью линейно проекции \textquotedblrightвытягивают\textquotedblright из 2D в 1D.
Также как и в задачах обработки текста, на вход в модель кроме патчей надо подать positional encoders.
Итоговые преобразования изображения выглядят так: \\

\noindent$z_{0} = [x_{class}; x^1_{p} \cdot E; x^2_{p} \cdot E; \dots x^N_{p} \cdot E] + E_{pos} $,\\
$x \in \mathds{R}^{H \times W \times C}$, $E \in \mathds{R}^{N \times (P^{2} \times C)}$,
где $C$ - кол-во каналов, $(P, P)$ - размер каждого патча, $N = HW / P^2$

\begin{figure}[h]
    \centering
    \includegraphics[scale=1]{img/vit/patches.png}
    \caption{Подготовка патчей}
\end{figure}

В качестве альтернативы патчам изображения, входная последовательность может быть сформирована
из карт характеристик (feature map) CNN. В этой гибридной модели проекция E применяется к участкам, извлеченным из карты объектов CNN, как частный случай, патчи могут иметь пространственный размер 1x1.
Патч с классом изображения и positional embeddings добавляются, как описано выше.

\subsection{Query-Key-Value вычисления}

TimeSFormer состоит из L блоков энкодера. В каждом блоке $l$, как и в оригинальном трансформере вычисляются Query, Key, Value для каждого патча из его прошлого состояния $z_{(p, t)}^{(l-1)}$

\begin{center}
$q_{(p, t)}^{(l, a)} = W_{Q}^{(l, a)} \cdot LN(z_{(p, t)}^{(l-1)}) \in \mathds{R}^{D_{h}}$,

$q_{(p, t)}^{(l, a)} = W_{K}^{(l, a)} \cdot LN(z_{(p, t)}^{(l-1)}) \in \mathds{R}^{D_{h}}$,

$q_{(p, t)}^{(l, a)} = W_{V}^{(l, a)} \cdot LN(z_{(p, t)}^{(l-1)}) \in \mathds{R}^{D_{h}}$,   
\end{center}
где $LN()$ означате LayerNorm, $a = 1, \dots, A$ - индексы голов в Multi-head attention.

\subsection{Attention weights}
Вычисление self-attention коэффициентов для $a_{(p, t)}^{(l, a)}$ для патча $(p, t)$ вычисляется как:

\begin{center}
$a_{(p, t)}^{(l, a)} = SoftMax(\frac{q_{(p, t)}^{(l, a)}}{\sqrt{D_{h}}}^{T} \cdot \left[k_{(0, 0)}^{(l, a)} {k_{(p', t')}^{(l, a)}}_{p' = 1, \dots, N, t' = 1, \dots, F} \right])$
\end{center}

\subsection{Encoding}

Закодированное представление $z_{(p, t)}^{(l)}$ в блоке $l$ получается путем вычисления взвешенной суммы векторов значений с использованием коэффициентов self-attention от каждой головы multi-head attention:

\begin{center}
$s_{(p, t),(0, 0)}^{(l, a)} = a_{(p, t),(0, 0)}^{(l, a)} \cdot v_{(0, 0)} + \sum\limits_{p'=1}^N \sum\limits_{t'=1}^F a_{(p, t),(p', t')}^{(l, a)} \cdot a_{(p', t')}^{(l, a)}$
\end{center}

Затем векторы каждого головы конкатенируются, проецируется и передается через MLP, используя residual connections после каждой операции:

\begin{center}
$z'_{(p, t)}^{(l)} = W_{O} \cdot \begin{bmatrix} 
  s_{(p, t)}^{(l, 1)} \\
  \vdots  \\
  s_{(p, t)}^{(l, A)}
\end{bmatrix} + z_{(p, t)}^{(l - 1)}$

$z_{(p, t)}^{(l)} = MLP(LN(z'_{(p, t)}^{(l)})) + z'_{(p, t)}^{(l)}$
\end{center}

Для финальной классификации используется MLP с 1 скрытым слоем. \\

Ниже представлен блок энкодера в TimeSFormer:

\begin{figure}[h]
    \centering
    \includegraphics[scale=.9]{img/vit/timesformer.png}
    \caption{Энкодер блок в TimeSFormer}
\end{figure}

\newpage
\section{Модель RoBERTa}

\subsection{BERT}

BERT был представлен в статье BERT: Pre-training of Deep Bidirectional Transformers for Language Understanding\cite{bert:2018}. Архитектура модели BERT - это энкодер Transformer. По новому подошли к процессу обучения и к тому, как BERT используется для последующих задач.

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.8]{img/bert/bert.png}
    \caption{Архитектура BERT}
\end{figure}

\begin{center}
\textbf{Новый способ обучения}    
\end{center}

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.6]{img/bert/bert_input.png}
    \caption{Входные данные}
\end{figure}

Входные данные для обучения: пары предложений со  разделенных специальным токеном-разделителем [SEP] (рис 2.15). Чтобы модель могла легко различать эти предложения, в дополнение к токеновым и позиционным вложениям она использует вложения сегментов. Также BERT имеет 2 цели обучения, для которых используются некоторые специальные токены([MASK], [CLS]). \\

\begin{center}
\textbf{Первая цель обучения обучения: Masked Language Modeling (MLM)}    
\end{center}
 Наиболее важной из двух целей обучения является цель моделирования замаскированного языка (MLM). На этапе обучения для MLM происходит следующее:
 \begin{itemize}
     \item Выбирается несколько токенов (каждый токен выбирается с вероятностью 15\%)
     \item Выбранные токены заменяются на [MASK] с вероятности 80\%, случайный токен - 15\% и остается без изменений - 10\%
     \item Модель должна предсказать исходный токен
\end{itemize}

На рисунке ниже показан пример шага обучения для одного предложения.

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.18]{img/bert/bert_tasks.png}
    \caption{Masked Language Modeling (MLM)}
\end{figure}
\newpage
\begin{center}
\textbf{Вторая цель обучения обучения: Next Sentence Prediction (NSP)}    
\end{center}

Задача «Предсказание следующего предложения» (NSP) представляет собой задачу бинарной классификации. На последнем уровне модели, исходя из токена [CLS], модель предсказывает, являются ли два предложения последовательными предложениями в некотором тексте или нет. В датасете для обучения, представленном в оригинальной статье сказанно, что при обучении 50\% примеров содержат последовательные предложения, извлеченные из обучающих текстов, а еще 50\% — случайную пару предложений.

Это задание учит модель понимать отношения между предложениями и позволяет BERT решать более сложные задачи.

\begin{center}
Пример из исходного датасета:    
\end{center}

Ввод: [CLS] мужчина пошел в [MASK] магазин [SEP] он купил галлон [MASK] молока [SEP]

Метка: Последовательные \\

Ввод: [CLS] мужчина пошел в магазин [MASK] [SEP] пингвин [MASK] летают \#\#less birds [SEP]

Метка: Непоследовательные\\

Следует также упомянуть, что BERT подходит для решения разных задач, ниже приведены основные типы задач:

\begin{itemize}
    \item Классификация отдельных предложений
    Для классификации отдельного предложения, вводятся токены предложения и [CLS] с классом предложения.
    \item Классификация пар предложений
    Для классифицикации пары предложений, вводятся токены предложений, как в обучении, и и [CLS] с классом предложения.
    \item Ответ на вопрос (Question answering)
    В задачах вопросов на ответы, на вход дают тект и вопрос. Ответ на этот вопрос содержится в исходном тексте, и задача состоит в том, чтобы найти отрывок из текста с ответом.
    \item Тэгирование слов в предложении
    В задачах тегирования прогнозируются теги для каждого токена. Например, в распознавании именованных объектов (NER) вы должны предсказать, является ли слово именованным объектом и его типом (например, местоположение, человек и т. д.).
\end{itemize}

\subsection{RoBERTa - улучшенная версия BERT}

Модель RoBERTa была предложена в книге RoBERTa: A Robustly Optimized BERT Pretraining Approach\cite{roberta:2019}. Он основан на модели Google BERT, выпущенной в 2018 году. Она основан на BERT и изменяет ключевые гиперпараметры, удаляя цель предварительного обучения в следующем предложении и обучая с гораздо большими мини-батчами и learning rate. \\

Список основных изменений представлен ниже:
\begin{itemize}
    \item Обучение большем объем данных (в 10 раз больше) и дольше по времени.
    \item Входные данные представляют собой более длинные последовательности, при оставшемся ограничении в 512 токенов, как и у BERT
    \item Динамическое маскирование токенов: токены маскируются по-разному в каждую эпоху, тогда как BERT делает это раз и навсегда
    \item Использование большого батча (относительно исходного BERT с батчем 256) при обучении модели
    \item Использование Byte-Pair Encoding с байтами в качестве субъединицы, а не с символами (из-за символов юникода)
\end{itemize}

\newpage
\section{Модель Wave2vec 2.0}

Фреймворк wave2vec 2.0\cite{w2v:2020} разработан для self-supervised обучения эмбеддингов на основе необработанных аудиоданных. Модель кодирует речевой звук с помощью многослойной сверточной нейронной сети, а затем маскирует участки результирующих скрытых речевых представлений, аналогично моделированию замаскированного языка (MLM), который был разобран в разделе про BERT. Скрытые представления передаются в Трансформер для построения контекстуализированных эмбеддингов, и модель обучается с помощью contrastive learning.

Как показано на рисунке ниже, модель обучается в два этапа. Первый этап проходит в режиме самоконтроля, который выполняется с использованием немаркированных данных и направлен на достижение наилучшего возможного представления речи. Вы можете думать об этом так же, как вы думаете о встраивании слов. Эмбеддинги слов также направлено на достижение наилучшего представления естественного языка. Основное отличие заключается в том, что Wav2Vec 2.0 обрабатывает аудио вместо текста.

Вторая фаза обучения - это контролируемая тонкая настройка, в ходе которой помеченные данные используются для обучения модели предсказывать определенные слова или фонемы (наименьшая возможная звуковая единица в определенном языке).

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.8]{img/wave2vec/w2v_stages.png}
    \caption{Этабы обучения wave2vec}
\end{figure}

Архитектура окончательной модели, используемой для прогнозирования, состоит из трех основных частей:

\begin{itemize}
    \item Cверточные слои, которые обрабатывают входные данные необработанной формы сигнала для получения скрытого представления - Z,
    \item Трансформер из n слоев, создающий эмбеддинги с контекстом - C,
    \item линейная проекция на выходные данные - Y
\end{itemize}

\subsection{Контрастивное обучение}
Контрастивное обучение - это концепция, в рамках которой входные данные преобразуются двумя различными способами. После этого модель обучается распознавать, являются ли два преобразования входных данных по-прежнему одним и тем же объектом. В Wav2Vec 2.0 трансформаторные слои являются первым способом преобразования, второй осуществляется путем квантования, что будет объяснено в дальнейшей части этой статьи. Более формально, для замаскированного скрытого представления $z_{t}$ мы хотели бы получить такое контекстное представление $c_{t}$, чтобы иметь возможность угадать правильное квантованное представление $q_{t}$ среди других квантованных представлений.

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.25]{img/wave2vec/wave2vec.png}
    \caption{Входные данные}
\end{figure}

\subsection{Квантизация}
Квантизация - это процесс преобразования значений из непрерывного пространства в конечный набор значений в дискретном пространстве.

 У нас есть вектор представления скрытой речи $z_{t}$ охватывает несколько фонем. Число фонем в языке конечно. Более того, число всех возможных пар фонем конечно. Это означает, что они могут быть идеально представлены одним и тем же скрытым речевым представлением. Кроме того, их число конечно, поэтому мы можем создать "кодовую книгу", содержащий все возможные пары фонем. Затем квантование сводится к выбору правильного кодового слова из "кодовую книгу". Однако вы можете себе представить, что количество всевозможных звуков огромно. Чтобы упростить обучение и использование, авторы Wav2Vec 2.0 создали G кодовых книг, каждая из которых состоит из V кодовых слов. Чтобы создать квантизированное представление, следует выбрать лучшее слово из каждой кодовой книги. Затем выбранные векторы объединяются и обрабатываются линейным преобразованием для получения квантизированное представления.

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.7]{img/wave2vec/w2v_books.png}
    \caption{Выбор токена их кодовых книг}
\end{figure}

Выбор правильного кодовго представления выбирается с помощью Gumbel softmax:

\begin{center}
$p_{g, v} = \frac{exp{\frac{sim(l_{g, v} + n_{v})}{\tau}}}{\displaystyle\sum_{k=1}^{V} exp{\frac{l_{g, k} + n_{k}}{\tau}}}$,
где 
\begin{itemize}
    \item sim — косинусное сходство,
    \item l — логиты, полученные свертками,
    \item $n_{k} = -log(-log(gl))$,
    \item $u_{k}$ - выборки из равномерного распределения U(0, 1),
    \item $\tau$ - температура
\end{itemize}
\end{center}

Поскольку это задача классификации, функция softmax представляется естественным выбором для выбора наилучшего кодового слова в каждой кодовой книге. Gumbel softmax поставляется с двумя улучшениями: рандомизацией и температурой $\tau$. Благодаря рандомизации модель более охотно выбирает разные кодовые слова во время обучения, а затем обновляет их веса. Важно, особенно в начале обучения, не допускать использования только подмножества кодовых книг.

\subsection{Diversity loss (Потеря разнообразия)}

Потеря разнообразия - это своего рода метод регуляризации. Авторы установили G=2 кодовых книги с V=320 кодовыми словами в каждой кодовой книге. Теоретически это дает $320 \cdot 320 = 102400$ возможных квантованных представлений. Но в процессе обучения можно  столкнутся с проблемой того, что модель использовать все эти возможности. Она научится использовать, например, только 100 кодовых слов из каждой кодовой книги и растратит весь потенциал кодовой книги впустую. Вот почему потеря разнообразия может быть полезной. Он основан на энтропии, которая может быть вычислена по следующей формуле:

\begin{center}
$H(x) = - \displaystyle\sum_{x} P(x)\cdot log(P(x))$, где
\end{center}

$x$ — возможный результат действия дискретной случайной величины $X$,

$P(x)$ — вероятность события $x$. \\

Энтропия принимает максимальное значение, когда распределение данных является равномерным. В случае wave2vec это означает, что все кодовые слова используются с одинаковой частотой. С помощью этого производится вычисление энтропию каждой кодовой книги по всей серии обучающих примеров, чтобы проверить, используются ли кодовые слова с одинаковой частотой. Максимизация этой энтропии побудит модель использовать преимущества всех кодовых слов. Максимизация равна минимизации отрицательной энтропии, которая является потерей разнообразия.\\

\begin{center}
$L_{d} = \frac{1}{GV} \cdot (-H(\overline{p}_{g})) = \frac{1}{GV}\displaystyle\sum_{g=1}^{G}\displaystyle\sum_{v=1}^{V} \overline{p}_{g, v} \cdot log(\overline{p}_{g, v})$ 
\end{center}

\newpage
\chapter{Практическая часть}

Основной задачей в работе будет классификация видео на наличие юмора с использованием разных моделей для разных модальностей.
Для работы с разными модальностями будут использованы 3 модели, которые были описаны в теоретической части:
\begin{itemize}
    \item Wave2Vec - извлечение аудио признаков,
    \item BERT - извлечение текстовых признаков,
    \item TimeSFormer - извлечение признаков из видео фрагментов,
\end{itemize}

Сначала модели попробуют предсказать классы в своих модальностях, после чего их предсказания будут обработаны вместе для улучшения результатов.

\section{Классификация по одной модальности}

Для извлечения признаков, как уже было описано, будут использоваться 3 различные модели. В силу ограниченных вычислительных мощностей, будут взяты предобученные модели с \href{https://huggingface.co/}{Hugging Face} и дообучены на наших данных.

Так как у нас ровное кол-во наблюдений положительного и отрицательного классов, делим выборку 70\% на дообучение и 30\% валидацию с одинаковым кол-вом примеров разных классов в обеих подвыборках.

Для каждой модели начала посчитаем качество модели по 3 метрикам: Accuracy (из-за отличной балансировки, можем использовать обычную accuracy), Precision, Recall.

\begin{itemize}
    \item Для работы с текстом будем использовать RoBERTa из библиотеки моделей HuggingFace: \href{https://huggingface.co/roberta-base}{roberta-base}. 
    \item Для работы с аудио будем использовать Wave2vec2 из библиотеки моделей HuggingFace: \href{https://huggingface.co/facebook/wav2vec2-base}{wav2vec2-base}.
    \item Для работы с видео будем использовать Wave2vec2 из библиотеки моделей HuggingFace: \href{https://huggingface.co/facebook/timesformer-base-finetuned-k400}{timesformer-base}.
\end{itemize}

С помощью предлагаемых Hugging Face интрументов, сделаем дообучение моделей под нашу задачу. \\

\subsection{Результаты}

\textbf{Далее в таблицах с результатами модальности будут называться в сокращенном виде: Текст - Т, Аудио - А, Видео - В.} \\

Результаты классификации по одной модальности: \\

\begin{center}
    \begin{tabular}{ |c||c|c|c|}
     \hline
     \multicolumn{4}{|c|}{Классификация по тексту} \\
         \hline
         Модальность & Accuracy & Precision & Recall \\
         \hline
         Т & 0.71377 & 0.70563 & 0.75064 \\
         В & 0.47836 & 0.50600 & 0.59729\\
         А &  0.50885 & 0.57865 & 0.78420 \\
         \hline
    \end{tabular} \\    
\end{center}

TimeSFormer показывает достаточно низкие значени метрик, относительно других моделей. Такие результаты связаны со спецификой задачи, моедль не фоксируется только на человеке, но на всем видеофрагменте. Возможным решением станет использование других моделей или инструментов, которые будут обрабатывать именно человека на видео.

Wave2vec показывает хороший recall (т.е модель хорошо угадывает, класс с юмором), но accuracy все еще плохой.

По результатам можем сделать вывод, что самым подходящей для детекции юмора оказалась текстовая модальност, с которой работала RoBERTa. Действительно по тексту легче всего понять, есть ли шутка, потому что не надо обрабатывать много второстепенной информации, как в видео, и "поведение"\ голоса разных людей во время шутки, как на аудио.

\section{Классификация видео с использованием нескольких модальностей}

Теперь попробуем собирать вместе голосование моделей в разных модальностях, сначала будем брать все комбинации 2 модальностей, а потом возьмем 3 модальности.

Подсчет результатов будем производить двумя разными сопобами:
\begin{itemize}
    \item Посчитать среднюю вероятность
    \item Выбирать голосованием, каждой модели за класс
\end{itemize} \\

\subsection{Результаты}

Результаты экспериментов с подсчетом средней вероятности: \\

\begin{center}
\begin{tabular}{ |c||c|c|c| }
     \hline
      Модальности & Accuracy & Precision & Recall \\
     \hline
      А + В & 0.49737 & 0.50512 & 0.60373 \\
      А + Т & 0.71377 & 0.70488 & 0.75257 \\
      В + Т & 0.63573 & 0.63569 & 0.66559 \\
      А + В + Т & 0.63081 & 0.62862 & 0.67074 \\
     \hline
\end{tabular} 
\end{center} \\

В голосованиях в парах и тройке результаты лучше, чем у А и В по одиночке.
А + Т слегка улучшает результат RoBERTa в одиночку, а в остальных парах и тройке, результаты улучшаются, но не достигают RoBERTa.

При спорных ситуациях в голосованием по классам, решающий голос за модальностями с лучшим результатом в прошлой частью (т.е. TimeSFormer имеет самый слабый голос, затем Wave2Vec и самый сильный голос у RoBERTa).

Результаты экспериментов с голосованием по классам: \\

\begin{center}
\begin{tabular}{ |c||c|c|c|}
     \hline
      Модальности & Accuracy & Precision & Recall \\
     \hline
      А + В & 0.49737 & 0.50512 & 0.60373 \\
      А + Т & 0.71377 & 0.70563 & 0.75064 \\
      В + Т & 0.71377 & 0.70563 & 0.75064 \\
      А + В + Т & 0.50885 & 0.50885 & 1.0 \\
    \hline
\end{tabular}
\end{center} \\

Здесь можно заметить, т.к. текст имеет решщающий голос, то в парах с текстом результаты одинаковые. 
В А + В + Т интересно заметить, что recall равен 1.0. Тоесть наша "тройка"\ относит все сэйплы к одному классу.

\chapter{Заключение}

В ходе проделанной работы были прочитаны статьи в разделе нейронных сетей, изучен механизм внимания и архитектура Трансформер с теоретической точки зрения. Также изучены модели, в основе которых лежит Трансформер: RoBerta, используемая для работы с текстовой модальностью, Wave2vec2, используемая для работы с звуковой модальностью, и TimeSFromer, используемая для работы с визуальной модальностью.

Далее были проведены эксперименты  на датасете UrFunnny с каждой из моделей. Изучив результаты можно увидеть, что RoBERTa показыает самые хорошие результаты. Предположения, почему получились такие результаты было сделано в пунке 3.1.1 под таблицей с результатами.

Что касается совмещения результатов моделей (голосование по вероятностям и финальным лэйблам), аудио и текст вместе улучшают результат RoBERTa в одиночку.

Для работы с моделями, обучение и инференс, был изучен фрэймворк HuggingFace, имеющий большую популярность в сфере глубокого обучения. Кроме того были использованы язык программирования Python, PyTorch - бекэнд для нейронных сетей, а также вспомогательные библиотеки librosa (работа со звуком) и Numpy.

Следующим шагом в исследовании детекции юмора может стать объединением моделей в одну, на примере CLIP от openAI или улучшение моделей в аудио и видео модальностях.

\putbibliography %Вместо этой команды будет вставлена библиография

\end{document}

### 2024-03-12 src/segmentation_module/data_utils.py

@@ -1,0 +1,160 @@
+## This data_processing script is initialized to process the DailyDal dataset as descripted in the SIGDIAL paper.
+## You can modify the code to set some settings off for your need.
+
+from torch.utils.data import Dataset
+import random
+import itertools
+
+###################### FOR DATALOADING ######################
+def load_txt(in_fname):
+    id2txt = {}
+    dup_dict = {}
+    dup_list = []
+    with open(in_fname) as in_file:
+        for idx, line in enumerate(in_file):
+            if line not in dup_dict.values():
+                id2txt[idx] = [utterance.replace(" __eou__","") for utterance in line.strip().split(" __eou__ ")]
+                dup_dict[idx] = line
+            else:
+                dup_list.append(idx)
+    return id2txt, dup_list
+
+def load_act(in_fname, dup_list):
+    id2act = {}
+    with open(in_fname) as in_file:
+        for idx, line in enumerate(in_file):
+            if idx not in dup_list:
+                id2act[idx] = line.strip().split(" ")
+    return id2act
+
+def load_topic(in_fname, dup_list):
+    id2topic = {}
+    with open(in_fname) as in_file:
+        for idx, line in enumerate(in_file):
+            if idx not in dup_list:
+                id2topic[idx] = int(line.strip())
+    return id2topic
+
+def load_meta(text_path, act_path, topic_path):
+    txt_dict, dup_list = load_txt(text_path)
+    topic_dict = load_topic(topic_path, dup_list)
+    act_dict = load_act(act_path, dup_list)
+
+    return txt_dict, topic_dict, act_dict
+
+def remove_duplicates(txt_dict, topic_dict, act_dict):
+    # Remove duplicated dialogues from all three dictionaries.
+    unique_utterances = {}
+    cleaned_topic_dict = {}
+    cleaned_act_dict = {}
+
+    for key, utterances in txt_dict.items():
+        if utterances not in unique_utterances.values():
+            unique_utterances[key] = utterances
+            cleaned_topic_dict[key] = topic_dict[key]
+            cleaned_act_dict[key] = act_dict[key]
+
+    return unique_utterances, cleaned_topic_dict, cleaned_act_dict
+
+
+################# FOR POS/NEG SAMPLES SELECTION ################
+def pesudo_generation_for_one_sample(utterances, acts, topic, txt_dict, act_dict, topic_dict):
+    sample_triple_for_this_dial = []
[…]

### 2024-03-12 src/segmentation_module/model_utils.py

@@ -1,0 +1,42 @@
+import torch
+from torch.utils.data import Dataset, DataLoader
+from transformers import BertTokenizer, BertModel
+import torch.nn.functional as F
+import torch.nn as nn
+
+
+class CoherenceNet(torch.nn.Module):
+    def __init__(self, bert_model, device):
+        super(CoherenceNet, self).__init__()
+        self.bert = bert_model
+        self.coherence_decoder = nn.Sequential(
+                                    nn.Linear(768, 768),
+                                    nn.ReLU(),
+                                    nn.Dropout(p=0.1),
+                                    nn.Linear(768, 2)
+                                )
+        self.device = device
+
+    def forward(self, batch):
+
+        output = []
+
+        for idx, sample in enumerate(batch):
+            pos_output = self.bert(**sample[0].to(self.device))
+            neg1_output = self.bert(**sample[1].to(self.device))
+            neg2_output = self.bert(**sample[2].to(self.device))
+
+            pos_output = pos_output.last_hidden_state[:, 0, :]
+            neg1_output = neg1_output.last_hidden_state[:, 0, :]
+            neg2_output = neg2_output.last_hidden_state[:, 0, :]
+
+            coherence_output_pos = self.coherence_decoder(pos_output)
+            coherence_output_neg1 = self.coherence_decoder(neg1_output)
+            coherence_output_neg2 = self.coherence_decoder(neg2_output)
+
+            output.append(torch.stack([F.softmax(coherence_output_pos.squeeze(0)), 
+                                       F.softmax(coherence_output_neg1.squeeze(0)), 
+                                       F.softmax(coherence_output_neg2.squeeze(0))], dim=0))
+
+        return torch.stack(output, dim=0)
+        

### 2024-03-12 src/segmentation_module/neural_texttiling.py

@@ -1,0 +1,324 @@
+import segeval
+from sklearn.metrics import f1_score
+import torch
+import numpy as np
+from sentence_transformers import SentenceTransformer
+
+
+def similarity_computing(texts, tokenizer, text_encoder, mode, device):
+    if mode == 'SC':
+        ### OPT 1: BI-ENCODER
+        #tokenization
+        inputs = tokenizer(texts, padding=True, truncation=True, return_tensors="pt")
+        # calculate the sentence embeddings by averaging the embeddings of non-padding words
+        with torch.no_grad():
+            embeddings = text_encoder(input_ids=inputs["input_ids"].to(device), attention_mask=inputs["attention_mask"].to(device))
+            attention_mask = inputs["attention_mask"].unsqueeze(-1).to(device)
+            embeddings = torch.sum(embeddings[0]*attention_mask, dim=1) / torch.sum(attention_mask, dim=1)  # mean-pooling
+            #embeddings = embeddings[0][:,0,:] # CLS
+        scores = torch.cosine_similarity(embeddings[:-1, :], embeddings[1:, :]).cpu().detach().numpy()
+
+    if mode == 'NSP': 
+        ### OPT 2: CROSS-ENCODER
+        scores = []
+        # calculate the NSP probability and use it as score
+        with torch.no_grad():
+            for i in range(len(texts)-1):
+                sent1 = texts[i]
+                sent2 = texts[i+1]
+                tokenized = tokenizer(sent1, sent2, padding='max_length', max_length = 128, truncation=True, return_tensors='pt')
+                logits = text_encoder(**tokenized.to(device)).logits
+                probabilities = torch.softmax(logits, dim=1)
+                probabilities = probabilities[0][0].item()
+                scores.append(probabilities)
+
+    if mode == 'CM':
+        ### OPT 3: COHERENCE SCORING
+        scores = []
+        # calculate the coherence score and use it as score
+        with torch.no_grad():
+            for i in range(len(texts)-1):
+                sent1 = texts[i]
+                sent2 = texts[i+1]
+                tokenized = tokenizer(sent1, sent2, padding='max_length', max_length = 128, truncation=True, return_tensors='pt')
+                output = text_encoder([[tokenized]*3])
+                coh_scores = output[:, :, 0]
+                scores.append(coh_scores.tolist()[0][0])
+
+    return scores
+
+
+def depth_computing(scores):
+
+    num_scores = len(scores)
+    depth_scores = []
+
+    for i in range(num_scores):
[…]

### 2024-03-12 src/segmentation_module/scripts/segmentation.sh

@@ -1,0 +1,7 @@
+DATASET=./data/eval/dialseg_711.json
+ENCODER_MODEL=dse-roberta-base
+
+python ./src/segmentation_module/segment.py \
+--dataset $DATASET \
+--text_encoder $ENCODER_MODEL \
+--mode NSP

### 2024-03-12 src/segmentation_module/scripts/train_coherence_model.sh

@@ -1,0 +1,11 @@
+DATASET=./datasets/dailydialog/
+ENCODER_MODEL=dse-roberta-base
+SAVE_PATH=./src/segmentation_module/checkpoints
+
+python ./src/segmentation_module/train.py \
+--dataset $DATASET \
+--epochs 1 \
+--batch_size 32 \
+--margin 1 \
+--text_encoder $ENCODER_MODEL \
+--checkpoints_path $SAVE_PATH 

### 2024-03-12 src/segmentation_module/train.py

@@ -1,0 +1,167 @@
+import os
+import argparse
+import torch
+
+from torch.utils.data import DataLoader, random_split
+from transformers import AdamW
+from transformers import get_linear_schedule_with_warmup
+from transformers import AutoModel, AutoTokenizer
+
+from data_utils import UtteranceDataset
+from model_utils import CoherenceNet
+
+from tqdm import tqdm
+import warnings
+warnings.filterwarnings('ignore')
+
+
+def parse_args():
+    parser = argparse.ArgumentParser(usage='python train.py -t path/to/dataset -e aws-ai/dse-bert-base -s path/to/checkpoints -m 1')
+    parser.add_argument('-ds', '--dataset', help='path of dataset for pseudo data generation', default='./data/train/dailydialog')
+    parser.add_argument('-ep', '--epochs', type=int, help='number of training epochs', default=10)
+    parser.add_argument('-bs', '--batch_size', type=int, help='batch size', default=24)
+    parser.add_argument('-m', '--margin', type=float, help='hyper-parameter of marginal ranking loss', default=1)
+    parser.add_argument('-encoder', '--text_encoder', help='text encoder for utterances', default='aws-ai/dse-bert-base')
+    parser.add_argument('-save', '--checkpoints_path', help='path to save checkpoints', default='./checkpoints/')
+    args = parser.parse_args()
+    return args
+
+def collate_fn(batch):
+    return batch
+
+def marginal_ranking_loss(batch, margin):
+    # Calculate pairwise differences
+    batch_tensor = batch[:, :, 0]
+
+    # Compute individual losses
+    # coh1 > coh2, coh1 > coh3, coh2 > coh3
+    loss_coh1_coh2 = torch.nn.functional.relu(margin - (batch_tensor[:, 0] - batch_tensor[:, 1]))
+    loss_coh1_coh3 = torch.nn.functional.relu(margin - (batch_tensor[:, 0] - batch_tensor[:, 2]))
+    loss_coh2_coh3 = torch.nn.functional.relu(margin - (batch_tensor[:, 1] - batch_tensor[:, 2]))
+
+    # Sum and average the losses
+    total_loss = (loss_coh1_coh2 + loss_coh1_coh3 + loss_coh2_coh3) / 3
+    total_loss = torch.mean(total_loss)
+
+    return total_loss
+
+def validation_metric(sample_list):
+    # one triple break down into 3 pairs: a>b, a>c, b>c
[…]

### 2024-04-17 paper/main.tex
paper start

--- 2024-03-12 paper/paper.tex
@@
-\documentclass[PMI,KR]{HSEUniversity}
+\documentclass[PMI,VKR]{HSEUniversity}
 % Возможные опции: KR или VKR; PI или BI
@@
 \author{Miakov Timofey Ilich}
-\supervisor{Приглашенный преподаватель НИУ ВШЭ-Нижний Новгород}{А.А. Шимко}
-\Year{2023}
+\supervisor{Professor NRU HSE - Nizhniy Novgorodw}{L.V. Savchenko }
+\Year{2024}
 \Abstract{
@@
 
-\chapter*{Введение}
+\chapter{Introduction}
 
-Юмор - это уникальное и творческое коммуникативное поведение, проявляемое во время социальных взаимодействий.
-Он создается мультимодальным способом, с использованием слов (текст), жестов (зрение) и просодических сигналов (акустика).
-Понимание юмора с помощью этих трех модальностей укладывается в рамки мультимодального языка - недавнего направление исследований в области обработки естественного языка (NLP).
-Хотя распознавание юмора является признанной областью исследований в НЛП, подход к этой проблеме со стороны мультимодального глубокого обучения недостаточно изученная область.
+\section{Formulation of the problem}
 
-В данной работе, будет проведено исследование по классификации видео на наличие юмора c использованием датасета UrFunny\cite{urfunny:2019}, который был создан специально для детектирования юмора с помощью разных модальностей.
-Кроме того, исследование будет сосредоточенно на использовании архитектуры Трансформер и его модификациях, которые будут использованны для извлечения признаков из разных модальностей и предсказании классов.
+\section{Relevance of the topic}
 
-\chapter{UrFunny датасет}
-Авторы используемого мной датасета, собрали 1866 видеороликов, а также их стенограммы с портала \href{https://www.ted.com/}{TED}.
+\section{The purpose of the diploma}
 
-TED talks - один из самых разнообразных каналов обмена идеями, как по выступающим, так и по темам.
-Спикеры из разных слоев общества, этнических групп и культур излагают свои мысли на широко популярном канале.
-Темы этих презентаций разнообразны: от научных открытий до повседневных событий.
-Благодаря разнообразию докладчиков и тем выступления на TED охватывают широкий спектр юмора.
-Таким образом, эта платформа представляет собой уникальный ресурс для изучения динамики юмора в мультимодальной среде.
+\section{Objectives}
 
-Видеоролики TED включают в себя ручные расшифровки и маркеры аудитории.
-Транскрипции отличаются высокой надежностью, что, в свою очередь, позволяет выровнять текст и аудио.
-Транскрипции также включают надежно аннотированные маркеры поведения аудитории.
-В частности, маркер “смех” использовался в исследованиях НЛП в качестве показателя юмора.
-Предыдущие исследования выявили важность, как кульминационного момента, так и контекста для понимания и моделирования юмор.
-В юмористическом сценарии контекст - это постепенное развитие сюжета, а кульминационный момент - внезапный поворот сюжета, вызывающий смех.
-
-В собранных данных, с помощью маркера смеха были размечены 8257 отрывков, где присутсвует юмор.
-из расшифровок.
-Контекст для этих видео извлекался из предыдущих предложений до кульминационной фразы.
-Используя аналогичный подход, 8257 отрывков, в которых не присутствует юмор, отбираются через случайные промежутки времени, где за последним предложением не сразу следует маркер смеха.
-В наборе данных наблюдается однородное 50\%-ное разделение между положительными и отрицательными примерами.
-
-Приведем немного статистическиз данных о датасете.
-
-Общая продолжительность всего набора данных составляет 90,23 часа.
-В UR-FUNNY насчитывается в общей сложности 1741 отдельный расказчиков и 417 различных тем.
-В пятерку наиболее частовстречающихся тем входяттехнологии, наука, культура, глобальные проблемы и дизайн.
-Средняя продолжительность каждого экземпляра данных составляет 19,67 секунды, при этом контекст в среднем из 14,7 сек.
-и кульминационный момент в среднем за 4,97 сек.
-Среднее количество слов в заголовке составляет 16,14, а среднее количество слов в контекстных предложениях - 14,80.
-
-\chapter{Теоретическая часть}
+\chapter{Theoretical explanation}
 
@@
 
-\section{NLP до Трансформера}
+\section{Dialogue Segmentation}
 
@@
 
-\section{Механизм внимания - первый шаг к Трансформеру}
+\section{Text embeddings extraction}
 
@@
 
-\subsection{Self-attention}
+\subsection{Architecture of BERT model}
 
@@
 \newpage
-\subsection{Query, Key, Value}
+\subsection{}
 

### 2024-04-18 paper/main.tex
update settings

@@
 \begin{document}
-
 \maketitle
@@
 
-\section{Formulation of the problem}
+\section{Motivation}
 
-\section{Relevance of the topic}
-
-\section{The purpose of the diploma}
-
-\section{Objectives}
+\section{Research Objectives}
 
@@
 
-Мультимодальные глубокое обучение — это сфера машинного обучения, целью которого является обучение моделей искусственного интеллекта обрабатывать и находить взаимосвязи между различными типами данных (модальностями) - как правило, изображениями или видео, аудио и текстом.
-Комбинируя различные модальности, модель глубокого обучения может более универсально воспринимать свою среду, поскольку некоторые сигналы существуют только в определенных модальностях.
+The pipeline of dynamic thematic modelling of audio dialogues, which will be considered in this paper, is divided into several main stages. 
+These include segmentation, topic extraction, and topic evolution. 
+The approaches and algorithms used in these stages will be described in detail from a theoretical point of view.
 
-Данный раздел предсавит архитектуру Трансформер, который изначально придуман для задач в области НЛП.
-А также её модификации для обработки текста (RoBERTa), видео (Vision Transformer) и звука (Wave2Vec), которые будут использоваться в практической части работы.
-
-\section{Dialogue Segmentation}
+\section{Dialogue Segmentation Stage}
 
@@
         Т           & 0.71377  & 0.70563   & 0.75064  \\
-         В & 0.47836 & 0.50600 & 0.59729\\
+        В           & 0.47836  & 0.50600   & 0.59729  \\
         А           & 0.50885  & 0.57865   & 0.78420  \\

### 2024-04-19 paper/main.tex
update paper

@@
 \begin{document}
+
 \maketitle
@@
 \section{Research Objectives}
+The objectives of the thesis work are:
+\begin{itemize}
+    \item A review of existing approaches to dynamic text and dialogue topic modelling
+    \item The definition of the main stages for DTM for audio dialogues
+    \item The identification of suitable training data sets and their collection
+    \item An exploration of the possibilities of using large language models for some stages and the selection of models suitable for training in limited resources
+    \item The development of the architecture of all stages and their interconnections
+    \item The identification of key indicators to assess the quality of each stage is also required
+    \item Visual analysis of the pipeline operation on pre-selected audio data
+\end{itemize}
 
@@
 The pipeline of dynamic thematic modelling of audio dialogues, which will be considered in this paper, is divided into several main stages. 
-These include segmentation, topic extraction, and topic evolution. 
+These include audio-to-text, dialogue segmentation, topic extraction, and topic evolution. 
 The approaches and algorithms used in these stages will be described in detail from a theoretical point of view.
+
+\section{Automatic Speech Recognition Stage}
 
@@
 
-Для решения задач в сфере обработки естественного языка до определенного момента широко применялись рекурентный нейронные сети (RNN).
-Основной архитектурой в таких задачах была модель encoder-decoder\cite{seq2seq:2014} (рис A), в которой каждый из блоков  представлял из себя RNN.
-На вход энкодеру подавались последовательно токены, после чего на выход получался ембеддинг, который в свою очоередь подавался декодеру.
-Декодер, используя токены изначального предложения, а также переданное представление от кодировщика, генерирует целевую последовательность.
+The dialogue segmentation stage represents the initial stage for textual type of data. 
+At this stage, a set of strings is input, with each line representing an utterance from one dialogue participant. 
+The result is a set of utterance indices that indicate the beginning of a thematically related group.
+
+In accordance with the majority of previous studies, we have utilise the best open-source solution based on the presented in papers metrics. 
+Let us now examine the algorithm in greater detail.
+
+The input of stage is a set of utterances $\{u1, u2, ..., uk\}$, splited into pairs of consecutive utterance. 
+All utterances text encode by encoder model.
+
+\subsection{BERT for embeddings extraction}
+
+The BERT model is a Transformer encoder that was first introduced in the article BERT: Pre-training of Deep Bidirectional Transformers for Language Understanding\cite{bert:2018}.  
+This article presents a new approach to the learning process and how BERT is used for downstream tasks.
 
@@
     \centering
-    \includegraphics[scale=0.21]{img/transformer/enc_dec.png}
-    \caption{Архитекутура encoder-decoder}
+    \includegraphics[scale=0.8]{img/bert.png}
+    \caption{BERT architecture}
 \end{figure}
 
-Но у такой архитекутры были некоторые проблемы.
-Во-первых, все предствлание текста сжималось в один вектор, чаще всего это из-за многообразие возможных входных данных и огранниченности вектора терялась важная информация.
-Во-вторых, декодеру могла понадобиться разная информация от энкодера на разных этапах генерации, но декодер видит только одно представление источника.
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.6]{img/bert_input.png}
+    \caption{}
+\end{figure}
 
-Чтобы исправить эти проблемы, представили механизм внимания (attention mechanism\cite{attention:2014}).
+The input data for training comprises pairs of sentences separated by a special token separator [SEP] (Figure 2.15). 
+In order to facilitate the model's ability to distinguish between these sentences, it employs segment embeddings in addition to token and position embeddings. 
+BERT also has two learning objectives for which some special tokens ([MASK], [CLS]) are used.
 
-\section{Text embeddings extraction}
+The first objective of the training process is the masked language modelling (MLM). 
+During the learning phase of MLM, the following occurs:
+
+\begin{itemize}
+    \item A number of tokens are selected with a probability of 15\% for each token.
+    \item The selected tokens are replaced by [MASK] with a probability of 80\%, a random token with a probability of 15\% and remain unchanged with a probability of 10\%.
+    \item The model must predict the original token.
+\end{itemize}
+
+The figure below illustrates an example of a training step for a single sentence.
+
+The Next Sentence Prediction (NSP) task is a binary classification task. 
+In order to understand relationship between two sentences, BERT training process also uses next sentence prediction.
+During training the model gets as input pairs of sentences and it learns to predict if the second sentence is the next sentence in the original text as well. 
+The training dataset presented in the original paper states that when trained, 50\% of the examples contain related sentences extracted from the training texts, while the remaining 50\% contain a random pair of sentences.
+
+In terms of our approach to topic segmentation, authors utilise Next Sentence Prediction (NSP) BERT as the encoder, due to the similarity of these two tasks. 
+Both tasks take a pair of sentences/utterances as input, with the objective of predicting the appropriate next sentence, which should be topically related.
+
+In more detail, the positive instances $(s_{i}, s_{t+i})$ and negative instances $(s_{i}, s_{t-i})$ are fed into the model in the form of the following sequence: 
+\begin{center}
+    $[CLS] s_{i} [SEP] s_{t+/-i} [SEP]$,
+\end{center}
+where the symbols $[CLS]$ and $[SEP]$ represent special tokens in BERT.
+
+Here, the contextualised representation of $[CLS]$ is employed as the topic-aware embedding to predict the degree of matching between the two input utterances in terms of topic. 
+The topical coherence score is estimated by passing the $[CLS]$ representation through another MLP.
+
+\subsection{Depth Scores calculation}
+
+Subsequently, an utterance-pair coherence scoring model is applied to $k - 1$ consecutive pairs, resulting in a sequence of coherence scores $\{c1, c2, ..., ck-1\}$. 
+The value of $c_i \in [0, 1]$ indicates the topical relatedness of the two utterances in the $i_th$ pair. 
+The algorithm utilises coherence scores as a foundation idea, yet enhances them through the sequence of "depth scores". $\{dp1, dp2, ..., dpk-1\}$ is calculated to measure the sharpness of a valley. 
+This is achieved by examining the highest coherence scores $h_l(i)$ and $h_r(i)$ on either side of interval $i$.
+
+The depth score for interval $i$, $dp_i$, is calculated as follows: 
+\begin{center}
+    $D_{pi} = \frac{ h_{l}(i) + h_{r}(i) - 2 \cdot c_{i}}{2}$. 
+\end{center}
+
+A higher depth score indicates a lower topical relatedness between the pair of utterances. 
+The threshold $\tau$, which is used to identify segment boundaries, is computed from the mean $\mu$ and standard deviation $\sigma$ of depth scores: $\tau = \mu - \frac{\sigma}{2}$. 
+A pair of utterances with a depth score exceeding $\tau$ will be selected to have a segment boundary between them.
+
+\section{Topic Extraction Stage}
+
+\subsection{Methods overview}
+
+\subsection{Large Language Models}
+
+\subsection{LLM for topic extraction}
+
+\section{Topic Evolution over time}
+
+\section{Attention mechanism}
 
@@
 
-\subsection{Architecture of BERT model}
+\subsection{Self-attention}
 
@@
 
-\section{Модель TimeSformer}
-\subsection{Новый подход к решению задач компьютерного зрения}
+\newpage
 
-Сверточные нейронные сети (CNN) являлись базовой архитектурой для задач распознавания изображений и показывали лучшие результаты во всех бенчмарках.
-Базовым слоем таких сетей являлись свертки, которые обрабатывали все изображения сразу.
-
-Однако в 2016 году Google опубликовали статью, в которой представили новую нейронную сеть для работы с изображениями - Vision Transformer\cite{vit:2016} (далее ViT).
-Основным отличием от других нейронных сетей для обработки изображений стало использование архитекутры Transformer вместо сверток.
-
-Ключевая идея, лежащая в основе ViT, заключается в обработке изображения как набор патчей (patches), а затем использовании модели на основе Transformer для обработки этих патчей.
-Участки сначала сглаживаются в набор векторов, которые затем обрабатываются.
-Это позволяет модели находить и запоминать долгосрочные зависимости между различными патчами, что потенциально приводит к повышению качества модели по сравнению с традиционными CNN.
-
-Так как основной темой работы, являются видео, основной моделью стал не ViT, а еще более новая модель, разработанная Facebook, которая называется TimeSFormer.
-TimeSformer\cite{timesformer:2021} адаптирует стандартную архитектуру Трансформер и приемы, используемые в ViT, к видео, позволяя изучать пространственно-временные характеристики непосредственно из последовательности исправлений на уровне кадров.
-
-\subsection{Обработка входного изображения}
-
-На вход в модель подается разбитое на N частей изображение.
-Т.к.
-Трансформер принимает на вход 1D-вектора,  части изображения с помощью линейно проекции \textquotedblrightвытягивают\textquotedblright из 2D в 1D.
-Также как и в задачах обработки текста, на вход в модель кроме патчей надо подать positional encoders.
-Итоговые преобразования изображения выглядят так: \\
-
-\noindent$z_{0} = [x_{class}; x^1_{p} \cdot E; x^2_{p} \cdot E; \dots x^N_{p} \cdot E] + E_{pos} $,\\
-$x \in \mathds{R}^{H \times W \times C}$, $E \in \mathds{R}^{N \times (P^{2} \times C)}$,
-где $C$ - кол-во каналов, $(P, P)$ - размер каждого патча, $N = HW / P^2$
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=1]{img/vit/patches.png}
-    \caption{Подготовка патчей}
-\end{figure}
-
-В качестве альтернативы патчам изображения, входная последовательность может быть сформирована
-из карт характеристик (feature map) CNN.
-В этой гибридной модели проекция E применяется к участкам, извлеченным из карты объектов CNN, как частный случай, патчи могут иметь пространственный размер 1x1.
-Патч с классом изображения и positional embeddings добавляются, как описано выше.
-
-\subsection{Query-Key-Value вычисления}
-
-TimeSFormer состоит из L блоков энкодера. В каждом блоке $l$, как и в оригинальном трансформере вычисляются Query, Key, Value для каждого патча из его прошлого состояния $z_{(p, t)}^{(l-1)}$
-
-\begin{center}
-    $q_{(p, t)}^{(l, a)} = W_{Q}^{(l, a)} \cdot LN(z_{(p, t)}^{(l-1)}) \in \mathds{R}^{D_{h}}$,
-
-    $q_{(p, t)}^{(l, a)} = W_{K}^{(l, a)} \cdot LN(z_{(p, t)}^{(l-1)}) \in \mathds{R}^{D_{h}}$,
-
-    $q_{(p, t)}^{(l, a)} = W_{V}^{(l, a)} \cdot LN(z_{(p, t)}^{(l-1)}) \in \mathds{R}^{D_{h}}$,
-\end{center}
-где $LN()$ означате LayerNorm, $a = 1, \dots, A$ - индексы голов в Multi-head attention.
-
-\subsection{Attention weights}
-Вычисление self-attention коэффициентов для $a_{(p, t)}^{(l, a)}$ для патча $(p, t)$ вычисляется как:
-
-\begin{center}
-    $a_{(p, t)}^{(l, a)} = SoftMax(\frac{q_{(p, t)}^{(l, a)}}{\sqrt{D_{h}}}^{T} \cdot \left[k_{(0, 0)}^{(l, a)} {k_{(p', t')}^{(l, a)}}_{p' = 1, \dots, N, t' = 1, \dots, F} \right])$
-\end{center}
-
-\subsection{Encoding}
-
-Закодированное представление $z_{(p, t)}^{(l)}$ в блоке $l$ получается путем вычисления взвешенной суммы векторов значений с использованием коэффициентов self-attention от каждой головы multi-head attention:
-
-\begin{center}
-    $s_{(p, t),(0, 0)}^{(l, a)} = a_{(p, t),(0, 0)}^{(l, a)} \cdot v_{(0, 0)} + \sum\limits_{p'=1}^N \sum\limits_{t'=1}^F a_{(p, t),(p', t')}^{(l, a)} \cdot a_{(p', t')}^{(l, a)}$
-\end{center}
-
-Затем векторы каждого головы конкатенируются, проецируется и передается через MLP, используя residual connections после каждой операции:
-
-\begin{center}
-    $z'_{(p, t)}^{(l)} = W_{O} \cdot \begin{bmatrix}
-            s_{(p, t)}^{(l, 1)} \\
-            \vdots              \\
-            s_{(p, t)}^{(l, A)}
-        \end{bmatrix} + z_{(p, t)}^{(l - 1)}$
-
-    $z_{(p, t)}^{(l)} = MLP(LN(z'_{(p, t)}^{(l)})) + z'_{(p, t)}^{(l)}$
-\end{center}
-
-Для финальной классификации используется MLP с 1 скрытым слоем. \\
-
-Ниже представлен блок энкодера в TimeSFormer:
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=.9]{img/vit/timesformer.png}
-    \caption{Энкодер блок в TimeSFormer}
-\end{figure}
+\section{Модель RoBERTa}
 
 \newpage
-\section{Модель RoBERTa}
-
-\subsection{BERT}
-
-BERT был представлен в статье BERT: Pre-training of Deep Bidirectional Transformers for Language Understanding\cite{bert:2018}.
-Архитектура модели BERT - это энкодер Transformer.
-По новому подошли к процессу обучения и к тому, как BERT используется для последующих задач.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.8]{img/bert/bert.png}
-    \caption{Архитектура BERT}
-\end{figure}
-
-\begin{center}
-    \textbf{Новый способ обучения}
-\end{center}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.6]{img/bert/bert_input.png}
-    \caption{Входные данные}
-\end{figure}
-
-Входные данные для обучения: пары предложений со  разделенных специальным токеном-разделителем [SEP] (рис 2.15).
-Чтобы модель могла легко различать эти предложения, в дополнение к токеновым и позиционным вложениям она использует вложения сегментов.
-Также BERT имеет 2 цели обучения, для которых используются некоторые специальные токены([MASK], [CLS]).
-\\
-
-\begin{center}
-    \textbf{Первая цель обучения обучения: Masked Language Modeling (MLM)}
-\end{center}
-Наиболее важной из двух целей обучения является цель моделирования замаскированного языка (MLM). На этапе обучения для MLM происходит следующее:
-\begin{itemize}
-    \item Выбирается несколько токенов (каждый токен выбирается с вероятностью 15\%)
-    \item Выбранные токены заменяются на [MASK] с вероятности 80\%, случайный токен - 15\% и остается без изменений - 10\%
-    \item Модель должна предсказать исходный токен
-\end{itemize}
-
-На рисунке ниже показан пример шага обучения для одного предложения.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.18]{img/bert/bert_tasks.png}
-    \caption{Masked Language Modeling (MLM)}
-\end{figure}
-\newpage
-\begin{center}
-    \textbf{Вторая цель обучения обучения: Next Sentence Prediction (NSP)}
-\end{center}
-
-Задача «Предсказание следующего предложения» (NSP) представляет собой задачу бинарной классификации.
-На последнем уровне модели, исходя из токена [CLS], модель предсказывает, являются ли два предложения последовательными предложениями в некотором тексте или нет.
-В датасете для обучения, представленном в оригинальной статье сказанно, что при обучении 50\% примеров содержат последовательные предложения, извлеченные из обучающих текстов, а еще 50\% — случайную пару предложений.
-
-Это задание учит модель понимать отношения между предложениями и позволяет BERT решать более сложные задачи.
-
-\begin{center}
-    Пример из исходного датасета:
-\end{center}
-
-Ввод: [CLS] мужчина пошел в [MASK] магазин [SEP] он купил галлон [MASK] молока [SEP]
-
-Метка: Последовательные \\
-
-Ввод: [CLS] мужчина пошел в магазин [MASK] [SEP] пингвин [MASK] летают \#\#less birds [SEP]
-
-Метка: Непоследовательные\\
-
-Следует также упомянуть, что BERT подходит для решения разных задач, ниже приведены основные типы задач:
-
-\begin{itemize}
-    \item Классификация отдельных предложений
-          Для классификации отдельного предложения, вводятся токены предложения и [CLS] с классом предложения.
-    \item Классификация пар предложений
-          Для классифицикации пары предложений, вводятся токены предложений, как в обучении, и и [CLS] с классом предложения.
-    \item Ответ на вопрос (Question answering)
-          В задачах вопросов на ответы, на вход дают тект и вопрос. Ответ на этот вопрос содержится в исходном тексте, и задача состоит в том, чтобы найти отрывок из текста с ответом.
-    \item Тэгирование слов в предложении
-          В задачах тегирования прогнозируются теги для каждого токена.
-Например, в распознавании именованных объектов (NER) вы должны предсказать, является ли слово именованным объектом и его типом (например, местоположение, человек и т.
-д.).
-\end{itemize}
-
-\subsection{RoBERTa - улучшенная версия BERT}
-
-Модель RoBERTa была предложена в книге RoBERTa: A Robustly Optimized BERT Pretraining Approach\cite{roberta:2019}.
-Он основан на модели Google BERT, выпущенной в 2018 году.
-Она основан на BERT и изменяет ключевые гиперпараметры, удаляя цель предварительного обучения в следующем предложении и обучая с гораздо большими мини-батчами и learning rate.
-\\
-
-Список основных изменений представлен ниже:
-\begin{itemize}
-    \item Обучение большем объем данных (в 10 раз больше) и дольше по времени.
-    \item Входные данные представляют собой более длинные последовательности, при оставшемся ограничении в 512 токенов, как и у BERT
-    \item Динамическое маскирование токенов: токены маскируются по-разному в каждую эпоху, тогда как BERT делает это раз и навсегда
-    \item Использование большого батча (относительно исходного BERT с батчем 256) при обучении модели
-    \item Использование Byte-Pair Encoding с байтами в качестве субъединицы, а не с символами (из-за символов юникода)
-\end{itemize}
-
-\newpage
-\section{Модель Wave2vec 2.0}
-
-Фреймворк wave2vec 2.0\cite{w2v:2020} разработан для self-supervised обучения эмбеддингов на основе необработанных аудиоданных.
-Модель кодирует речевой звук с помощью многослойной сверточной нейронной сети, а затем маскирует участки результирующих скрытых речевых представлений, аналогично моделированию замаскированного языка (MLM), который был разобран в разделе про BERT.
-Скрытые представления передаются в Трансформер для построения контекстуализированных эмбеддингов, и модель обучается с помощью contrastive learning.
-
-Как показано на рисунке ниже, модель обучается в два этапа.
-Первый этап проходит в режиме самоконтроля, который выполняется с использованием немаркированных данных и направлен на достижение наилучшего возможного представления речи.
-Вы можете думать об этом так же, как вы думаете о встраивании слов.
-Эмбеддинги слов также направлено на достижение наилучшего представления естественного языка.
-Основное отличие заключается в том, что Wav2Vec 2.0 обрабатывает аудио вместо текста.
-
-Вторая фаза обучения - это контролируемая тонкая настройка, в ходе которой помеченные данные используются для обучения модели предсказывать определенные слова или фонемы (наименьшая возможная звуковая единица в определенном языке).
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.8]{img/wave2vec/w2v_stages.png}
-    \caption{Этабы обучения wave2vec}
-\end{figure}
-
-Архитектура окончательной модели, используемой для прогнозирования, состоит из трех основных частей:
-
-\begin{itemize}
-    \item Cверточные слои, которые обрабатывают входные данные необработанной формы сигнала для получения скрытого представления - Z,
-    \item Трансформер из n слоев, создающий эмбеддинги с контекстом - C,
-    \item линейная проекция на выходные данные - Y
-\end{itemize}
-
-\subsection{Контрастивное обучение}
-Контрастивное обучение - это концепция, в рамках которой входные данные преобразуются двумя различными способами.
-После этого модель обучается распознавать, являются ли два преобразования входных данных по-прежнему одним и тем же объектом.
-В Wav2Vec 2.0 трансформаторные слои являются первым способом преобразования, второй осуществляется путем квантования, что будет объяснено в дальнейшей части этой статьи.
-Более формально, для замаскированного скрытого представления $z_{t}$ мы хотели бы получить такое контекстное представление $c_{t}$, чтобы иметь возможность угадать правильное квантованное представление $q_{t}$ среди других квантованных представлений.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.25]{img/wave2vec/wave2vec.png}
-    \caption{Входные данные}
-\end{figure}
-
-\subsection{Квантизация}
-Квантизация - это процесс преобразования значений из непрерывного пространства в конечный набор значений в дискретном пространстве.
-
-У нас есть вектор представления скрытой речи $z_{t}$ охватывает несколько фонем.
-Число фонем в языке конечно.
-Более того, число всех возможных пар фонем конечно.
-Это означает, что они могут быть идеально представлены одним и тем же скрытым речевым представлением.
-Кроме того, их число конечно, поэтому мы можем создать "кодовую книгу", содержащий все возможные пары фонем.
-Затем квантование сводится к выбору правильного кодового слова из "кодовую книгу".
-Однако вы можете себе представить, что количество всевозможных звуков огромно.
-Чтобы упростить обучение и использование, авторы Wav2Vec 2.0 создали G кодовых книг, каждая из которых состоит из V кодовых слов.
-Чтобы создать квантизированное представление, следует выбрать лучшее слово из каждой кодовой книги.
-Затем выбранные векторы объединяются и обрабатываются линейным преобразованием для получения квантизированное представления.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.7]{img/wave2vec/w2v_books.png}
-    \caption{Выбор токена их кодовых книг}
-\end{figure}
-
-Выбор правильного кодовго представления выбирается с помощью Gumbel softmax:
-
-\begin{center}
-    $p_{g, v} = \frac{exp{\frac{sim(l_{g, v} + n_{v})}{\tau}}}{\displaystyle\sum_{k=1}^{V} exp{\frac{l_{g, k} + n_{k}}{\tau}}}$,
-    где
-    \begin{itemize}
-        \item sim — косинусное сходство,
-        \item l — логиты, полученные свертками,
-        \item $n_{k} = -log(-log(gl))$,
-        \item $u_{k}$ - выборки из равномерного распределения U(0, 1),
-        \item $\tau$ - температура
-    \end{itemize}
-\end{center}
-
-Поскольку это задача классификации, функция softmax представляется естественным выбором для выбора наилучшего кодового слова в каждой кодовой книге.
-Gumbel softmax поставляется с двумя улучшениями: рандомизацией и температурой $\tau$.
-Благодаря рандомизации модель более охотно выбирает разные кодовые слова во время обучения, а затем обновляет их веса.
-Важно, особенно в начале обучения, не допускать использования только подмножества кодовых книг.
-
-\subsection{Diversity loss (Потеря разнообразия)}
-
-Потеря разнообразия - это своего рода метод регуляризации.
-Авторы установили G=2 кодовых книги с V=320 кодовыми словами в каждой кодовой книге.
-Теоретически это дает $320 \cdot 320 = 102400$ возможных квантованных представлений.
-Но в процессе обучения можно  столкнутся с проблемой того, что модель использовать все эти возможности.
-Она научится использовать, например, только 100 кодовых слов из каждой кодовой книги и растратит весь потенциал кодовой книги впустую.
-Вот почему потеря разнообразия может быть полезной.
-Он основан на энтропии, которая может быть вычислена по следующей формуле:
-
-\begin{center}
-    $H(x) = - \displaystyle\sum_{x} P(x)\cdot log(P(x))$, где
-\end{center}
-
-$x$ — возможный результат действия дискретной случайной величины $X$,
-
-$P(x)$ — вероятность события $x$. \\
-
-Энтропия принимает максимальное значение, когда распределение данных является равномерным.
-В случае wave2vec это означает, что все кодовые слова используются с одинаковой частотой.
-С помощью этого производится вычисление энтропию каждой кодовой книги по всей серии обучающих примеров, чтобы проверить, используются ли кодовые слова с одинаковой частотой.
-Максимизация этой энтропии побудит модель использовать преимущества всех кодовых слов.
-Максимизация равна минимизации отрицательной энтропии, которая является потерей разнообразия.\\
-
-\begin{center}
-    $L_{d} = \frac{1}{GV} \cdot (-H(\overline{p}_{g})) = \frac{1}{GV}\displaystyle\sum_{g=1}^{G}\displaystyle\sum_{v=1}^{V} \overline{p}_{g, v} \cdot log(\overline{p}_{g, v})$
-\end{center}
-
-\newpage
-\chapter{Практическая часть}
+\chapter{Experiments}
 

### 2024-04-19 src/segmentation/neural_texttiling.py
update segment code

@@ -1,8 +1,8 @@
-import segeval
-from sklearn.metrics import f1_score
 import torch
+import segeval
+
+from sklearn.metrics import f1_score
 import numpy as np
-from sentence_transformers import SentenceTransformer
 
 
 def similarity_computing(texts, tokenizer, text_encoder, mode, device):
@@ -49,7 +49,6 @@ def similarity_computing(texts, tokenizer, text_encoder, mode, device):
 
 
 def depth_computing(scores):
-
     num_scores = len(scores)
     depth_scores = []
 
@@ -302,7 +301,7 @@ def get_loss_for_each_utterance(words, losses):
     return utterances, loss_for_each_u
 
 
-def TextTiling_glm(dialogue, reference, text_decoder, tokenizer, threshold, device):
+def text_tiling_llm(dialogue, reference, text_decoder, tokenizer, threshold, device):
 
     dialogs = combine_near_two_utterances([dialogue])
 

### 2024-04-19 src/segmentation/scripts/bert_segmentation.sh

@@ -1,7 +1,7 @@
-DATASET=./datasets/eval/dialseg_711.json
-ENCODER_BERT_MODEL=dse-roberta-base
+DATASET=./datasets/eval/superseg.json
+ENCODER_BERT_MODEL=FacebookAI/roberta-large
 
-python ./src/segmentation_module/segment_bert.py \
+python ./src/segmentation/segment_bert.py \
 --dataset $DATASET \
 --text_encoder $ENCODER_BERT_MODEL \
---mode NSP
+--mode SC

### 2024-04-19 src/segmentation/scripts/llm_segmentation.sh

@@ -1,6 +1,6 @@
 #!/bin/bash
-DATASET=./datasets/eval/dialseg_711.json
-ENCODER_LLM_MODEL=microsoft/DialoGPT-medium
+DATASET=./datasets/eval/superseg.json
+ENCODER_LLM_MODEL=microsoft/DialoGPT-large
 
 python ./src/segmentation/segment_llm.py \
 --dataset $DATASET \

### 2024-04-19 src/segmentation/segment_bert.py

@@ -56,7 +56,8 @@ if __name__ == "__main__":
     text_encoder_name = args.text_encoder
     mode = args.mode
     device = 'cuda' if torch.cuda.is_available() else 'cpu'
-
+    print('[INFO] Device:', device)
+    
     # load encoder model
     if mode == 'SC':
         tokenizer = AutoTokenizer.from_pretrained(text_encoder_name)
@@ -81,31 +82,31 @@ if __name__ == "__main__":
             test_data.append(dialogue)
             
     # Hyper-parameter (alpha) search on dev set
-    best_alpha, best_pk = alpha_search(
-        dev_data, text_encoder, tokenizer, mode, device, -2, 2, 0.1)
-    print('[INFO] The loaded text encoder is: ', text_encoder_name)
+    #best_alpha, _ = alpha_search(
+    #    dev_data, text_encoder, tokenizer, mode, device, -2, 2, 1)
+    best_alpha = 1
     print('[INFO] The best hyper-parameter (alpha): ', best_alpha)
 
-    # Evaluation on test set
     total_pk = 0
     total_wd = 0
     total_f1 = 0
     num_samples = len(test_data)
     for i, dialogue in tqdm(enumerate(test_data), total=len(test_data), desc='Evaluating TEST set'):
-        pk, wd, f1, pred_segments = TextTiling(
-            dialogue['utterances'], dialogue['segments'], text_encoder, tokenizer, best_alpha, mode, device)
+        pk, wd, f1, pred_segments = TextTiling(dialogue['utterances'], 
+                                               dialogue['segments'], 
+                                               text_encoder, 
+                                               tokenizer,
+                                               best_alpha,
+                                               mode, 
+                                               device)
         total_pk += pk
         total_wd += wd
         total_f1 += f1
 
-    # Compute the mean scores
     mean_pk = total_pk / num_samples
     mean_wd = total_wd / num_samples
     mean_f1 = total_f1 / num_samples
 
-    # Print or return the mean scores
-    print('-----------------------------------')
     print(f"Mean P_k score: {mean_pk}")
     print(f"Mean WindowDiff score: {mean_wd}")
     print(f"Mean F1 score: {mean_f1}")
-    print('-----------------------------------')

### 2024-04-19 src/segmentation/segment_llm.py

@@ -5,7 +5,7 @@ import torch
 import numpy as np
 from transformers import AutoModelForCausalLM, AutoTokenizer
 
-from neural_texttiling import TextTiling_glm
+from neural_texttiling import text_tiling_llm
 
 from tqdm import tqdm
 import warnings
@@ -27,11 +27,15 @@ def threshold_search(dialogue_data, text_decoder, tokenizer, device, lowerbound,
         num_samples = len(dialogue_data)
         
         for dialogue in dialogue_data:
-            pk, _, _, _ = TextTiling_glm(dialogue['utterances'], dialogue['segments'], text_decoder, tokenizer, threshold, device)
+            pk, _, _, _ = text_tiling_llm(dialogue['utterances'], 
+                                          dialogue['segments'], 
+                                          text_decoder, 
+                                          tokenizer, 
+                                          threshold, 
+                                          device)
             total_pk += pk
 
         mean_pk = total_pk / num_samples
-
         if mean_pk < best_pk:
             best_pk = mean_pk
             best_threshold = threshold
@@ -51,9 +55,8 @@ if __name__ == "__main__":
     data = args.dataset
     text_decoder_name = args.text_decoder
     device = 'cuda' if torch.cuda.is_available() else 'cpu'
-    print('Device:', device)
+    print('[INFO] Device:', device)
 
-    # load generative model
     text_decoder = AutoModelForCausalLM.from_pretrained(text_decoder_name).to(device)
     tokenizer = AutoTokenizer.from_pretrained(text_decoder_name)
 
@@ -62,24 +65,38 @@ if __name__ == "__main__":
     dev_data, test_data = [], []
     
     for dialogue in dialogue_data:
-        if dialogue['set'] == 'dev': dev_data.append(dialogue)
-        else: test_data.append(dialogue)
+        if dialogue['set'] == 'dev': 
+            dev_data.append(dialogue)
+        else: 
+            test_data.append(dialogue)
 
     ## Threshold search on dev set
-    best_threshold, best_pk = threshold_search(dev_data, text_decoder, tokenizer, device, 0, 5, 1)
-    print('[INFO] The loaded text decoder is: ', text_decoder_name)
-    print('[INFO] The best threshold: ', best_threshold)
+    #best_threshold, _ = threshold_search(dev_data, text_decoder, tokenizer, device, 0, 5, 1)
+    best_threshold = 4
[…]

### 2024-04-21 paper/main.tex
add transformers description

@@
 
-\section{Motivation}
+\section{Relevance of the topic}
 
@@
 The objectives of the thesis work are:
-\begin{itemize}
+\begin{enumerate}
     \item A review of existing approaches to dynamic text and dialogue topic modelling
@@
     \item Visual analysis of the pipeline operation on pre-selected audio data
+\end{enumerate}
+
+\chapter{Literature review}
+
+\section{Transformer architecture}
+
+Prior to 2017 (the publication of the original Transformer article), the main approach to working with sequences was to use Recurrent Neural Networks (RNN).
+However, this approach has several well-known drawbacks:
+
+First, RNN hold all information about the sequence in a hidden state that is updated with each step. 
+If the model needs to "remember" something that happened hundreds of steps ago, that information must be stored in a hidden state and not replaced with something new. 
+Therefore, you either have to have a very large hidden state, or accept the loss of information.
+
+Second, training recurrent networks is difficult to parallelize: to get the hidden state of the RNN layer for step $i$, you have to compute the state for step $i - 1$.
+
+Both problems make it difficult to apply RNN to really long sequences: even if you wait until the end of training, your model will somehow lose information about what was at the beginning of the text.
+It is necessary to have method to "read" the sequence so that at any point in time it is possible to refer to an arbitrary point in the past in constant time and without loss of information. 
+This is the self-attention mechanism underlying Transformer, introduced in "Attention is all you need" paper.
+
+\subsection{Self-attention}
+
+Self-attention represents a key component of the model in which tokens' information interact with each other.
+At each stage of the encoder, tokens exchange information with one another, gather contextual data, and refine their previous representations (Figure 2.3).
+
+For each input token trainable projections $W_Q$, $W_K$, $W_V$ obtained three representations interpreted by humans as a query ($q_i$), a key ($k_i$), and a value ($v_i$).
+\begin{itemize}
+    \item $q_i$ - query for other tokens about the information;
+    \item $k_i$ - keys that will be used to extract information about the token;
+    \item $v_i$ - values of information;
 \end{itemize}
 
-\chapter{Theoretical explanation}
+Next are the attention weights that will be applied to the values from which the important information will be extracted:
+\begin{center}
+    $Weights_{i} = softmax(\frac{q_i \cdot k_{1}^T}{C}, \frac{q_i \cdot k_{2}^T}{C}, \dots)$,
+\end{center}
+
+where $C$ is some normalization constant, $\sqrt[]{d}$ of the dimension of keys and values was taken as a normalization constant.
+
+The output of the self-attention layer is a weighted sum of values with coefficients in the form of attention weights. In vectorized form, you can write:
+
+\begin{center}
+    $Attention(q, k, v) = softmax(\frac{Q \cdot K^T}{\sqrt{d_k}}) \cdot V $
+\end{center}
+
+The figure below shows the complete algorithm for the interaction of Q, K, V in the self-attention mechanism.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=1]{img/query-key-value.png}
+    \caption{Query, Key, Value in Self-Attention mechanism}
+\end{figure}
+
+In the decodert part the structure of the attention mechanism is different. 
+They use Masked self-attention. In the form of attention described above, each token will "look" at the entire sequence, which is undesirable for the decoder. 
+Indeed, in the generation stage, we will generate one token per step, and access to subsequent steps in the training stage will lead to information leakage in the decoder and poor model quality. 
+To avoid this problem, you need to apply an autoregressive mask when training attention, by manually adjusting the
+weights for future tokens before softmax, so that their probabilities become zero after softmax. As you can see in the image below (source), this mask has a lower triangular appearance.
+
+\subsection{Multi-head attention}
+
+One set of $Q, K, V$ can reflect only one type of dependency between tokens, and matrices extract only a limited amount of information from input representations. 
+To compensate for this suboptimality, the authors of the architecture proposed an approach with multiple "heads" of attention (multi-head attention): 
+in fact, instead of one layer of attention, we use several parallel ones with different weights, and then aggregate the results. The figure below shows what multi-head attention looks like:
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.7]{img/multi-head.png}
+    \caption{Query, Key, Value in Self-Attention mechanism}
+\end{figure}
+
+\subsection{Feed forward and Add \& Norm}
+
+In addition to the attention blocks, the Transformer also contains feed-forward layers consisting of a linear layer and a ReLU activation function.
+The FFN layer can be expressed as follows: \\
+$FFN(x) = act(xW_{1} + b_{1}) \dot W_{2} + b_{2}$,
+
+Intermediate activations in FFN are different: it all started with the well-known ReLU, but at some point the community switched to GELU (Gaussian Error Linear Unit) with a formula $x\Phi(x)$, where $\Phi$ - is the distribution function of a standard normal random variable
+\\
+
+The second important layer is LayerNorm, which independently normalizes the vector representation of each sequence in the batch. 
+The vector representation of each token is normalized in the Transformer.
+In addition, LayerNorm has trainable parameters, $scale$ and $bias$, which are used after normalization to scale the output of the layer.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=1]{img/layer_norm.png}
+    \caption{LayerNorm слой}
+\end{figure}
+
+In Figure 2.8, note that $\mu_{k}$ and $\sigma_{k}$ are evaluated for each sequence, unlike $scale$ and $bias$, which are layer parameters.
+
+\subsection{Positional encoding}
+
+Since the model does not know the order of the input tokens, because all tokens are processed simultaneously, therefore we must explicitly show the model the position of the tokens. 
+To do this, we have two sets of attachments: for the tokens themselves and for their positions. 
+Then the input representation of the token is the sum of the two attachments.
+Positional encodings can be trained, but the authors of the original article found that the presence of corrected attachments does not degrade the quality of the model.
+The final formula for positional embeddings looks like this:
+
+\begin{center}
+    $PE_{pos,2i} = \sin(pos/10000^{2i/d_{model}})$, \\
+    $PE_{pos,2i+1} = \cos(pos/10000^{2i/d_{model}})$, \\
+\end{center}
+where $pos$ is a position and $i$ is a vector dimension.
+Each measure of position encoding corresponds to a sine wave, and the wavelengths form a geometric progression from $2\pi$ to 10,000$\cdot 2\pi$.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=1]{img/pos_encoding.png}
+    \caption{Positional encoding}
+\end{figure}
+
+A newer version of positional encoding is rotary embeddings.
+Rotary Position Embedding, or RoPE, is a type of position embedding that encodes absolute position information with a rotation matrix and naturally incorporates explicit relative position dependency into the self-attention formulation.
+In particular, RoPE has valuable properties such as the flexibility to be extended to arbitrary sequence lengths, decaying inter-token dependency with increasing relative distances, and the ability to augment linear self-attention with relative position encoding.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.7]{img/rotary_embs.png}
+    \caption{Rotary embeddings}
+\end{figure}
+
+\section{BERT}
+
+The BERT is a language model based on the transformer architecture, that was first introduced in the article BERT: Pre-training of Deep Bidirectional Transformers for Language Understanding\cite{bert:2018}.
+From an architectural point of view, BERT consists of 12 encoder blocks, including a Self-Attention mechanism, Normalization and Feed-forward layers.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.8]{img/bert.png}
+    \caption{BERT architecture}
+\end{figure}
+
+\subsection[short]{Training tasks}
+BERT has two pretrain objective described in article: \\
+The first objective of the training process is the masked language modelling (MLM). 
+During the training phase of MLM, the following occurs:
+
+\begin{itemize}
+    \item A number of tokens are selected with a probability of 15\% for each token.
+    \item The selected tokens are replaced by [MASK] with a probability of 80\%, a random token with a probability of 15\% and remain unchanged with a probability of 10\%.
+    \item The model must predict the original token.
+\end{itemize}
+
+Secondly, Next Sentence Prediction (NSP) task is a binary classification task. 
+In order to understand relationship between two sentences, BERT training process also uses next sentence prediction.
+During training the model gets as input pairs of sentences and it learns to predict if the second sentence is the next sentence in the original text as well. 
+The training dataset presented in the original paper states that when trained, 50\% of the examples contain related sentences extracted from the training texts, while the remaining 50\% contain a random pair of sentences.
+
+\section{Large Language models}
+
+\subsection{LLaMA}
+
+\subsection{Mistral}
 
@@
 The dialogue segmentation stage represents the initial stage for textual type of data. 
-At this stage, a set of strings is input, with each line representing an utterance from one dialogue participant. 
 The result is a set of utterance indices that indicate the beginning of a thematically related group.
@@
 In accordance with the majority of previous studies, we have utilise the best open-source solution based on the presented in papers metrics. 
+It is algorithm from Improving Unsupervised Dialogue Topic Segmentation with Utterance-Pair Coherence Scoring by Linzi Xing and Giuseppe Carenini (2020)
 Let us now examine the algorithm in greater detail.
-
-The input of stage is a set of utterances $\{u1, u2, ..., uk\}$, splited into pairs of consecutive utterance. 
-All utterances text encode by encoder model.
-
-\subsection{BERT for embeddings extraction}
-
-The BERT model is a Transformer encoder that was first introduced in the article BERT: Pre-training of Deep Bidirectional Transformers for Language Understanding\cite{bert:2018}.  
-This article presents a new approach to the learning process and how BERT is used for downstream tasks.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.8]{img/bert.png}
-    \caption{BERT architecture}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.6]{img/bert_input.png}
-    \caption{}
-\end{figure}
-
-The input data for training comprises pairs of sentences separated by a special token separator [SEP] (Figure 2.15). 
-In order to facilitate the model's ability to distinguish between these sentences, it employs segment embeddings in addition to token and position embeddings. 
-BERT also has two learning objectives for which some special tokens ([MASK], [CLS]) are used.
-
-The first objective of the training process is the masked language modelling (MLM). 
-During the learning phase of MLM, the following occurs:
-
-\begin{itemize}
-    \item A number of tokens are selected with a probability of 15\% for each token.
-    \item The selected tokens are replaced by [MASK] with a probability of 80\%, a random token with a probability of 15\% and remain unchanged with a probability of 10\%.
-    \item The model must predict the original token.
-\end{itemize}
-
-The figure below illustrates an example of a training step for a single sentence.
-
-The Next Sentence Prediction (NSP) task is a binary classification task. 
-In order to understand relationship between two sentences, BERT training process also uses next sentence prediction.
-During training the model gets as input pairs of sentences and it learns to predict if the second sentence is the next sentence in the original text as well. 
-The training dataset presented in the original paper states that when trained, 50\% of the examples contain related sentences extracted from the training texts, while the remaining 50\% contain a random pair of sentences.
 
@@
 Here, the contextualised representation of $[CLS]$ is employed as the topic-aware embedding to predict the degree of matching between the two input utterances in terms of topic. 
-The topical coherence score is estimated by passing the $[CLS]$ representation through another MLP.
+The topic coherence scores are estimated by passing the $[CLS]$ representation through another MLP.
 
@@
 
+The input of stage is a set of utterances $\{u1, u2, ..., uk\}$. 
 Subsequently, an utterance-pair coherence scoring model is applied to $k - 1$ consecutive pairs, resulting in a sequence of coherence scores $\{c1, c2, ..., ck-1\}$. 
@@
 \begin{center}
-    $D_{pi} = \frac{ h_{l}(i) + h_{r}(i) - 2 \cdot c_{i}}{2}$. 
+    $D_{pi} = \frac{h_{l}(i) + h_{r}(i) - 2 \cdot c_{i}}{2}$. 
 \end{center}
@@
 
-\subsection{Methods overview}
-
-\subsection{Large Language Models}
-
 \subsection{LLM for topic extraction}
@@
 \section{Topic Evolution over time}
-
-\section{Attention mechanism}
-
-Механизм внимания является частью нейронной сети.
-На каждом этапе декодирования он решает, какие части входных данных важнее всего.
-При этом кодировщику не нужно сжимать весь исходный код в один вектор - он создает представления для всех токенов источника.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.3]{img/transformer/attention_mech1.png}
-    \caption{Схема работы механизма внимания}
-\end{figure}
-
-\newpage
-Алгоритм работы механизма внимания в encoder-decoder модели приведен ниже.
-
-\begin{itemize}
-    \item На каждом этапе декодирования механизм внимания получает на вход: состояние декодера $h_t$ и все состояния кодировщика $s_1, s_2, \dots, s_m$
-    \item Вычисляет коэффициенты внимания (attention scores) для каждого состояния кодеровщика $s_k$ механизм внимания вычисляет свою "релевантность" для этого состояния декодера $h_t$. \\
-          Для вычисления можно применить любую функцию, однако, есть несколько популярных и простых вариантов, которые работают довольно хорошо.
-          \begin{itemize}
-              \item $scores(h_t, s_k) = h_t^T \cdot s_k $
-              \item $scores(h_t, s_k) = h_t^T \cdot W \cdot s_k $, где $W$ - матрица (обучаемый параметр)
-              \item $scores(h_t, s_k) = w_2^T \cdot \tanh(W \cdot [h_t^T, s_k]) $, где $W, w_2$ - матрицы (обучаемые параметры), а $[h_t^T, s_k]$ означает конкатенацию векторов
-          \end{itemize}
-
-    \item Вычисляет веса внимания (attention weights): распределение вероятностей - softmax, применяемое к показателям внимания;
-    \item Вычисляет вывод внимания : взвешенная сумма состояний декодера с весами внимания.
-\end{itemize}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=1]{img/transformer/self-attention.png}
-    \caption{Схема работы self-attention механизма}
-\end{figure}
-
-\subsection{Self-attention}
-
-Self-attention - является одним из ключевых компонентов модели, в которой токены взаимодействуют друг с другом.
-На каждом шаге в энкодере токены обмениваются информацией между собой, собирает контекст и обновляет предыдущее представление себя (рис 2.3).
-Такой подход полностью исключает последовательную обработку, как в RNN, что ускоряет работу нейронной сети и дает возможность обрабатывать каждое слово параллельно.
-
-\newpage
-\subsection{}
-
-Каждый входной токен в self-attention получает три представления, соответствующие ролям, которые он может играть:
-\begin{itemize}
-    \item query - запрашивает информацию;
-    \item key - показывает, какая информация есть у него;
-    \item value - дает запрашиваемую информацию;
-\end{itemize}
-
-От слова, которое мы сейчас подаем на вход (центральное), требуется query вектор, этим вектором мы запрашиваем у других информацию о текущем слове в данном контексте.
-От каждого другого слова нам будет нужен вектор key, который показывает, какую информацию данное слово может сказать о смысле центрального слова, а также вектор value, из которого мы и получим нужную нам информацию.
-
-Получив информацию от каждого вектора, мы подаем это на вход softmax функции и получаем итоговый эмбеддинг центрального слова.
-На рисунке нижен представлен полный алгоритм взаимодействия Q, K, V в self-attention механизме.
-
-\newpage
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=1]{img/transformer/query-key-value.png}
-    \caption{Query, Key, Value в self-attention механизма}
-\end{figure}
-
-Формула для вычисления выходных данных внимания следующая:
-
-\begin{center}
-    $Attention(q, k, v) = softmax(\frac{q \cdot k^T}{\sqrt{d_k}}) \cdot v $
-\end{center}
-
-\subsection{Masked self-attention}
-
-Для энкодера и декодера используются разные способы получения токенов, если энкодер получает все токены сразу и каждый токен может смотреть на любой другой токен во входном предложении, в декодере, мы генерируем по одному токену за раз: во время генерации мы не знаем, какие токены мы будем генерировать в дальше.
-Чтобы запретить декодеру заглядывать вперед, модель использует masked-self attention, который видит токены, только стоящие до него во входной последовательности.
-
-Работа masked self-attention отличается во время обучения и инференса модели.
-Во время второго декодер не заглядывает вперед - так как, мы не знаем, что будет дальше.
-Но при обучении используются уже известные нам данные.
-Поэтому при обучении на вход декодеру подется все целевое предложение целиком - без масок.
-
-На рисунке нижне приведен алгоритм работы masked self-attention механизма.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.9]{img/transformer/masked self-attention.png}
-    \caption{Схема работы masked Self-attention механизма}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=1]{img/transformer/multi-head.png}
-    \caption{Схема Multi-head attention}
-\end{figure}
-
-\subsection{Multi-Head Attention}
-
-Обычно понимание роли слова в предложении требует понимания того, как оно связано с разными частями предложения. Это важно не только при обработке исходного предложения, но и при генерации целевого.
-Поэтому мы должны позволить модели сосредоточиться на разных вещах: для этого сделали multi-head attention.
-Вместо одного attention механизма multi-head attention имеет несколько «голов», которые работают независимо.
-
-\begin{center}
-    $MultiHead(Q, K, V) = Concat(head_{1}, head_{2}, \dots, head_{n})$ \\
-    $head_{i} = Attention(QW^{i}_{Q}, KW^{i}_{K}, VW^{i}_{V})$
-\end{center}
-
-Формально это реализовано в виде нескольких механизмов внимания, результаты которых объединяются. (рис 2.6)
-
-\newpage
-\section{Трансформер: Архитектура модель}
-
-Трансформер - это модель, основанная только на механизме винмания и оказавшая огромное влияние на NLP в частности и машинном обучении в целом.
-Впервые представленная в статье Attention is All You Need\cite{allyouneed:2017} в 2017 году.
-Модель сразу показала лучшие результаты в задаче перевода, в сравнении с encoder-decoder модели.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.2]{img/transformer/transformer.png}
-    \caption{Схема архитектуры Траснформер}
-\end{figure}
-
-Трансформер состоит из частей, которое ранне описывались.
-Encoder состоит из multi-head self-attention механизмов и генерирует представление слов, а decoder из masked self-attention механизма, который уже генерирует новую последовательность.
-Это происходит в несколько слоев, обычно 6.
-
-Теперь рассмотрими другие составляющие архитекутры:\\
-
-\subsection{Feed forward and Add $\&$ Norm}
-
-Кроме блоков внимания, в Трансформере также присутствуют feed forward слои, которые состоят из линейного слоя и фукнции активации ReLU.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=1]{img/transformer/feed_forward.png}
-    \caption{Feed-forward слой}
-\end{figure}
-
-Вторым важным слоем является LayerNorm, которой независимо нормализует векторное представление каждой последовательности в батче.
-В Трансформере нормализуются векторное представление каждого токена.
-Кроме того, LayerNorm имеет обучаемые параметры, $scale$ и $bias$, которые используются после нормализации для масштабирования выходных данных слоя.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=1]{img/transformer/layer_norm.png}
-    \caption{LayerNorm слой}
-\end{figure}
-
-На рис 2.8 стоит обратить внимание на то, что $\mu_{k}$ и $\sigma_{k}$ оцениваются для каждой последовательности, в отличии от $scale$ и $bias$, которые являются параметрами слоя.
-
-\subsection{Positional encoding}
-
-Поскольку Transformer не знает порядок входных токенов, т.к все токены обрабатываются одновременно, поэтому мы должны явно показать модели положения токенов.
-Для этого у нас есть два набора вложений: для самихтокенов и для их позиций.
-Тогда входное представление токена является сумма двух вложений.
-Positional encodings можно обучить, но авторы оригинальной статьи обнаружили, что наличие исправленных вложений не ухудшает качество модели. Итоговая формула для позиционные вложения выглядят так:
-\begin{center}
-    $PE_{pos,2i} = \sin(pos/10000^{2i/d_{model}})$, \\
-    $PE_{pos,2i+1} = \cos(pos/10000^{2i/d_{model}})$, \\
-\end{center}
-
-где $pos$ это позиция и $i$ является векторным измерением.
-Каждое измерение позиционного кодирования соответствует синусоиде, а длины волн образуют геометрическую прогрессию от $2\pi$ до 10000 $\cdot 2\pi$.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=1]{img/transformer/pos_encoding.png}
-    \caption{Positional encoding}
-\end{figure}
-
-\subsection{Residual connection}
-
-Как и в сверточных сетях, в Трансформере применяется residual connections.
-Смысл residual connection очень прост, добавляет входные данные к выходным данным модели, но в то же время это очень полезная вещь: он борется с проблемой затухания градиента по сети и позволяют использовать много слоев.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.9]{img/transformer/res_block.png}
-    \caption{Residual connection}
-\end{figure}
-
-В трансформаторе остаточные соединения используются после каждого блока внимания и полносвязного блока.
-
-\newpage
-
-\section{Модель RoBERTa}
 

### 2024-04-22 paper/main.tex
fix cites and reduce percent of plagiarism

@@
 \begin{enumerate}
-    \item A review of existing approaches to dynamic text and dialogue topic modelling
+    \item Research of existing solutions for dynamic text and dialogue topic modelling
     \item The definition of the main stages for DTM for audio dialogues
@@
     \item The development of the architecture of all stages and their interconnections
-    \item The identification of key indicators to assess the quality of each stage is also required
+    \item Selection of appropriate metrics and measurement of the quality of the approaches used at each stage
     \item Visual analysis of the pipeline operation on pre-selected audio data
@@
 
-Prior to 2017 (the publication of the original Transformer article), the main approach to working with sequences was to use Recurrent Neural Networks (RNN).
-However, this approach has several well-known drawbacks:
+The main architecture for working with text was Recurrent Neural Networks. However, this approach has several well-known disadvantages:
 
-First, RNN hold all information about the sequence in a hidden state that is updated with each step. 
-If the model needs to "remember" something that happened hundreds of steps ago, that information must be stored in a hidden state and not replaced with something new. 
-Therefore, you either have to have a very large hidden state, or accept the loss of information.
+First of all, RN processes tokens and updates the hidden state sequentially, with this approach, the model begins to forget what happened before or it has to have a huge hidden state.
+Secondly, it is very difficult to start distributed training of recurrent networks, because the hidden state at each step must be considered separately.
 
-Second, training recurrent networks is difficult to parallelize: to get the hidden state of the RNN layer for step $i$, you have to compute the state for step $i - 1$.
+In order to avoid the issues previously outlined, the article "Attention is all you need" \cite{attention:2017} presented the Transformer architecture and its principal component, 
+the Attention Mechanism. The objective of this methodology is to permit the elements of the sequence to be accessed at any time, thereby reducing the loss of information.
 
-Both problems make it difficult to apply RNN to really long sequences: even if you wait until the end of training, your model will somehow lose information about what was at the beginning of the text.
-It is necessary to have method to "read" the sequence so that at any point in time it is possible to refer to an arbitrary point in the past in constant time and without loss of information. 
-This is the self-attention mechanism underlying Transformer, introduced in "Attention is all you need" paper.
-
-\subsection{Self-attention}
+\subsection{Self-attention mechanism}
 
@@
 \begin{center}
-    $Weights_{i} = softmax(\frac{q_i \cdot k_{1}^T}{C}, \frac{q_i \cdot k_{2}^T}{C}, \dots)$,
+    $Attn_Weights_{i} = softmax(\frac{q_i \cdot k_{1}^T}{C}, \frac{q_i \cdot k_{2}^T}{C}, \dots)$,
 \end{center}
 
-where $C$ is some normalization constant, $\sqrt[]{d}$ of the dimension of keys and values was taken as a normalization constant.
+where $C$ is some normalization constant.
 
-The output of the self-attention layer is a weighted sum of values with coefficients in the form of attention weights. In vectorized form, you can write:
+The Self-Attention result is a weighted sum of Values with coefficients in the form of attention weights. In vectorized form, you can write:
 
@@
 \end{center}
+where $\sqrt[]{d}$ constant of the dimension of $K$ and $V$.
 
-The figure below shows the complete algorithm for the interaction of Q, K, V in the self-attention mechanism.
+The figure below shows the complete algorithm for the interaction of Q, K, V.
 
@@
 They use Masked self-attention. In the form of attention described above, each token will "look" at the entire sequence, which is undesirable for the decoder.
-Indeed, in the generation stage, we will generate one token per step, and access to subsequent steps in the training stage will lead to information leakage in the decoder and poor model quality. 
-To avoid this problem, you need to apply an autoregressive mask when training attention, by manually adjusting the
-weights for future tokens before softmax, so that their probabilities become zero after softmax. As you can see in the image below (source), this mask has a lower triangular appearance.
+
+In the decoder, the generation of tokens occurs sequentially.
+If the attention mechanism has access to all tokens, not just those already used, this will result in information leakage in the decoder at the training stage and a reduction in the quality of the model.
+To avoid this problem, Masked Self-Attention is used, which masks tokens that follow the current one and reduces their probability in the softmax function to 0.
 
@@
 
-One set of $Q, K, V$ can reflect only one type of dependency between tokens, and matrices extract only a limited amount of information from input representations. 
-To compensate for this suboptimality, the authors of the architecture proposed an approach with multiple "heads" of attention (multi-head attention): 
-in fact, instead of one layer of attention, we use several parallel ones with different weights, and then aggregate the results. The figure below shows what multi-head attention looks like:
+To cover more information and dependencies, the authors propose the use of multiple levels of attention in parallel and the summarisation of the values obtained. 
+This is referred to as Multi-head attention. The figure below illustrates what a multi-head approach looks like:
 
@@
 
-\subsection{Feed forward and Add \& Norm}
+\subsection{Feed Forward Network}
 
-In addition to the attention blocks, the Transformer also contains feed-forward layers consisting of a linear layer and a ReLU activation function.
-The FFN layer can be expressed as follows: \\
-$FFN(x) = act(xW_{1} + b_{1}) \dot W_{2} + b_{2}$,
+Transformer block also contains feed-forward network consisting of a linear layer and a activation function.
+The FFN layer can be expressed as follows:
+\begin{center}
+    $FFN(x) = act(x \dot W_{1} + b_{1}) \dot W_{2} + b_{2}$, \\    
+\end{center}
+where $act$ is the activation function such as ReLU = $\max(0, x)$ or more contemporary GELU = $x\Phi(x)$.
 
-Intermediate activations in FFN are different: it all started with the well-known ReLU, but at some point the community switched to GELU (Gaussian Error Linear Unit) with a formula $x\Phi(x)$, where $\Phi$ - is the distribution function of a standard normal random variable
-\\
-
-The second important layer is LayerNorm, which independently normalizes the vector representation of each sequence in the batch. 
+In terms of LayerNorm, it is normalizes the vector representation of each sequence in the batch. 
 The vector representation of each token is normalized in the Transformer. 
@@
 
-In Figure 2.8, note that $\mu_{k}$ and $\sigma_{k}$ are evaluated for each sequence, unlike $scale$ and $bias$, which are layer parameters.
+\subsection{Positional encodings and Rotary embedding}
 
-\subsection{Positional encoding}
+A sequence of tokens is fed into the model, but the model itself does not understand in which order the tokens are located. 
+Therefore, positional embeddings are added to the tokens, which are designed to show the model the input order of the elements.
+Positional embeddings are given by formulas:
+\begin{center}
+    $PE_{position, 2i} = \sin(\frac{position}{10000^{2i/d_{model}}})$, \\
+    $PE_{position, 2i+1} = \cos(\frac{position}{10000^{2i/d_{model}}})$, \\
+\end{center}
 
-Since the model does not know the order of the input tokens, because all tokens are processed simultaneously, therefore we must explicitly show the model the position of the tokens. 
-To do this, we have two sets of attachments: for the tokens themselves and for their positions. 
-Then the input representation of the token is the sum of the two attachments.
-Positional encodings can be trained, but the authors of the original article found that the presence of corrected attachments does not degrade the quality of the model.
-The final formula for positional embeddings looks like this:
-
-\begin{center}
-    $PE_{pos,2i} = \sin(pos/10000^{2i/d_{model}})$, \\
-    $PE_{pos,2i+1} = \cos(pos/10000^{2i/d_{model}})$, \\
-\end{center}
-where $pos$ is a position and $i$ is a vector dimension.
-Each measure of position encoding corresponds to a sine wave, and the wavelengths form a geometric progression from $2\pi$ to 10,000$\cdot 2\pi$.
+Each measure of position encoding corresponds to a sin wave, and the wavelengths form a geometric progression from $2\pi$ to 10,000$\cdot 2\pi$.
 
@@
 A newer version of positional encoding is rotary embeddings.
-Rotary Position Embedding, or RoPE, is a type of position embedding that encodes absolute position information with a rotation matrix and naturally incorporates explicit relative position dependency into the self-attention formulation.
-In particular, RoPE has valuable properties such as the flexibility to be extended to arbitrary sequence lengths, decaying inter-token dependency with increasing relative distances, and the ability to augment linear self-attention with relative position encoding.
+Rotary position embedding represents a distinct approach to incorporating relative position information into the attention matrix. 
+This approach differs from other approaches in that it first multiplies queries and keys with a rotation matrix. This rotation matrix is a function of absolute position. 
+Calculating the inner products of rotated queries $W_Q$ and keys $W_K$ results in an attention matrix that is a function of relative position information only.
 
@@
 
-\subsection[short]{Training tasks}
+\subsection{Training tasks}
 BERT has two pretrain objective described in article: \\
@@
 
-Secondly, Next Sentence Prediction (NSP) task is a binary classification task. 
+Secondly, Next Sentence Prediction (NSP) pretraining task is a binary classification task. 
 In order to understand relationship between two sentences, BERT training process also uses next sentence prediction.
@@
 
-In terms of our approach to topic segmentation, authors utilise Next Sentence Prediction (NSP) BERT as the encoder, due to the similarity of these two tasks. 
-Both tasks take a pair of sentences/utterances as input, with the objective of predicting the appropriate next sentence, which should be topically related.
+In terms of our approach to topic segmentation, authors utilise Next Sentence Prediction (NSP) BERT as the encoder, due to the similar type of task. 
+Input of both tasks an utterances, with the objective of predicting the appropriate next sentence, which should be topically related.
 
-In more detail, the positive instances $(s_{i}, s_{t+i})$ and negative instances $(s_{i}, s_{t-i})$ are fed into the model in the form of the following sequence: 
+In more detail, the positive instances $(s_{i}, s_{t+i})$ and negative instances $(s_{i}, s_{t-i})$ use in the model's fine-tuning stage  in the form described below: 
 \begin{center}
@@
 
-Here, the contextualised representation of $[CLS]$ is employed as the topic-aware embedding to predict the degree of matching between the two input utterances in terms of topic. 
-The topic coherence scores are estimated by passing the $[CLS]$ representation through another MLP.
+In order to indicate the thematic similarity of two statements in a dialogue, the $[CLS]$ token is employed, which performs a similar role in the usual BERT. 
+The resulting value is subjected to a linear dictionary and a final similarity value is obtained, which is then utilised to calculate Coherence scores.
 
@@
 
-The input of stage is a set of utterances $\{u1, u2, ..., uk\}$. 
-Subsequently, an utterance-pair coherence scoring model is applied to $k - 1$ consecutive pairs, resulting in a sequence of coherence scores $\{c1, c2, ..., ck-1\}$. 
-The value of $c_i \in [0, 1]$ indicates the topical relatedness of the two utterances in the $i_th$ pair. 
-The algorithm utilises coherence scores as a foundation idea, yet enhances them through the sequence of "depth scores". $\{dp1, dp2, ..., dpk-1\}$ is calculated to measure the sharpness of a valley. 
-This is achieved by examining the highest coherence scores $h_l(i)$ and $h_r(i)$ on either side of interval $i$.
+With the help of a fine-tuned BERT (the training algorithm is described above) coherence scores $c_i$, where $i \in [0, n - 1]$ are calculate.
+According to texttiling algorithm description $c_i$ indicates the topical relatedness of the two adjacent utterances.
+The model uses the obtained values as base values, but processes them for final predictions and gets Depth scores $ds_i$, where $i \in [0, n - 1]$  - measuring the sharpness of a valley.
 
-The depth score for interval $i$, $dp_i$, is calculated as follows: 
+The depth score is calculated as shows below: 
 \begin{center}
-    $D_{pi} = \frac{h_{l}(i) + h_{r}(i) - 2 \cdot c_{i}}{2}$. 
+    $D_{pi} = \frac{h_{l}(i) + h_{r}(i) - 2 \cdot c_{i}}{2}$,
 \end{center}
+where $h_{l}, h_{r}$ - the highest cs for $i$ pair in choosen range.
 
-A higher depth score indicates a lower topical relatedness between the pair of utterances. 
-The threshold $\tau$, which is used to identify segment boundaries, is computed from the mean $\mu$ and standard deviation $\sigma$ of depth scores: $\tau = \mu - \frac{\sigma}{2}$. 
-A pair of utterances with a depth score exceeding $\tau$ will be selected to have a segment boundary between them.
+The greater the depth score between two statements, the less likely it is that they share a thematic relationship. 
+To identify the pairs along which the boundaries of the thematic segments lie, the threshold value is calculated as follows: $threshold = \mu - \frac{\sigma}{2}$, where $\mu$ is the average value and $\sigma$ is the standard deviation.
+Pairs with a depth score above the $threshold$ value will be choosen as boundaries.
+
+\subsection{Stage Evaluation}
+
+After defining the basic segmentation algorithm, it is worth talking about evaluation metrics for this stage:
+\begin{itemize}
+    \item \textbf{F1}\\
+    Given that this is a binary classification problem, it is tempting to use Precision and Recall, and their combination $F_{\beta}$ score as an evaluation metric. \\ 
+     - Metric is defined as the boundaries ratio identified as true boundaries; \\
+       $Precision = \frac{true \: positive}{true \: positive + false \: positive}$ \\
+     - Recall is defined as the true boundaries ratio identified by the model. \\
+       $Recall = \frac{true \: positive}{true \: positive + false \: negative}$ \\
+    $F_{\beta}$ score is defined as the weighted harmonic mean of precision and recall: \\
+    $F_{\beta} = (1 + \beta^{2}) \cdot \frac{precision \cdot recall}{(\beta^{2} \cdot precision) + recall}$
+    \item \textbf{Pk score} is calculated using a sliding window-based method.
+The window size is typically set to half of the average true segment number.
+As the window is slide, the algorithm determines whether the two ends of the window are in the same or different segments in the ground truth segmentation.
+If there is a mismatch, the algorithm increases a counter.
+The final score is calculated by scaling the penalty between 0 and 1 and dividing the number of measurements.
+    \item \textbf{WindowDiff} is also calculated by a sliding window.
+In this case, for each position of the window of size k, we simply compare the number of boundaries in the ground truth with the number of boundaries predicted by the Topic Segmentation model.\\
+    $WD(pred, true) = \frac{1}{N - k}\sum_{i = 1}^{N - k}(\left\lvert b (true_{i}, true_{i + k}) - b(pred_{i}, pred_{i + k}) \right\rvert > 0)$, \\ 
+    where $b(i, j)$ represents the number of boundaries between two positions $i$ and $j$ in the text, $N$ represents the number of sentences in the text.
+\end{itemize}
+
+In practice, both $Pk$ and $WindowDiff$ scores are used to evaluate a model. 
+A lower score means predictions are closer to the actual boundaries.
 
 \section{Topic Extraction Stage}
+
+\subsection{PromptTopic}
+
+\subsection{Stage Evaluation}
+As for the quality evaluation of the extracted topics, we will use two metrics: Topic coherence and Topic diversity
+\begin{itemize}
+    \item \textbf{Topic diversity} is a metric that assesses the ratio of unique keywords across all topic representations.
+This metric is used to measure the capabilities of a topic model covering a wide range of topics.
+The diversity score is a value between 0 and 1, with a score of 0 indicating repetitive topics and a score of 1 indicating diverse topics.
+    \item \textbf{Topic coherence} is a measure of the degree of interconnection of words within a topic.
+In our experiments, we employed normalized point-to-point mutual information (PMI) as a metric for thematic consistency.
+A higher score for a given name indicates greater consistency, with an absolute correlation estimated at 1 point.
+    \end{itemize}
 
@@
 
-Основной задачей в работе будет классификация видео на наличие юмора с использованием разных моделей для разных модальностей.
-Для работы с разными модальностями будут использованы 3 модели, которые были описаны в теоретической части:
-\begin{itemize}
-    \item Wave2Vec - извлечение аудио признаков,
-    \item BERT - извлечение текстовых признаков,
-    \item TimeSFormer - извлечение признаков из видео фрагментов,
-\end{itemize}
+The experiments will be divided into stages, with each stage accompanied by a description of the stage and the algorithms used. 
+This description is included in the theoretical part of the work.
 
-Сначала модели попробуют предсказать классы в своих модальностях, после чего их предсказания будут обработаны вместе для улучшения результатов.
+\subsection{ASR experiments}
 
-\section{Классификация по одной модальности}
+\subsection{Text clustering}
 
-Для извлечения признаков, как уже было описано, будут использоваться 3 различные модели.
-В силу ограниченных вычислительных мощностей, будут взяты предобученные модели с \href{https://huggingface.co/}{Hugging Face} и дообучены на наших данных.
+\subsection{Topic extraction}
 
-Так как у нас ровное кол-во наблюдений положительного и отрицательного классов, делим выборку 70\% на дообучение и 30\% валидацию с одинаковым кол-вом примеров разных классов в обеих подвыборках.
+\chapter{Conclusion}
 
-Для каждой модели начала посчитаем качество модели по 3 метрикам: Accuracy (из-за отличной балансировки, можем использовать обычную accuracy), Precision, Recall.
-
-\begin{itemize}
-    \item Для работы с текстом будем использовать RoBERTa из библиотеки моделей HuggingFace: \href{https://huggingface.co/roberta-base}{roberta-base}.
-    \item Для работы с аудио будем использовать Wave2vec2 из библиотеки моделей HuggingFace: \href{https://huggingface.co/facebook/wav2vec2-base}{wav2vec2-base}.
-    \item Для работы с видео будем использовать Wave2vec2 из библиотеки моделей HuggingFace: \href{https://huggingface.co/facebook/timesformer-base-finetuned-k400}{timesformer-base}.
-\end{itemize}
-
-С помощью предлагаемых Hugging Face интрументов, сделаем дообучение моделей под нашу задачу. \\
-
-\subsection{Результаты}
-
-\textbf{Далее в таблицах с результатами модальности будут называться в сокращенном виде: Текст - Т, Аудио - А, Видео - В.} \\
-
-Результаты классификации по одной модальности: \\
-
-\begin{center}
-    \begin{tabular}{ |c||c|c|c|}
-        \hline
-        \multicolumn{4}{|c|}{Классификация по тексту} \\
-        \hline
-        Модальность & Accuracy & Precision & Recall   \\
-        \hline
-        Т           & 0.71377  & 0.70563   & 0.75064  \\
-        В           & 0.47836  & 0.50600   & 0.59729  \\
-        А           & 0.50885  & 0.57865   & 0.78420  \\
-        \hline
-    \end{tabular} \\
-\end{center}
-
-TimeSFormer показывает достаточно низкие значени метрик, относительно других моделей.
-Такие результаты связаны со спецификой задачи, моедль не фоксируется только на человеке, но на всем видеофрагменте.
-Возможным решением станет использование других моделей или инструментов, которые будут обрабатывать именно человека на видео.
-
-Wave2vec показывает хороший recall (т.е модель хорошо угадывает, класс с юмором), но accuracy все еще плохой.
-
-По результатам можем сделать вывод, что самым подходящей для детекции юмора оказалась текстовая модальност, с которой работала RoBERTa.
-Действительно по тексту легче всего понять, есть ли шутка, потому что не надо обрабатывать много второстепенной информации, как в видео, и "поведение"\ голоса разных людей во время шутки, как на аудио.
-
-\section{Классификация видео с использованием нескольких модальностей}
-
-Теперь попробуем собирать вместе голосование моделей в разных модальностях, сначала будем брать все комбинации 2 модальностей, а потом возьмем 3 модальности.
-
-Подсчет результатов будем производить двумя разными сопобами:
-\begin{itemize}
-    \item Посчитать среднюю вероятность
-    \item Выбирать голосованием, каждой модели за класс
-\end{itemize} \\
-
-\subsection{Результаты}
-
-Результаты экспериментов с подсчетом средней вероятности: \\
-
-\begin{center}
-    \begin{tabular}{ |c||c|c|c| }
-        \hline
-        Модальности & Accuracy & Precision & Recall  \\
-        \hline
-        А + В       & 0.49737  & 0.50512   & 0.60373 \\
-        А + Т       & 0.71377  & 0.70488   & 0.75257 \\
-        В + Т       & 0.63573  & 0.63569   & 0.66559 \\
-        А + В + Т   & 0.63081  & 0.62862   & 0.67074 \\
-        \hline
-    \end{tabular}
-\end{center} \\
-
-В голосованиях в парах и тройке результаты лучше, чем у А и В по одиночке.
-А + Т слегка улучшает результат RoBERTa в одиночку, а в остальных парах и тройке, результаты улучшаются, но не достигают RoBERTa.
-
-При спорных ситуациях в голосованием по классам, решающий голос за модальностями с лучшим результатом в прошлой частью (т.е.
-TimeSFormer имеет самый слабый голос, затем Wave2Vec и самый сильный голос у RoBERTa).
-
-Результаты экспериментов с голосованием по классам: \\
-
-\begin{center}
-    \begin{tabular}{ |c||c|c|c|}
-        \hline
-        Модальности & Accuracy & Precision & Recall  \\
-        \hline
-        А + В       & 0.49737  & 0.50512   & 0.60373 \\
-        А + Т       & 0.71377  & 0.70563   & 0.75064 \\
-        В + Т       & 0.71377  & 0.70563   & 0.75064 \\
-        А + В + Т   & 0.50885  & 0.50885   & 1.0     \\
-        \hline
-    \end{tabular}
-\end{center} \\
-
-Здесь можно заметить, т.к. текст имеет решщающий голос, то в парах с текстом результаты одинаковые.
-В А + В + Т интересно заметить, что recall равен 1.0. Тоесть наша "тройка"\ относит все сэйплы к одному классу.
-
-\chapter{Заключение}
-
-В ходе проделанной работы были прочитаны статьи в разделе нейронных сетей, изучен механизм внимания и архитектура Трансформер с теоретической точки зрения.
-Также изучены модели, в основе которых лежит Трансформер: RoBerta, используемая для работы с текстовой модальностью, Wave2vec2, используемая для работы с звуковой модальностью, и TimeSFromer, используемая для работы с визуальной модальностью.
-
-Далее были проведены эксперименты  на датасете UrFunnny с каждой из моделей.
-Изучив результаты можно увидеть, что RoBERTa показыает самые хорошие результаты.
-Предположения, почему получились такие результаты было сделано в пунке 3.1.1 под таблицей с результатами.
-
-Что касается совмещения результатов моделей (голосование по вероятностям и финальным лэйблам), аудио и текст вместе улучшают результат RoBERTa в одиночку.
-
-Для работы с моделями, обучение и инференс, был изучен фрэймворк HuggingFace, имеющий большую популярность в сфере глубокого обучения.
-Кроме того были использованы язык программирования Python, PyTorch - бекэнд для нейронных сетей, а также вспомогательные библиотеки librosa (работа со звуком) и Numpy.
-
-Следующим шагом в исследовании детекции юмора может стать объединением моделей в одну, на примере CLIP от openAI или улучшение моделей в аудио и видео модальностях.
-
-\putbibliography %Вместо этой команды будет вставлена библиография
+\putbibliography
 

### 2024-04-22 paper/main.tex
fix ttile of paper + limitations

@@
 \author{Miakov Timofey Ilich}
-\supervisor{Professor NRU HSE - Nizhniy Novgorodw}{L.V. Savchenko }
+\supervisor{Professor NRU HSE - Nizhniy Novgorod}{L.V. Savchenko}
 \Year{2024}
-\Abstract{
-
-}
+\Abstract{}
 
@@
 \section{Research Objectives}
+
 The objectives of the thesis work are:
@@
 
+\subsection{LayerNorm}
+
+%do it
+
 \subsection{Positional encodings and Rotary embedding}
@@
 \section{BERT}
+\subsection{Model overview}
 
@@
 
-\section{Large Language models}
+\section{Large Language Model Meta AI}
 
-\subsection{LLaMA}
-
-\subsection{Mistral}
+\section{Mistral LLM}
 
@@
 
+\subsection{Utterance-Pair Coherence Scoring approach}
 The dialogue segmentation stage represents the initial stage for textual type of data. 
@@
 
-\subsection{Depth Scores calculation}
+\subsection{Depth Scores}
 
@@
 
+\subsection{LLM for topic extraction}
+
 \subsection{Stage Evaluation}
@@
 
-\subsection{LLM for topic extraction}
-
 \section{Topic Evolution over time}
@@
 
+\subsection{Experiments settings}
+
 \subsection{ASR experiments}
@@
 
+The input data at this stage are dialogues divided into separate utterances of the participants ${ut_1, ut_2, \dots, ut_n}$.
+
 \subsection{Topic extraction}
+
+\subsection{Topic Evolution}
+
+\chapter{Limitations}
+
+The LLM and other models require a significant amount of computing resources, including GPUs.
+Due to the limited availability of resources, the models were not larger than 8 billion parameters and did not exceed 16 GB during training and inference at all stages.
+Additionally, due to the lack of financial opportunities, open-source models were employed in the work.
+Should it be possible to reproduce all stages on more advanced models (such as GPT 4 or Lama 80b), the experimental results would be considerably more favourable.
 

### 2024-04-23 paper/main.tex
llm section added + new imgs

@@
 
-The objectives of the thesis work are:
+The main objectives of the research work:
 \begin{enumerate}
-    \item Research of existing solutions for dynamic text and dialogue topic modelling
+    \item Investigate existing solutions for dynamic text and dialogue topic modelling
     \item The definition of the main stages for DTM for audio dialogues
@@
 
-In terms of LayerNorm, it is normalizes the vector representation of each sequence in the batch. 
-The vector representation of each token is normalized in the Transformer. 
+\subsection{LayerNorm}
+
+In terms of LayerNorm (Figure 2.3), it is normalizes the vector representation of each sequence in the batch. 
 In addition, LayerNorm has trainable parameters, $scale$ and $bias$, which are used after normalization to scale the output of the layer.
@@
 
-\subsection{LayerNorm}
-
-%do it
+The application of layer normalisation ensures that all parameters within a given layer exhibit a uniform distribution across all features for a specific input.
+The LayerNorm can be expressed as:
+\begin{center}
+    $y = \frac{x - \mu}{\sqrt{\sigma^{2} + \epsilon}} \odot \gamma + \beta$,
+\end{center}
+where $\mu$ - mean of last D dimension, $\sigma^2$ - variance of last D dimension, $\gamma$ and $\beta$ - learnable params;
 
@@
 
-A newer version of positional encoding is rotary embeddings.
+A newer version of positional encoding is Rotary Positional Embeddings.
 Rotary position embedding represents a distinct approach to incorporating relative position information into the attention matrix. 
@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.7]{img/rotary_embs.png}
-    \caption{Rotary embeddings}
-\end{figure}
+In order to illustrate the concept, we will consider an examples with two features at position $m$.
+\begin{center}
+    $RoPE(x_1, x_2) =   \begin{pmatrix}
+                            \cos m\theta & -\sin m\theta\\
+                            \sin\theta & \cos m\theta
+                        \end{pmatrix} 
+                        \begin{pmatrix}
+                            x_1 \\
+                            x_2
+                        \end{pmatrix} 
+                        \begin{pmatrix}
+                            x_1 \cos m\theta - x_2\sin m\theta\\
+                            x_2 \cos m\theta + x_1\sin m\theta
+                        \end{pmatrix},
+    $
+\end{center}
+where $\theta$ is a constant angle.
 
@@
 
-\section{Large Language Model Meta AI}
+\section{LLM}
 
-\section{Mistral LLM}
+LLM is a type of artificial intelligence(AI) algorithm that use deep learning techniques and massively large data sets to understand, summarize, generate, and predict new content. 
+LLMs are trained with immense amounts of data and use self-supervised learning to predict the next token in a sentence, given the surrounding context. 
+Once an LLM has been trained, it can be fine-tuned for a wide range of NLP tasks, including generating and classifying text, answering questions.
+The following section will examine the main LLMs used in the article.
+
+\subsection{LLaMA}
+
+In July 2023, an open-source collection of models, called LLaMA 2, was released by Mata AI Team. The collection consists of models with 7B and 70B parameters.
+
+In order to create a collection of LLaMA 2 models, the authors employed the architecture of an optimised auto-regressive transformer (Touvron et al.(2023)) with RMSNorm normalisation, Rotary Embeddings and SwiGLU activation function in FFN, but implemented a number of alterations with the intention of enhancing the quality of the model and its performance, such as:
+\begin{itemize}
+    \item Meticulous data preparation
+    \item Larger dataset
+    \item Increased the length of the context in compare with earlier approaches
+    \item Incorporation of grouped-query attention (GQA)
+\end{itemize}
+
+\textbf{Grouped-query attention} can be thought of as a way to optimize the attention mechanism in transformer-based models. 
+Instead of computing attention for each query independently, GQA groups queries together and computes their attention jointly.
+This reduces the number of attention computations, leading to faster inference times.
+
+The GQA method has been applied to speed up inference on large language models without significantly sacrificing quality. 
+It's a promising technique for improving the efficiency of transformer models, particularly in the context of generative AI.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=1]{img/gqa.png}
+    \caption{Self-attention and Grouped-query attention}
+\end{figure}
+
+\section{Mistral}
+
+In September 2023, the MistralAI team introduced the Mistral 7B model, which has been designed for high performance and efficiency. 
+In comparison to the 13B Llama 2, the Mistral 7B model has demonstrated superior performance in all benchmarks related to reasoning, mathematics, and code generation. 
+The Mistral 7B model employs grouped-query attention (GQA) for rapid inference and sliding window attention (SWA) for handling long sequences at a reduced cost.
+
+As other models, Mistral is a transformer based model. However, it differs from the Llama model in several key respects.
+\begin{itemize}
+    \item Sliding Window Attention (SWA): This method allows the model to attend to information beyond a set window size by leveraging the stacked layers of a transformer. 
+    With a window size of 4096, it can attend to around 131K tokens. 
+    \item Rolling Buffer Cache: It is limits the size of the cache using a rolling buffer.
+The cache has a fixed size, and as new data is added, older data is overwritten once the cache exceeds its size.
+    This method reduces the memory usage of the cache by 8x for sequences of 32k tokens without compromising the quality of the model.
+    \item In the context of sequence generation, the model employs a process of pre-filling and chunking. 
+    This involves the prediction of tokens in a sequential manner, given that each token is contingent upon the previous one.
+However, given that the prompts are known in advance, the cache can be pre-filled with them.
+    In the event that a prompt is of a considerable length, it can be divided into smaller units, with the cache being filled in this manner. 
+    The attention mask operates over both the cache and the chunk
+\end{itemize}
 

### 2024-04-23 paper/main.tex
add some new text

@@
 
-\section{Mistral}
+\subsection{Mistral}
 
@@
     $F_{\beta} = (1 + \beta^{2}) \cdot \frac{precision \cdot recall}{(\beta^{2} \cdot precision) + recall}$
-    \item \textbf{Pk score} is calculated using a sliding window-based method.
-The window size is typically set to half of the average true segment number.
-As the window is slide, the algorithm determines whether the two ends of the window are in the same or different segments in the ground truth segmentation.
-If there is a mismatch, the algorithm increases a counter.
-The final score is calculated by scaling the penalty between 0 and 1 and dividing the number of measurements.
-    \item \textbf{WindowDiff} is also calculated by a sliding window.
-In this case, for each position of the window of size k, we simply compare the number of boundaries in the ground truth with the number of boundaries predicted by the Topic Segmentation model.\\
-    $WD(pred, true) = \frac{1}{N - k}\sum_{i = 1}^{N - k}(\left\lvert b (true_{i}, true_{i + k}) - b(pred_{i}, pred_{i + k}) \right\rvert > 0)$, \\ 
-    where $b(i, j)$ represents the number of boundaries between two positions $i$ and $j$ in the text, $N$ represents the number of sentences in the text.
+    \item \textbf{Pk score} is uses the sliding window method and counts the number of discrepancies between the ends of the predicted segments and the true segments.
+The final score is the counter divided by the number of measurements.
+The lower the value, the better.
+    \item In \textbf{WindowDiff} this algorithm, for each position of a sliding window of size k, comparing the number of boundaries in the ground truth with the number of boundaries predicted by the Topic Segmentation model.\\
+    $WD(pred, true) = \frac{1}{N - k}\sum_{i = 1}^{N - k}(\left\lvert bounds(true_{i}, true_{i + k}) - bounds(pred_{i}, pred_{i + k}) \right\rvert > 0)$, \\ 
+    where $bounds(s, s + i)$ represents the number of boundaries between two positions $i$ and $j$ in the text, $N$ represents the number of sentences in the text.
 \end{itemize}
@@
 
-\subsection{Experiments settings}
+\section{Experiments settings}
 
-\subsection{ASR experiments}
+\section{ASR experiments}
 
-\subsection{Text clustering}
+\section{Text clustering}
 
 The input data at this stage are dialogues divided into separate utterances of the participants ${ut_1, ut_2, \dots, ut_n}$.
+At this stage, experiments will be conducted using the encoder model for the Utterance-Coherence Scoring algorithm. 
+In addition, algorithms for the time of model inference will be performed throughout the pipeline to select the fastest solution.
 
-\subsection{Topic extraction}
+\begin{center}
+    \begin{tabular}{ | c | c | c | c | c | c | c | c |} 
+        \hline
+        \multicolumn{2}{| c |}{Country List} & \multicolumn{3}{| c |}{Country List} &  \multicolumn{3}{| c |}{Country List}  \\
+        \hline
+        model & dataset & Pk & wd & f1 & Pk & wd & f1 \\ 
+        \hline
+        bert-base NSP & committee & Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-base CLS & committee & Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-large NSP & isci & Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-base CLS & isci & Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-base NSP & ami & Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-base CLS & ami & Pk & wd & f1 & Pk & wd & f1 \\ 
+        \hline
+    \end{tabular}
+\end{center}
 
-\subsection{Topic Evolution}
+\section{Topic extraction}
 
-\chapter{Limitations}
+\begin{center}
+    \begin{tabular}{ |c|c|c| } 
+     \hline
+     model & Pk & wd \\ 
+     \hline
+     bert-base NSP & cell5 & cell6 \\ 
+     bert-base CLS & cell8 & cell9 \\ 
+     \hline
+    \end{tabular}
+\end{center}
+
+\section{Topic Evolution}
+
+\section{Limitations}
 
@@
 Additionally, due to the lack of financial opportunities, open-source models were employed in the work.
-Should it be possible to reproduce all stages on more advanced models (such as GPT 4 or Lama 80b), the experimental results would be considerably more favourable.
+Should it be possible to reproduce all stages on more advanced models (such as GPT4 or LlaMA2 80b), the experimental results would be considerably more favourable.
 
@@
 \putbibliography
-
 \end{document}

### 2024-04-24 paper/main.tex
update bib

@@
 
+Natural Language Processing (NLP) has emerged as a crucial field in computer science, enabling machines to comprehend and respond to human language in a meaningful way. 
+Among the various tasks within NLP, dialogue understanding, which involves analyzing and interpreting natural language conversations, has gained significant attention. 
+One of the key challenges in dialogue understanding is identifying and tracking different topics discussed throughout the conversation.
+
+Topic modeling is a statistical technique that helps uncover topic structures within textual data. 
+Traditional topic modeling approaches, however, suffer from limitations such as topic drift, where the topics change over time, and the inability to capture dynamic relationships between topics. 
+In order to analyse time-dependent text corpora, the Dynamic Topic Modeling (DTM) approach is utilised, which will be subjected to further investigation in the forthcoming work.
+
+In this paper, we explore current approaches to (DTM) and apply them to text and audio dialogue datasets. 
+Furthermore, we will endeavour to combine several of the most effective open-source methodologies into pipeline from text clustering to topic evolution over time. 
+A significant aspect of our research will be to investigate the potential for integrating Large Language Models into our DTM pipeline, with a view to comparing it with other existing solutions.
+
+The quality and performance of the pipeline will be evaluated using specific metrics at both the overall and stage-specific levels.
+Following this, a comparison will be made with other existing solutions and models.
+
 \section{Relevance of the topic}
+
+Dynamic Topic Modeling is a powerful tool for voice dialogue understanding, offering a dynamic and comprehensive representation of the conversation structure. 
+It has the potential to significantly enhance the performance of voice dialogue systems and improve the overall user experience.
+In addition, exploring the possibilities of using large language models, which are now showing significant results in many NLP tasks, can further improve the model.
 
@@
 
-In the decodert part the structure of the attention mechanism is different. 
-They use Masked self-attention. In the form of attention described above, each token will "look" at the entire sequence, which is undesirable for the decoder.
-
-In the decoder, the generation of tokens occurs sequentially.
-If the attention mechanism has access to all tokens, not just those already used, this will result in information leakage in the decoder at the training stage and a reduction in the quality of the model.
-To avoid this problem, Masked Self-Attention is used, which masks tokens that follow the current one and reduces their probability in the softmax function to 0.
+The decoder uses a modified Self-attention mechanism called Masked Self-attention.
+Because a new sequence is generated in the decoder, the tokens that come after the current one are masked so that the model does not see future tokens and does not rely on information about them.
+Masking leads to the fact that after softmax in the attention mechanism, the probability of future tokens is zero.
 
@@
 
-To cover more information and dependencies, the authors propose the use of multiple levels of attention in parallel and the summarisation of the values obtained. 
-This is referred to as Multi-head attention. The figure below illustrates what a multi-head approach looks like:
+To cover more information and dependencies, the authors propose Multi-head attention, that use of multiple levels of attention in parallel and the summarisation of the values obtained. 
+The figure below illustrates what a multi-head approach looks like:
 
@@
 
-A sequence of tokens is fed into the model, but the model itself does not understand in which order the tokens are located. 
+Model itself does not understand in which order the input tokens tokens are located. 
 Therefore, positional embeddings are added to the tokens, which are designed to show the model the input order of the elements.
@@
 
-A newer version of positional encoding is Rotary Positional Embeddings.
+A newer version of positional encoding is Rotary Positional Embeddings \cite{rotary:2021}.
 Rotary position embedding represents a distinct approach to incorporating relative position information into the attention matrix. 
@@
 
-\section{BERT}
-\subsection{Model overview}
+\section{Transformer based models}
+
+Several models based on the Transformer architecture will be required in the work at different stages. 
+More specifically, text clustering will require an encoder for text information based on BERT, and LLM will be used for topic processing.
+
+\subsection{BERT}
 
@@
 
-\subsection{Training tasks}
-BERT has two pretrain objective described in article: \\
+\textbf{BERT has two pretrain objective} described in article: \\
 The first objective of the training process is the masked language modelling (MLM). 
@@
 During training the model gets as input pairs of sentences and it learns to predict if the second sentence is the next sentence in the original text as well. 
-The training dataset presented in the original paper states that when trained, 50\% of the examples contain related sentences extracted from the training texts, while the remaining 50\% contain a random pair of sentences.
+The training dataset presented in the original paper states that when trained, 50\% of the examples contain related sentences extracted from the training texts, while the remaining 50\% contain a random pair of sentences.\\
 
-\section{LLM}
-
-LLM is a type of artificial intelligence(AI) algorithm that use deep learning techniques and massively large data sets to understand, summarize, generate, and predict new content. 
+\newpage
+\textbf{LLM} is a type of artificial intelligence(AI) algorithm that use deep learning techniques and massively large data sets to understand, summarize, generate, and predict new content. 
 LLMs are trained with immense amounts of data and use self-supervised learning to predict the next token in a sentence, given the surrounding context. 
@@
 
-In July 2023, an open-source collection of models, called LLaMA 2, was released by Mata AI Team. The collection consists of models with 7B and 70B parameters.
+In July 2023, an open-source collection of models, called LLaMA2 \cite{llama:2023}, was released by Mata AI Team. The collection consists of models with 7B and 70B parameters.
 
@@
 
-In September 2023, the MistralAI team introduced the Mistral 7B model, which has been designed for high performance and efficiency. 
+In September 2023, the MistralAI \cite{mistral:2023} team introduced the Mistral 7B model, which has been designed for high performance and efficiency. 
 In comparison to the 13B Llama 2, the Mistral 7B model has demonstrated superior performance in all benchmarks related to reasoning, mathematics, and code generation. 
@@
 
-\section{Automatic Speech Recognition Stage}
+\section{ASR approach}
 
-\section{Dialogue Segmentation Stage}
+\section{Improving Utterance-Pair Coherence Scoring}
 
-\subsection{Utterance-Pair Coherence Scoring approach}
 The dialogue segmentation stage represents the initial stage for textual type of data. 
@@
 Let us now examine the algorithm in greater detail.
+
+\subsection{BERT for coherence scoring}
 
@@
 
-\subsection{Stage Evaluation}
+\subsection{Evaluation Metrics}
 
@@
 
-\section{Topic Extraction Stage}
-
-\subsection{PromptTopic}
+\section{PromptTopic}
 
@@
     \end{itemize}
-
-\section{Topic Evolution over time}
 
@@
 Due to the limited availability of resources, the models were not larger than 8 billion parameters and did not exceed 16 GB during training and inference at all stages.
-Additionally, due to the lack of financial opportunities, open-source models were employed in the work.
+Additionally, because of the absence of financial opportunities, open-source models were employed in the work.
 Should it be possible to reproduce all stages on more advanced models (such as GPT4 or LlaMA2 80b), the experimental results would be considerably more favourable.

### 2024-04-25 paper/main.tex
update bib and some fixes in paper

@@
     \includegraphics[scale=0.7]{img/multi-head.png}
-    \caption{Query, Key, Value in Self-Attention mechanism}
+    \caption{Multi-head Attention mechanism}
 \end{figure}
@@
 In accordance with the majority of previous studies, we have utilise the best open-source solution based on the presented in papers metrics. 
-It is algorithm from Improving Unsupervised Dialogue Topic Segmentation with Utterance-Pair Coherence Scoring by Linzi Xing and Giuseppe Carenini (2020)
+It is algorithm from Improving Unsupervised Dialogue Topic Segmentation with Utterance-Pair Coherence Scoring \cite{depthscore:2021} 
 Let us now examine the algorithm in greater detail.
@@
     $F_{\beta} = (1 + \beta^{2}) \cdot \frac{precision \cdot recall}{(\beta^{2} \cdot precision) + recall}$
-    \item \textbf{Pk score} is uses the sliding window method and counts the number of discrepancies between the ends of the predicted segments and the true segments.
+    \item \textbf{Pk score} \cite{pk:1999} is uses the sliding window method and counts the number of discrepancies between the ends of the predicted segments and the true segments.
 The final score is the counter divided by the number of measurements.
 The lower the value, the better.
-    \item In \textbf{WindowDiff} this algorithm, for each position of a sliding window of size k, comparing the number of boundaries in the ground truth with the number of boundaries predicted by the Topic Segmentation model.\\
+    \item In \textbf{WindowDiff} \cite{wd:2002} this algorithm, for each position of a sliding window of size k, comparing the number of boundaries in the ground truth with the number of boundaries predicted by the Topic Segmentation model.
+The lower the value, the better.\\
     $WD(pred, true) = \frac{1}{N - k}\sum_{i = 1}^{N - k}(\left\lvert bounds(true_{i}, true_{i + k}) - bounds(pred_{i}, pred_{i + k}) \right\rvert > 0)$, \\ 
-    where $bounds(s, s + i)$ represents the number of boundaries between two positions $i$ and $j$ in the text, $N$ represents the number of sentences in the text.
+    where $bounds(s, s + i)$ is the number of boundaries between $i$ and $j$ positions, $N$ - number of sentences in the text.
 \end{itemize}
-
-In practice, both $Pk$ and $WindowDiff$ scores are used to evaluate a model. 
-A lower score means predictions are closer to the actual boundaries.
 
@@
 \begin{itemize}
-    \item \textbf{Topic diversity} is a metric that assesses the ratio of unique keywords across all topic representations.
+    \item \textbf{Topic diversity} \cite{topicdiversity:2019} is a metric that assesses the ratio of unique keywords across all topic representations.
 This metric is used to measure the capabilities of a topic model covering a wide range of topics.
 The diversity score is a value between 0 and 1, with a score of 0 indicating repetitive topics and a score of 1 indicating diverse topics.
-    \item \textbf{Topic coherence} is a measure of the degree of interconnection of words within a topic.
+    \item \textbf{Topic coherence} \cite{topiccoh:2009} is a measure of the degree of interconnection of words within a topic.
 In our experiments, we employed normalized point-to-point mutual information (PMI) as a metric for thematic consistency.
@@
     \end{itemize}
-
 \newpage

### 2024-04-27 paper/main.tex
add whisper

@@
 
+Since one of the significant tasks is to use DTM for audio data, the first, but optional, stage of our pipeline will be a model for automatic speech recognition. 
+In this section, we will look at the whisper model from OpenAI and the MODEL 2.
+
+\subsection{Whisper}
+
+%need perephrace
+
+Whisper is an automatic speech recognition (ASR) system trained on 680,000 hours of multilingual and multitask supervised data collected from the web. 
+We show that the use of such a large and diverse dataset leads to improved robustness to accents, background noise and technical language.
+Moreover, it enables transcription in multiple languages, as well as translation from those languages into English.
+We are open-sourcing models and inference code to serve as a foundation for building useful applications and for further research on robust speech processing.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.7  ]{img/whisper.png}
+    \caption{Whisper architecture}
+\end{figure}
+
+The Whisper architecture is a simple end-to-end approach, implemented as an encoder-decoder Transformer. 
+Input audio is split into 30-second chunks, converted into a log-Mel spectrogram, and then passed into an encoder. 
+A decoder is trained to predict the corresponding text caption, intermixed with special tokens that direct the single model to perform tasks such as language identification, phrase-level timestamps, multilingual speech transcription, and to-English speech translation.
+
 \section{Improving Utterance-Pair Coherence Scoring}
@@
 
+\textbf{Datasets}
+
+For the experiments we will use two types of datasets: text dialogues and audio dialogues. 
+For text dialogues, we will use tiage, compiled from the personal chat corpus and manually annotated, and superdialseg, compiled from several smaller datasets: doc2dial and multidoc2dial. 
+As for the audio dialogues, we will use QMSum, which includes the audio recording corpuses - AMI and ISCI and is divided into three types of meetings (product, academic, committee). \\
+
+\textbf{Baseline Models}
+
+At the text clustering stage, comparisons are made between models of non-alternative baselines, because at this stage we are using a ready-made algorithm and making a minimum of changes.
+As for topic extraction, our LLM approach will be compared with BertTopic and LSA.
+
+\textbf{Pipeline configuration}
+
+MODEL1 will be used in the ASR for audio datasets. 
+At the dialogue clustering stage, BERT-based fine-tuning models will be used. 
+For topic extraction, the LlaMA 3 8B and Mistral 7B models will be used, in addition to experiments with a smaller model - phi3.
+
 \section{ASR experiments}
@@
 In addition, algorithms for the time of model inference will be performed throughout the pipeline to select the fastest solution.
+
+\begin{center}
+    \begin{tabular}{ | c | c | c | c | c | c | c | c |} 
+        \hline
+        \multicolumn{2}{| c |}{Country List} & \multicolumn{3}{| c |}{Country List} &  \multicolumn{3}{| c |}{Country List}  \\
+        \hline
+        model & dataset & Pk & wd & f1 & Pk & wd & f1 \\ 
+        \hline
+        bert-base NSP & tiage & Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-base CLS & superseg & Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-large NSP & tiage & Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-base CLS & superseg & Pk & wd & f1 & Pk & wd & f1 \\ 
+        \hline
+    \end{tabular}
+\end{center}
 
@@
 
+Summarizing all the results of the article:
+
+The study encompassed a range of approaches to topic modelling, including techniques such as segmentation of dialogues, topic extraction and topic evolution for dynamic topic modelling.
+For each of these stages, open source approaches with high reported metrics and open source models were studied.
+
 \putbibliography

### 2024-05-01 paper/main.tex
write topic extraction + exps

@@
 
-\section{PromptTopic}
+\section{Topic Extraction}
 
-\subsection{LLM for topic extraction}
+The next stage in the pipeline is topic extraction. At this stage, $n$ topics will be extracted from each segment in the dialogues. 
+At this stage, past approaches have used probabilistic or embedding approaches, like BertTopic \cite{berttopic:2022}. 
+In our work, we will apply a generative approach using Large Language models, as in PromptTopic \cite{prompttopic:2023}.
+
+We will try several Large Language Models by submitting a dialogue segment as input. The prompt for LLM will look like this:
+
+\begin{lstlisting}
+    <|user|> 
+    utterance_1
+    utterance_2
+    ...
+    utterance_n. 
+    You have dialogue between two people, 
+    extract N topic words. 
+    <|assistant|>
+
+\end{lstlisting}
+
+where $utterance$ is one phrase of the participant in the dialogue, \\$M$ - is the number of words. The appearance of the keywords <|user|>, <|assistant|> dependes on type of LLM.
 
 \subsection{Stage Evaluation}
+
 As for the quality evaluation of the extracted topics, we will use two metrics: Topic coherence and Topic diversity
@@
 \end{itemize}
-\newpage
+
 \chapter{Experiments}
@@
 
+\begin{table}[htb]
+    \begin{center}
+        \begin{tabular}{ | c | c | c | c |} 
+            \hline
+            dataset & train/val/test & Avg Turns & Avg tokens \\ 
+            \hline
+            SuperDialseg & 6984/1322/1322 & 13 & 239 \\ 
+            Tiage &  7939/-/1000 & 14.7  & null \\ 
+            Product Meetings (QMSum) & 137 & 535.6 & null \\ 
+            Academic Meetings (QMSum) & 59 & 819.0 & null \\ 
+            Comemitie Meetings (QMSum) & 36 & 207.7 & null \\
+            \hline
+        \end{tabular}
+    \end{center}
+\end{table}
+
 \textbf{Baseline Models}
@@
 At the text clustering stage, comparisons are made between models of non-alternative baselines, because at this stage we are using a ready-made algorithm and making a minimum of changes.
-As for topic extraction, our LLM approach will be compared with BertTopic and LSA.
+As for topic extraction, our LLM approach will be compared with BertTopic and LSA. \\
 
@@
         \hline
-        \multicolumn{2}{| c |}{Country List} & \multicolumn{3}{| c |}{Country List} &  \multicolumn{3}{| c |}{Country List}  \\
+        \multicolumn{1}{| c |}{} & \multicolumn{3}{| c |}{Tiage} &  \multicolumn{3}{| c |}{SuperDialseg}  \\
         \hline
-        model & dataset & Pk & wd & f1 & Pk & wd & f1 \\ 
+        model &  Pk & wd & f1 & Pk & wd & f1 \\ 
         \hline
-        bert-base NSP & tiage & Pk & wd & f1 & Pk & wd & f1 \\ 
-        bert-base CLS & superseg & Pk & wd & f1 & Pk & wd & f1 \\ 
-        bert-large NSP & tiage & Pk & wd & f1 & Pk & wd & f1 \\ 
-        bert-base CLS & superseg & Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-base NSP & Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-base CLS & Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-large NSP & Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-base CLS & Pk & wd & f1 & Pk & wd & f1 \\ 
         \hline
@@
 
+In terms of textual datasets ...
+
 \begin{center}
-    \begin{tabular}{ | c | c | c | c | c | c | c | c |} 
+    \begin{tabular}{ | c | c | c | c | c | c | c | c | c | c |} 
         \hline
-        \multicolumn{2}{| c |}{Country List} & \multicolumn{3}{| c |}{Country List} &  \multicolumn{3}{| c |}{Country List}  \\
+        \multicolumn{1}{| c |}{} & \multicolumn{3}{| c |}{Product} &  \multicolumn{3}{| c |}{Academic} & \multicolumn{3}{| c |}{Committee} \\
         \hline
-        model & dataset & Pk & wd & f1 & Pk & wd & f1 \\ 
+        model  & Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
         \hline
-        bert-base NSP & committee & Pk & wd & f1 & Pk & wd & f1 \\ 
-        bert-base CLS & committee & Pk & wd & f1 & Pk & wd & f1 \\ 
-        bert-large NSP & isci & Pk & wd & f1 & Pk & wd & f1 \\ 
-        bert-base CLS & isci & Pk & wd & f1 & Pk & wd & f1 \\ 
-        bert-base NSP & ami & Pk & wd & f1 & Pk & wd & f1 \\ 
-        bert-base CLS & ami & Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-base NSP &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-base CLS &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1  \\ 
+        bert-large NSP & Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-base CLS &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-base NSP &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-base CLS &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
         \hline
@@
 
+Speaking about results on audio corpora ...
+
 \section{Topic extraction}
-
 \begin{center}
-    \begin{tabular}{ |c|c|c| } 
+    \begin{tabular}{ | c | c | c | c | c | } 
         \hline
-     model & Pk & wd \\ 
+        \multicolumn{1}{| c |}{} & \multicolumn{2}{| c |}{Tiage} &  \multicolumn{2}{| c |}{SuperDialseg}  \\
         \hline
-     bert-base NSP & cell5 & cell6 \\ 
-     bert-base CLS & cell8 & cell9 \\ 
+        model & TC & TD & TC & TD \\ 
+        \hline
+        bert-base NSP & c & d & c & d \\ 
+        bert-base CLS & c & d & c & d \\ 
+        bert-large NSP & c & d & c & d \\ 
+        bert-base CLS & c & d & c & d \\ 
         \hline
@@
 \end{center}
+
+In terms of textual datasets ...
+
+\begin{center}
+    \begin{tabular}{ | c | c | c | c | c | c | c |} 
+        \hline
+        \multicolumn{1}{| c |}{} & \multicolumn{2}{| c |}{Product} &  \multicolumn{2}{| c |}{Academic} & \multicolumn{2}{| c |}{Committee} \\
+        \hline
+        model  & TC & TD & TC & TD & TC & TD  \\ 
+        \hline
+        LLaMA2 7b & c & d & c & d & c & d \\ 
+        LLaMA3 8b & c & d & c & d & c & d \\ 
+        Mistral 7B & c & d & c & d & c & d \\ 
+        phi3 & c & d & c & d & c & d \\ 
+        bert-base NSP & c & d & c & d & c & d \\ 
+        bert-base CLS & c & d & c & d & c & d \\ 
+        \hline
+    \end{tabular}
+\end{center}
+
+Speaking about results on audio corpora ...
 
@@
 Summarizing all the results of the article:
-
-The study encompassed a range of approaches to topic modelling, including techniques such as segmentation of dialogues, topic extraction and topic evolution for dynamic topic modelling.
-For each of these stages, open source approaches with high reported metrics and open source models were studied.
+\begin{itemize}
+    \item Various approaches for dialogue segmentation, topic mining, and topic evolution have been explored;
+    \item The search for suitable text and audio datasets was conducted, and the selected datasets were analysed and preprocessed for experiments at all stages and for the entire pipeline.
+    \item In the context of the dialog segmentation stage, the optimal open-source solution was identified and the BERT model was subjected to detailed analysis.
+This approach has been prepared and integrated into the general pipeline.
+    \item At the topic extraction stage, new approaches utilising generative models were investigated.
+Furthermore, a range of large language models, including LlaMA and Mistral, were examined, along with novel features that enhance the default architecture of the Transformer.
+    \item Topic Evolution
+    \item A suitable set of metrics was identified and selected to evaluate the quality of each stage.
+In order to utilise the metrics, a library with an implementation was Investigated and integrated into pipeline.
+\end{itemize}
 

### 2024-05-02 paper/main.tex
fix title and some new cites

@@
 \author{Miakov Timofey Ilich}
-\supervisor{Professor NRU HSE - Nizhniy Novgorod}{L.V. Savchenko}
-\Year{2024}
 \Abstract{}
@@
     \item Selection of appropriate metrics and measurement of the quality of the approaches used at each stage
-    \item Visual analysis of the pipeline operation on pre-selected audio data
+    \item Visual evaluation of the pipeline inference on choosen audio samples
 \end{enumerate}
@@
 Since one of the significant tasks is to use DTM for audio data, the first, but optional, stage of our pipeline will be a model for automatic speech recognition. 
-In this section, we will look at the whisper model from OpenAI and the MODEL 2.
+In this section, we will look at the Whisper \cite{whisper:2022} model from OpenAI.
 
 \subsection{Whisper}
-
-%need perephrace
 
@@
 
-\section{ASR experiments}
-
 \section{Text clustering}

### 2024-05-02 paper/main.tex
new info

@@
 
+\newpage
+In april 2024 was released LLaMA3 8B and 70B models with some new features: larger tokenizer vocabulary 128k agains 32k, the dataset for pre-training contains 8 times more sampels on over 15 trillion tokens on a new mix of publicly available online data
+
 \subsection{Mistral}
@@
         bert-large NSP & Pk & wd & f1 & Pk & wd & f1 \\ 
-        bert-base CLS & Pk & wd & f1 & Pk & wd & f1 \\ 
+        opt-128 & Pk & wd & f1 & Pk & wd & f1 \\ 
+        dialoGPT-medium & Pk & wd & f1 & Pk & wd & f1 \\ 
         \hline
@@
         bert-large NSP & Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
-        bert-base CLS &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
-        bert-base NSP &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
-        bert-base CLS &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
+        opt-128 &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
+        dialoGPT-medium &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
         \hline
@@
         \hline
-        bert-base NSP & c & d & c & d \\ 
-        bert-base CLS & c & d & c & d \\ 
-        bert-large NSP & c & d & c & d \\ 
-        bert-base CLS & c & d & c & d \\ 
+        LLaMA3 8B & c & d & c & d \\ 
+        Qwen1.5 7B & c & d & c & d \\ 
+        Mistral 7B & c & d & c & d \\ 
+        \hline
+        BertTopic & c & d & c & d \\
+        LSA  & c & d & c & d \\
         \hline
@@
         \hline
-        LLaMA2 7b & c & d & c & d & c & d \\ 
-        LLaMA3 8b & c & d & c & d & c & d \\ 
+        LLaMA3 8B & c & d & c & d & c & d \\ 
+        Qwen1.5 7B & c & d & c & d & c & d \\ 
         Mistral 7B & c & d & c & d & c & d \\ 
-        phi3 & c & d & c & d & c & d \\ 
-        bert-base NSP & c & d & c & d & c & d \\ 
-        bert-base CLS & c & d & c & d & c & d \\ 
+        \hline
+        BertTopic & c & d & c & d & c & d \\ 
+        LSA & c & d & c & d & c & d \\ 
         \hline

### 2024-05-03 paper/main.tex
add qmsum
add qwen

@@
 Once an LLM has been trained, it can be fine-tuned for a wide range of NLP tasks, including generating and classifying text, answering questions.
-The following section will examine the main LLMs used in the article.
+The following section will examine the main LLMs used in the experiments: LLaMA, Mistral, Qwen.
 
@@
 \end{itemize}
+
+\subsection{Qwen}
+
+The Qwen LLMs have been extensively trained on up to 3 trillion tokens of diverse text and code examples from a wide range of domains. 
+Qwen demonstrate superior performance in various downstream tasks.
+
+In terms of architecture: model based on LLaMA, activation function is SwiGLU, Rotary Embeddings, Pre-Norm and RMSNorm instead LayerNorm.
+
+The pre-training phase involves learning vast amounts of data to gain a comprehensive understanding of the world and its various complexities. 
+This includes not only basic language skills, but also advanced skills such as arithmetic, coding and reasoning. 
+In this section we introduce the data, model design and scaling, and the extensive evaluation results on benchmark datasets.
+
+The Qwen Series encompasses models that employ SFT and RLHF to align Qwen with human preference. 
+Consequently, the Qwen-Chat and its enhanced variant, the Qwen-Chat-RLHF, have been developed. Furthermore, we have created bespoke models for coding and mathematics.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=1]{img/qwen.png}
+    \caption{Qwen Series}
+\end{figure}
+
+For now, there are 5 models of different sizes, 4 of which are opensourced. Specially, Qwen-1.8B, Qwen-7B, Qwen-14B, and Qwen-72B.
 
@@
     utterance_n. 
-    You have dialogue between two people, 
-    extract N topic words. 
+    Extract N key words or pharases 
+    that describe the given dialog between two peoples, 
+    do not use more than 2 wrods 
     <|assistant|>
@@
 \section{Topic extraction}
+
 \begin{center}

### 2024-05-07 paper/main.tex
update methodology

@@
 
-\chapter{Introduction}
+\chapter*{Introduction}
 
@@
 
-The main objectives of the research work:
+The main objectives of the research work is to create and apply Dynamic Topic Modeling pipeline for audio and text dialogues. 
+To solve this problem, you will need to complete the following tasks 
 \begin{enumerate}
-    \item Investigate existing solutions for dynamic text and dialogue topic modelling
-    \item The definition of the main stages for DTM for audio dialogues
-    \item The identification of suitable training data sets and their collection
+    \item The identification of suitable datasets for experiments
+    \item Investigate open-source solutions for ASR, dialogue clustering, topic extraction and topic evolution.
     \item An exploration of the possibilities of using large language models for some stages and the selection of models suitable for training in limited resources
@@
     \item Selection of appropriate metrics and measurement of the quality of the approaches used at each stage
-    \item Visual evaluation of the pipeline inference on choosen audio samples
 \end{enumerate}
@@
 
-The pipeline of dynamic thematic modelling of audio dialogues, which will be considered in this paper, is divided into several main stages. 
-These include audio-to-text, dialogue segmentation, topic extraction, and topic evolution. 
-The approaches and algorithms used in these stages will be described in detail from a theoretical point of view.
-
-\section{ASR approach}
+\section{Automatic Speech Recognition}
 
@@
 A decoder is trained to predict the corresponding text caption, intermixed with special tokens that direct the single model to perform tasks such as language identification, phrase-level timestamps, multilingual speech transcription, and to-English speech translation.
+
+\subsection{Speaker-diarization}
 
@@
 
-\subsection{Evaluation Metrics}
+\section{Evaluation Metrics}
+
+\subsection{Dialogue segmentation}
 
@@
 
-\section{Topic Extraction}
-
-The next stage in the pipeline is topic extraction. At this stage, $n$ topics will be extracted from each segment in the dialogues. 
-At this stage, past approaches have used probabilistic or embedding approaches, like BertTopic \cite{berttopic:2022}. 
-In our work, we will apply a generative approach using Large Language models, as in PromptTopic \cite{prompttopic:2023}.
-
-We will try several Large Language Models by submitting a dialogue segment as input. The prompt for LLM will look like this:
-
-\begin{lstlisting}
-    <|user|> 
-    utterance_1
-    utterance_2
-    ...
-    utterance_n. 
-    Extract N key words or pharases 
-    that describe the given dialog between two peoples, 
-    do not use more than 2 wrods 
-    <|assistant|>
-
-\end{lstlisting}
-
-where $utterance$ is one phrase of the participant in the dialogue, \\$M$ - is the number of words. The appearance of the keywords <|user|>, <|assistant|> dependes on type of LLM.
-
-\subsection{Stage Evaluation}
+\subsection{Topic extraction}
 
@@
 \end{itemize}
+
+\chapter{Methodology}
+
+\section{Pipeline configuration}
+
+The pipeline of dynamic thematic modelling of audio dialogues, which will be considered in this paper, is divided into several main stages. 
+These include ASR, dialogue segmentation, topic extraction and topic evolution.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.7]{img/pipeline.png}
+    \caption{DTM pipeline}
+\end{figure}
+
+All the models in the pipeline were taken from the \href{https://huggingface.co/}{HuggingFace} hub. 
+Tools from \href{https://huggingface.co/docs/transformers/index}{transformers}, which is also part of HuggingFace, 
+will be used to create the pipeline.
+
+\section{ASR}
+
+The first optional stage is the Automatic Speech Recognition. 
+For audio dialogues, text is extracted from the audio track before segmentation. 
+For this stage, open-source:
+\begin{itemize}
+    \item OpenAI Whisper \\
+    Link to model: \href{https://huggingface.co/openai/whisper-large-v3}{whisper-large-v3};
+    \item pyannote SpeakerDiarization \\
+    Link to model: \href{https://huggingface.co/pyannote/speaker-diarization-3.0}{speaker-diarization-3.0};
+\end{itemize}
+
+\section{Dialogue Segmentation}
+
+\section{Topic Extraction}
+
+The next stage in the pipeline is topic extraction. At this stage, $n$ topics will be extracted from each segment in the dialogues. 
+At this stage, past approaches have used probabilistic or embedding approaches, like BertTopic \cite{berttopic:2022}. 
+In our work, we will apply a generative approach using Large Language models, as in PromptTopic \cite{prompttopic:2023}.
+
+We will try several Large Language Models:
+\begin{itemize}
+    \item MetaAI LLaMA3 8b \\
+    Link to model: \href{https://huggingface.co/meta-llama/Meta-Llama-3-8B-Instruct}{meta-llama3-8b}
+    \item Qwen1.5 7b \\
+    Link to model: \href{https://huggingface.co/Qwen/Qwen1.5-7B-Chat}{Qwen1.5 7B}
+    \item Mistral 7b \\
+    Link to model: \href{https://huggingface.co/mistralai/Mistral-7B-Instruct-v0.1}{Mistral 7B}
+\end{itemize}
+
+The prompt for LLM will look like this:
+
+\begin{lstlisting}
+    <|user|> 
+    utterance_1
+    utterance_2
+    ...
+    utterance_n. 
+    Please list of high-level topic words 
+    in the following dialogue:
+    <|assistant|>
+
+\end{lstlisting}
+
+where $utterance$ is one phrase of the participant in the dialogue, \\$M$ - is the number of words. The appearance of the keywords <|user|>, <|assistant|> dependes on type of LLM.
+
+\section{Topic Evolution}
 
@@
     \end{center}
+    \captionof{table}{Datasets}
 \end{table}
@@
 
-At the text clustering stage, comparisons are made between models of non-alternative baselines, because at this stage we are using a ready-made algorithm and making a minimum of changes.
+At the text clustering stage, comparisons are made between models of non-alternative baselines, 
+because at this stage we are using a ready-made algorithm and making a minimum of changes.
 As for topic extraction, our LLM approach will be compared with BertTopic and LSA. \\
-
-\textbf{Pipeline configuration}
-
-MODEL1 will be used in the ASR for audio datasets. 
-At the dialogue clustering stage, BERT-based fine-tuning models will be used. 
-For topic extraction, the LlaMA 3 8B and Mistral 7B models will be used, in addition to experiments with a smaller model - phi3.
 
@@
 
-\chapter{Conclusion}
+\chapter*{Conclusion}
 

### 2024-05-07 paper/main.tex
metrics added

@@
 
-The main objectives of the research work is to create and apply Dynamic Topic Modeling pipeline for audio and text dialogues. 
+The main objectives of the research work is to create and apply Dynamic Topic Modeling pipeline based on LLM for topic extraction for audio and text dialogues. 
 To solve this problem, you will need to complete the following tasks 
 \begin{enumerate}
-    \item The identification of suitable datasets for experiments
-    \item Investigate open-source solutions for ASR, dialogue clustering, topic extraction and topic evolution.
-    \item An exploration of the possibilities of using large language models for some stages and the selection of models suitable for training in limited resources
-    \item The development of the architecture of all stages and their interconnections
-    \item Selection of appropriate metrics and measurement of the quality of the approaches used at each stage
+    \item The identification of suitable datasets for experiments;
+    \item The investigation of the best suitable open-source solutions for ASR, dialogue clustering, topic extraction and topic evolution;
+    \item The exploration of the possibilities of using Large Language Models for some stages and the selection of models suitable for training in limited resources;
+    \item The selection of appropriate metrics and measurement of the quality of the approaches used at each stage;
 \end{enumerate}
@@
 
-\newpage
 \textbf{LLM} is a type of artificial intelligence(AI) algorithm that use deep learning techniques and massively large data sets to understand, summarize, generate, and predict new content. 
@@
 \newpage
-In april 2024 was released LLaMA3 8B and 70B models with some new features: larger tokenizer vocabulary 128k agains 32k, the dataset for pre-training contains 8 times more sampels on over 15 trillion tokens on a new mix of publicly available online data
+In april 2024 was released LLaMA3 8B and 70B models with some new features: larger tokenizer vocabulary 128k agains 32k, the dataset for pre-training contains 8 times more sampels on over 15 trillion tokens on a new mix of publicly available online data.
 
@@
     \centering
-    \includegraphics[scale=0.7  ]{img/whisper.png}
+    \includegraphics[scale=0.7]{img/whisper.png}
     \caption{Whisper architecture}
@@
 \section{Evaluation Metrics}
+
+\subsection{ASR}
+
+Word Error Rate (WER) is the metric for speech recognition. This metric counts errors at the word level. 
+A substitutions (S), insertions (I) and deletions (D) error are counted for each word.
+For example:
+
+\begin{figure}[h]
+    \includegraphics[scale=1]{img/wer.png}
+\end{figure}
+
+Word "sat" predicted as "sit", it is substitutions error, and the word "mat" and was not recognized at all, it is deletion error.\\
+To calculate the metric, take the formula:
+
+$WER = \frac{S + I + D}{N}$, \\
+where N is total number of words in the example.
+
+Note that the word "sit" has only one incorrect character, but the whole word is marked as incorrect.
+This is a defining feature of the WER: spelling errors are heavily penalised, no matter how minor.
 
@@
     <|user|> 
-    utterance_1
-    utterance_2
-    ...
-    utterance_n. 
+    <dialogue> 
     Please list of high-level topic words 
@@
     <|assistant|>
-
 \end{lstlisting}
 
-where $utterance$ is one phrase of the participant in the dialogue, \\$M$ - is the number of words. The appearance of the keywords <|user|>, <|assistant|> dependes on type of LLM.
+The appearance of the keywords <|user|>, <|assistant|> dependes on type of LLM.
 

### 2024-05-13 paper/main.tex
add speaker diarizer

@@
     \item The identification of suitable datasets for experiments;
-    \item The investigation of the best suitable open-source solutions for ASR, dialogue clustering, topic extraction and topic evolution;
+    \item The investigation of the best suitable open-source solutions for speaker diarisation, dialogue clustering, topic extraction and topic evolution;
     \item The exploration of the possibilities of using Large Language Models for some stages and the selection of models suitable for training in limited resources;
@@
 
-\section{Automatic Speech Recognition}
+\section{Speaker diarization}
 
-Since one of the significant tasks is to use DTM for audio data, the first, but optional, stage of our pipeline will be a model for automatic speech recognition. 
-In this section, we will look at the Whisper \cite{whisper:2022} model from OpenAI.
+Since one of the significant tasks is to use DTM for audio dialogues, the first, but optional, stage of our pipeline will be a model for speaker diarization.
+
+Speaker diarisation is the process of automatically segmenting and identifying different speakers in an audio recording. 
+The goal of speaker diarisation is to divide the audio stream into non-homogeneous segments, each segment corresponding to a specific speaker or speaker turn.
+
+The fundamental process of building a speaker diarization system involves the following steps:
+
+\begin{itemize}
+    \item Segmentation: The audio recording is divided into differente segments based on conditions like silence, pauses, and changes in speaker turns.
+    \item Speaker Embedding: Embeddings are extracted for each segment to represent the speech features of unique speaker.
+    \item Clustering: The extracted embeddings are clustered based on similarity, with segments from the same speaker forming a single cluster.
+    \item Labeling: Once the clustering is complete, each segment is assigned a label representing the speaker's identity.
+Segments in the same cluster are assigned the same label, indicating they belong to the same speaker.
+\end{itemize}
+
+As ASR model for this stage, we will look at the Whisper \cite{whisper:2022} model from OpenAI.
 
@@
 A decoder is trained to predict the corresponding text caption, intermixed with special tokens that direct the single model to perform tasks such as language identification, phrase-level timestamps, multilingual speech transcription, and to-English speech translation.
-
-\subsection{Speaker-diarization}
 
@@
 Word Error Rate (WER) is the metric for speech recognition. This metric counts errors at the word level. 
-A substitutions (S), insertions (I) and deletions (D) error are counted for each word.
+A substitutions (S), insertions (I) and deletions (D) error are counted for each word. \\
 For example:
@@
 The lower the value, the better.\\
-    $WD(pred, true) = \frac{1}{N - k}\sum_{i = 1}^{N - k}(\left\lvert bounds(true_{i}, true_{i + k}) - bounds(pred_{i}, pred_{i + k}) \right\rvert > 0)$, \\ 
+    $WD(pred, true) = \frac{1}{N - k}\sum_{i = 1}^{N - k}(\left\lvert bounds(true_{i}, true_{i + k}) - bounds(pred_{i}, pred_{i + k})\right\rvert > 0)$, \\ 
     where $bounds(s, s + i)$ is the number of boundaries between $i$ and $j$ positions, $N$ - number of sentences in the text.
@@
 
-\section{ASR}
+\section{Speaker Diarization}
 
-The first optional stage is the Automatic Speech Recognition. 
-For audio dialogues, text is extracted from the audio track before segmentation. 
-For this stage, open-source:
+The first optional stage is the Speaker Diarization. 
+For audio dialogues, text is extracted from the audio track before topic segmentation. 
+For this stage we use:
 \begin{itemize}
@@
     Link to model: \href{https://huggingface.co/openai/whisper-large-v3}{whisper-large-v3};
-    \item pyannote SpeakerDiarization \\
+    \item pyannote SpeakerDiarization library \\
     Link to model: \href{https://huggingface.co/pyannote/speaker-diarization-3.0}{speaker-diarization-3.0};
@@
 \textbf{Datasets}
-
 For the experiments we will use two types of datasets: text dialogues and audio dialogues. 
@@
 \begin{itemize}
-    \item Various approaches for dialogue segmentation, topic mining, and topic evolution have been explored;
+    \item Various approaches for topic diarization, dialogue segmentation, topic extraction, and topic evolution have been explored;
     \item The search for suitable text and audio datasets was conducted, and the selected datasets were analysed and preprocessed for experiments at all stages and for the entire pipeline.

### 2024-05-13 paper/main.tex
add lib + change picture
add seg

@@
 \section{Improving Utterance-Pair Coherence Scoring}
+\label{seg}
 
@@
     Link to model: \href{https://huggingface.co/openai/whisper-large-v3}{whisper-large-v3};
-    \item pyannote SpeakerDiarization library \\
-    Link to model: \href{https://huggingface.co/pyannote/speaker-diarization-3.0}{speaker-diarization-3.0};
+    \item pyannote - speaker diarization library \cite{pyannote:2023} \\
+    Link to model: \href{https://github.com/pyannote/pyannote-audio}{[pyannote]};
 \end{itemize}
@@
 \section{Dialogue Segmentation}
+
+The algorithm \ref{seg} will be used to dialogues segmentation. 
+The implementation of the algorithm will be taken from the \href{https://github.com/lxing532/Dialogue-Topic-Segmenter}{official github repository}. 
+Part of the code with the model inference will be adapted to the pipeline.
 

### 2024-05-14 paper/main.tex
change pipeline and metrics

@@
 
-\subsection{ASR}
+\subsection{Speaker diarization}
 
-Word Error Rate (WER) is the metric for speech recognition. This metric counts errors at the word level. 
-A substitutions (S), insertions (I) and deletions (D) error are counted for each word. \\
-For example:
+The standard metric for speaker diarization problem is the Diarization error rate \:
+This metric computes the error in diarization in terms of a duration ratio, which means that a perfectly aligned hypothesis and reference diarization would result in DER=0. It sums up 3 error:
+\begin{itemize}
+    \item False Alarm - speech segment predicted where there is no speaker. It is error from ASR models.
+    \item Missed Detection - no speech detected where there is a speaker. It is error from ASR models.
+    \item Confusion	Speech - the wrong cluster. It is error from the clustering model
+\end{itemize}
 
 \begin{figure}[h]
-    \includegraphics[scale=1]{img/wer.png}
+    \centering
+    \includegraphics[scale=0.8]{img/der.png}
+    \caption{Diarization error rate }
 \end{figure}
 
-Word "sat" predicted as "sit", it is substitutions error, and the word "mat" and was not recognized at all, it is deletion error.\\
-To calculate the metric, take the formula:
-
-$WER = \frac{S + I + D}{N}$, \\
-where N is total number of words in the example.
-
-Note that the word "sit" has only one incorrect character, but the whole word is marked as incorrect.
-This is a defining feature of the WER: spelling errors are heavily penalised, no matter how minor.
+Mathematically, the metric can be written: 
+\begin{center}
+    $DER = \frac{False Alarm  +  Missed Detection  +  Confusion}{Total}$    
+\end{center}
 
@@
             Tiage &  7939/-/1000 & 14.7  & null \\ 
+            \hline
             Product Meetings (QMSum) & 137 & 535.6 & null \\ 
@@
 As for topic extraction, our LLM approach will be compared with BertTopic and LSA. \\
+
+\section{Speaker diarization}
+
+\begin{center}
+    \begin{tabular}{ | c | c | c | c | c | c | c | c | c | c |} 
+        \hline
+        \multicolumn{1}{| c |}{} & \multicolumn{3}{| c |}{Product} &  \multicolumn{3}{| c |}{Academic} & \multicolumn{3}{| c |}{Committee} \\
+        \hline
+        model  & Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
+        \hline
+        whisper-large +  &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-base CLS &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1  \\ 
+        bert-large NSP & Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
+        opt-128 &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
+        dialoGPT-medium &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
+        \hline
+    \end{tabular}
+\end{center}
 

### 2024-05-16 paper/main.tex
update experimetns

@@
     <dialogue> 
-    Please list of high-level topic words 
-    in the following dialogue:
+    Give me N topic words 
+    from the following dialogue:
     <|assistant|>
@@
 
-The appearance of the keywords <|user|>, <|assistant|> dependes on type of LLM.
+where N is number of topic words, the appearance of the keywords <|user|>, <|assistant|> dependes on type of LLM.
+
+Moreover, in experiments, models are tested not only in zero-hour mode, but also in few-shot.
+A few-short is a technique in which there are examples of a query with an answer before the main question.
+In this case the prompt for LLM will look like this:
+
+\begin{lstlisting}
+    <|user|> 
+    <example_dialogue1> 
+    Give me N topic words 
+    from the following dialogue:
+    <example_answer1>
+
+    <example_dialogue2> 
+    Give me N topic words 
+    from the following dialogue:
+    <example_answer2>
+
+    <dialogue> 
+    Give me N topic words 
+    from the following dialogue:
+    <|assistant|>
+\end{lstlisting}
 
@@
             SuperDialseg & 6984/1322/1322 & 13 & 239 \\ 
-            Tiage &  7939/-/1000 & 14.7  & null \\ 
+            Tiage &  7939/-/1000 & 14.7  & - \\ 
             \hline
-            Product Meetings (QMSum) & 137 & 535.6 & null \\ 
-            Academic Meetings (QMSum) & 59 & 819.0 & null \\ 
-            Comemitie Meetings (QMSum) & 36 & 207.7 & null \\
+            Product Meetings (QMSum) & 137 & 535.6 & - \\ 
+            Academic Meetings (QMSum) & 59 & 819.0 & - \\ 
+            Comemitie Meetings (QMSum) & 36 & 207.7 & - \\
             \hline
@@
 \end{table}
-
-\textbf{Baseline Models}
 
@@
 \begin{center}
-    \begin{tabular}{ | c | c | c | c | c | c | c | c | c | c |} 
+    \begin{tabular}{ | c | c | c | c |} 
         \hline
-        \multicolumn{1}{| c |}{} & \multicolumn{3}{| c |}{Product} &  \multicolumn{3}{| c |}{Academic} & \multicolumn{3}{| c |}{Committee} \\
+        \multicolumn{1}{| c |}{} & \multicolumn{1}{| c |}{Product} &  \multicolumn{1}{| c |}{Academic} & \multicolumn{1}{| c |}{Committee} \\
         \hline
-        model  & Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
+        model  & \multicolumn{3}{| c |}{DER} \\ 
         \hline
-        whisper-large +  &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
-        bert-base CLS &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1  \\ 
-        bert-large NSP & Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
-        opt-128 &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
-        dialoGPT-medium &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
+        whisper-large & der & der & der \\ 
+        pyannote-diarisation &  der & der & der \\ 
+        bert-large NSP & der & der & der \\ 
         \hline
     \end{tabular}
+    \captionof{table}{Speaker diarization evaluation}
 \end{center}
@@
         \hline
-        model &  Pk & wd & f1 & Pk & wd & f1 \\ 
+        model & Pk $\downarrow$ & wd $\downarrow$ & f1 & Pk $\downarrow$ & wd $\downarrow$ & f1 \\ 
+        \hline
+        GreedySeg & 0.492 & 0.508 & 0.181 & 0.497 & 0.501 & 0.359 \\
+        TeT & 0.469 & 0.488 & 0.204 & 0.446 & 0.458 & 0.378 \\
+        TeT + Embedding & 0.469 & 0.488 & 0.204 & 0.446 & 0.458 & 0.378 \\ 
+        TeT + NSP & 0.425 & 0.439 & 0.285 & 0.522 & 0.531 & 0.201 \\
         \hline
         bert-base NSP & Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-large NSP & Pk & wd & f1 & Pk & wd & f1 \\ 
         bert-base CLS & Pk & wd & f1 & Pk & wd & f1 \\ 
-        bert-large NSP & Pk & wd & f1 & Pk & wd & f1 \\ 
         opt-128 & Pk & wd & f1 & Pk & wd & f1 \\ 
@@
     \end{tabular}
+    \captionof{table}{Text datasets segmentation}
 \end{center}
-
-In terms of textual datasets ...
 
@@
         \hline
-        model  & Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
+        model  & Pk $\downarrow$ & wd $\downarrow$ & f1 & Pk $\downarrow$ & wd $\downarrow$ & f1 & Pk $\downarrow$ & wd $\downarrow$ & f1 \\ 
+        \hline
+        GreedySeg &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
+        TeT & Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1  \\ 
+        TeT + Embedding & Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
+        TeT + NSP &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\  
         \hline
@@
     \end{tabular}
+    \captionof{table}{Audio datasets segmentation}
 \end{center}
@@
         \hline
+        BertTopic & c & d & c & d \\
+        LSA  & c & d & c & d \\
+        \hline
         LLaMA3 8B & c & d & c & d \\ 
@@
         \hline
-        BertTopic & c & d & c & d \\
-        LSA  & c & d & c & d \\
-        \hline
     \end{tabular}
+    \captionof{table}{Text datasets topics extraction}
 \end{center}
@@
         \hline
+        BertTopic & c & d & c & d & c & d \\ 
+        LSA & c & d & c & d & c & d \\ 
+        \hline
         LLaMA3 8B & c & d & c & d & c & d \\ 
@@
         \hline
-        BertTopic & c & d & c & d & c & d \\ 
-        LSA & c & d & c & d & c & d \\ 
-        \hline
     \end{tabular}
+    \captionof{table}{Audion datasets topics extraction}
 \end{center}
@@
 \begin{itemize}
-    \item Various approaches for topic diarization, dialogue segmentation, topic extraction, and topic evolution have been explored;
+    \item Various approaches for topic diarization, dialogue segmentation, topic extraction, and topic evolution have been explored and the best of them were adopted to DTM pipeline;
     \item The search for suitable text and audio datasets was conducted, and the selected datasets were analysed and preprocessed for experiments at all stages and for the entire pipeline.

### 2024-05-18 paper/main.tex
update paper

@@
 For text dialogues, we will use tiage, compiled from the personal chat corpus and manually annotated, and superdialseg, compiled from several smaller datasets: doc2dial and multidoc2dial. 
-As for the audio dialogues, we will use QMSum, which includes the audio recording corpuses - AMI and ISCI and is divided into three types of meetings (product, academic, committee). \\
+As for the audio dialogues, we will use QMSum, which includes the audio recording corpuses - AMI is product meetings and ICSI is academic meetings. \\
 
@@
             \hline
-            Product Meetings (QMSum) & 137 & 535.6 & - \\ 
-            Academic Meetings (QMSum) & 59 & 819.0 & - \\ 
-            Comemitie Meetings (QMSum) & 36 & 207.7 & - \\
+            Product Meetings (AMI) & 137 & 535.6 & - \\ 
+            Academic Meetings (ICSI) & 59 & 819.0 & - \\ 
             \hline
@@
 \begin{center}
-    \begin{tabular}{ | c | c | c | c |} 
+    \begin{tabular}{ | c | c | c |} 
         \hline
-        \multicolumn{1}{| c |}{} & \multicolumn{1}{| c |}{Product} &  \multicolumn{1}{| c |}{Academic} & \multicolumn{1}{| c |}{Committee} \\
+        \multicolumn{1}{| c |}{model} & \multicolumn{1}{| c |}{AMI} &  \multicolumn{1}{| c |}{ICSI} \\
         \hline
-        model  & \multicolumn{3}{| c |}{DER} \\ 
-        \hline
-        whisper-large & der & der & der \\ 
-        pyannote-diarisation &  der & der & der \\ 
-        bert-large NSP & der & der & der \\ 
+        pyannote-speaker-diarization-3.1  & 18.5 & 17.38 \\  
         \hline
@@
         TeT & 0.469 & 0.488 & 0.204 & 0.446 & 0.458 & 0.378 \\
-        TeT + Embedding & 0.469 & 0.488 & 0.204 & 0.446 & 0.458 & 0.378 \\ 
         TeT + NSP & 0.425 & 0.439 & 0.285 & 0.522 & 0.531 & 0.201 \\
@@
         bert-base CLS & Pk & wd & f1 & Pk & wd & f1 \\ 
-        opt-128 & Pk & wd & f1 & Pk & wd & f1 \\ 
         dialoGPT-medium & Pk & wd & f1 & Pk & wd & f1 \\ 
@@
 \begin{center}
-    \begin{tabular}{ | c | c | c | c | c | c | c | c | c | c |} 
+    \begin{tabular}{ | c | c | c | c | c | c | c | } 
         \hline
-        \multicolumn{1}{| c |}{} & \multicolumn{3}{| c |}{Product} &  \multicolumn{3}{| c |}{Academic} & \multicolumn{3}{| c |}{Committee} \\
+        \multicolumn{1}{| c |}{} & \multicolumn{3}{| c |}{AMI} &  \multicolumn{3}{| c |}{ICSI} \\
         \hline
-        model  & Pk $\downarrow$ & wd $\downarrow$ & f1 & Pk $\downarrow$ & wd $\downarrow$ & f1 & Pk $\downarrow$ & wd $\downarrow$ & f1 \\ 
+        model  & Pk $\downarrow$ & wd $\downarrow$ & f1 & Pk $\downarrow$ & wd $\downarrow$ & f1 \\ 
         \hline
-        GreedySeg &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
-        TeT & Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1  \\ 
-        TeT + Embedding & Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
-        TeT + NSP &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\  
+        GreedySeg &  Pk & wd & f1 & Pk & wd & f1 \\ 
+        TeT & Pk & wd & f1 & Pk & wd & f1 \\ 
+        TeT + NSP &  Pk & wd & f1 & Pk & wd & f1 \\  
         \hline
-        bert-base NSP &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
-        bert-base CLS &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1  \\ 
-        bert-large NSP & Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
-        opt-128 &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
-        dialoGPT-medium &  Pk & wd & f1 & Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-base NSP &  Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-base CLS &  Pk & wd & f1 & Pk & wd & f1 \\ 
+        bert-large NSP & Pk & wd & f1 & Pk & wd & f1 \\ 
+        opt-128 &  Pk & wd & f1 & Pk & wd & f1 \\ 
+        dialoGPT-medium &  Pk & wd & f1 & Pk & wd & f1 \\ 
         \hline
@@
 
+\subsection{Zero-shot prompting}
+
 \begin{center}
@@
         \hline
-        BertTopic & c & d & c & d \\
-        LSA  & c & d & c & d \\
+        LSA  & 0.00951 & 0.25824 & -0.01929 & 0.420026 \\
+        BertTopic & \cellcolor{green!60} 0.02751 & 0.34823 & \cellcolor{green!60} 0.01929 & 0.420026 \\
         \hline
-        LLaMA3 8B & c & d & c & d \\ 
-        Qwen1.5 7B & c & d & c & d \\ 
-        Mistral 7B & c & d & c & d \\ 
+        LLaMA3 8B & \cellcolor{green!20} 0.01621 & \cellcolor{green!60} 0.80707 & \cellcolor{green!20} 0.005887 & \cellcolor{green!60} 0.49525 \\ 
+        Qwen1.5 7B & 0.015453 & \cellcolor{green!20} 0.72169 & -0.07682 & \cellcolor{green!20} 0.47777 \\ 
+        Mistral 7B & 0.002248 & 0.63344 & -0.01689 & 0.034215 \\ 
         \hline
@@
 
-In terms of textual datasets ...
+\begin{center}
+    \begin{tabular}{ | c | c | c | c | c | c | c |} 
+        \hline
+        \multicolumn{1}{| c |}{} & \multicolumn{2}{| c |}{Product} &  \multicolumn{2}{| c |}{Academic} \\
+        \hline
+        model  & TC & TD & TC & TD  \\ 
+        \hline
+        LSA & 0.001375 & 0.009648 & -0.00172 & 0.013490 \\ 
+        BertTopic & \cellcolor{green!60} 0.02485 & 0.013996 & \cellcolor{green!60} 0.27210 & 0.029090 \\ 
+        \hline
+        LLaMA3 8B & \cellcolor{green!20} 0.11409 & \cellcolor{green!60} 0.50863 & \cellcolor{green!20} 0.14734 & \cellcolor{green!60} 0.76137\\ 
+        Qwen1.5 & 0.10328 & \cellcolor{green!20}0.38641 & 0.00237 & \cellcolor{green!20}0.7185 \\ 
+        Mistral 7B & 0.084103 & 0.29852 & -0.11318 & 0.5761 \\ 
+        \hline
+    \end{tabular}
+    \captionof{table}{Audio datasets topics extraction}
+\end{center}
+
+The result of experiments in zero-shot that LlaMA 3 8B demonstrate the best performance, it is far ahead of everyone in Topic Diversity and competes in Topic Coherence with BertTopic. 
+A strong drawdown in quality relative to extractive model was also noticed in the article PromptTopic \cite{prompttopic:2023}, where the models also lost to BertTopic in Topic Coherence.
+
+\subsection{Few-shot prompting}
 
@@
         \hline
-        \multicolumn{1}{| c |}{} & \multicolumn{2}{| c |}{Product} &  \multicolumn{2}{| c |}{Academic} & \multicolumn{2}{| c |}{Committee} \\
+        \multicolumn{1}{| c |}{} & \multicolumn{2}{| c |}{AMI} &  \multicolumn{2}{| c |}{ICSI} \\
         \hline
-        model  & TC & TD & TC & TD & TC & TD  \\ 
+        LLaMA3 8B zero-shot & 0.11409 & 0.50863 & 0.14734 & 0.76137\\ 
+        LLaMA3 8B few-shot & \cellcolor{green!20} 0.12129 & \cellcolor{green!20} 0.51291 & \cellcolor{red!10} 0.13985 & \cellcolor{green!60} 0.78201\\ 
         \hline
-        BertTopic & c & d & c & d & c & d \\ 
-        LSA & c & d & c & d & c & d \\ 
+        Qwen1.5 zero-shot & 0.10328 & 0.38641 & 0.00237 & 0.7185 \\ 
+        Qwen1.5 few-shot & 0.10190 & 0.38875 &  0.00239 & \cellcolor{green!20}0.7196 \\ 
         \hline
-        LLaMA3 8B & c & d & c & d & c & d \\ 
-        Qwen1.5 7B & c & d & c & d & c & d \\ 
-        Mistral 7B & c & d & c & d & c & d \\ 
+        Mistral zero-shot & 0.084103 & 0.29852 & -0.11318 & 0.5761 \\ 
+        Mistral few-shot & \cellcolor{red!10} 0.065381 & \cellcolor{green!20} 0.31920 & \cellcolor{green!20} -0.08673 & 0.5761 \\ 
         \hline
     \end{tabular}
-    \captionof{table}{Audion datasets topics extraction}
+    \captionof{table}{Comparisons of zero-shot and few-shot results}
 \end{center}
 
-Speaking about results on audio corpora ...
+Experiments with few-short prompts did not show a significant increase in metrics, but it became much easier to get topic words from the model output because the model understood the format in which we wanted to get a list of keywords.
+As a result, it doesn not mean much on its own, but as part of a pipeline where everything is working together, it's a significant improvement.
+
+\subsection{Fine-tuning}
 
@@
 The LLM and other models require a significant amount of computing resources, including GPUs.
-Due to the limited availability of resources, the models were not larger than 8 billion parameters and did not exceed 16 GB during training and inference at all stages.
+Due to the limited availability of resources, the models were not larger than 8 billion parameters and did not exceed 32 GB during training and inference at all stages.
 Additionally, because of the absence of financial opportunities, open-source models were employed in the work.

### 2024-05-18 src/segmentation/scripts/bert_segmentation.sh

@@ -1,7 +1,8 @@
-DATASET=./datasets/eval/dialseg_711.json
-ENCODER_BERT_MODEL=dse-roberta-base
+DATASET=./datasets/eval/meeting_ami.json
+ENCODER_BERT_MODEL=aws-ai/dse-bert-base
 
-python ./src/segmentation_module/segment_bert.py \
+python src/segmentation/segment_bert.py \
 --dataset $DATASET \
 --text_encoder $ENCODER_BERT_MODEL \
---mode NSP
+--mode SC \
+--alpha 2

### 2024-05-18 src/segmentation/segment_bert.py

@@ -50,11 +50,15 @@ if __name__ == "__main__":
     parser.add_argument('-m', '--mode', 
                         help='sequence classification (SC) / next sentence prediction (NSP) / coherence model (CM)', 
                         default='CM')
+    parser.add_argument('-a', '--alpha',
+                        help='choose alpha for alpha-search', 
+                        default=1)
     args = parser.parse_args()
     
     data = args.dataset
     text_encoder_name = args.text_encoder
     mode = args.mode
+    best_alpha = np.float64(args.alpha)
     device = 'cuda' if torch.cuda.is_available() else 'cpu'
 
     # load encoder model
@@ -79,10 +83,7 @@ if __name__ == "__main__":
             dev_data.append(dialogue)
         else:
             test_data.append(dialogue)
-            
-    # Hyper-parameter (alpha) search on dev set
-    best_alpha, best_pk = alpha_search(
-        dev_data, text_encoder, tokenizer, mode, device, -2, 2, 0.1)
+
     print('[INFO] The loaded text encoder is: ', text_encoder_name)
     print('[INFO] The best hyper-parameter (alpha): ', best_alpha)
 

### 2024-05-18 data/dailydialog/dailydialog_parser.py
add daily_dial

@@ -1,0 +1,107 @@
+import os, sys, getopt, gzip
+
+def parse_data(in_dir, out_dir):
+
+    # Finding files
+    if in_dir.endswith('train'):
+        dial_dir = os.path.join(in_dir, 'dialogues_train.txt')
+        emo_dir = os.path.join(in_dir, 'dialogues_emotion_train.txt')
+        act_dir = os.path.join(in_dir, 'dialogues_act_train.txt')
+    elif in_dir.endswith('validation'):
+        dial_dir = os.path.join(in_dir, 'dialogues_validation.txt')
+        emo_dir = os.path.join(in_dir, 'dialogues_emotion_validation.txt')
+        act_dir = os.path.join(in_dir, 'dialogues_act_validation.txt')
+    elif in_dir.endswith('test'):
+        dial_dir = os.path.join(in_dir, 'dialogues_test.txt')
+        emo_dir = os.path.join(in_dir, 'dialogues_emotion_test.txt')
+        act_dir = os.path.join(in_dir, 'dialogues_act_test.txt')
+    else:
+        print("Cannot find directory")
+        sys.exit()
+
+    out_dial_dir = os.path.join(out_dir, 'dial.txt')
+    out_emo_dir = os.path.join(out_dir, 'emo.txt')
+    out_act_dir = os.path.join(out_dir, 'act.txt')
+
+    # Open files
+    in_dial = open(dial_dir, 'r')
+    in_emo = open(emo_dir, 'r')
+    in_act = open(act_dir, 'r')
+
+    out_dial = open(out_dial_dir, 'w')
+    out_emo = open(out_emo_dir, 'w')
+    out_act = open(out_act_dir, 'w')
+
+
+    for line_count, (line_dial, line_emo, line_act) in enumerate(zip(in_dial, in_emo, in_act)):
+        seqs = line_dial.split('__eou__')
+        seqs = seqs[:-1]
+
+        emos = line_emo.split(' ')
+        emos = emos[:-1]
+
+        acts = line_act.split(' ')
+        acts = acts[:-1]
+        
+        seq_count = 0
+        seq_len = len(seqs)
+        emo_len = len(emos)
+        act_len = len(acts)
+    
+        if seq_len != emo_len or seq_len != act_len:
+            print("Different turns btw dialogue & emotion & acttion! ", line_count+1, seq_len, emo_len, act_len)
+            sys.exit()
+
+        for seq, emo, act in zip(seqs, emos, acts):
+
[…]

### 2024-05-19 paper/main.tex
update experiments

@@
         GreedySeg & 0.492 & 0.508 & 0.181 & 0.497 & 0.501 & 0.359 \\
-        TeT & 0.469 & 0.488 & 0.204 & 0.446 & 0.458 & 0.378 \\
-        TeT + NSP & 0.425 & 0.439 & 0.285 & 0.522 & 0.531 & 0.201 \\
-        \hline
-        bert-base NSP & Pk & wd & f1 & Pk & wd & f1 \\ 
-        bert-large NSP & Pk & wd & f1 & Pk & wd & f1 \\ 
-        bert-base CLS & Pk & wd & f1 & Pk & wd & f1 \\ 
-        dialoGPT-medium & Pk & wd & f1 & Pk & wd & f1 \\ 
+        GraphSeg & 0.462 & 0.466 & 0.239 & 0.496 & 0.515 & 0.238 \\
+        TextTilling & 0.469 & 0.488 & 0.204 & 0.446 & 0.458 & 0.378 \\
+        TeT + BERT & 0.425 & 0.439 & 0.285 & 0.522 & 0.531 & 0.201 \\
+        \hline\ 
+        Our solution BERT-NSP & Pk & wd & f1 & Pk & wd & f1 \\ 
+        Our solution BERT-CLS & Pk & wd & f1 & Pk & wd & f1 \\ 
         \hline
@@
         \hline
-        GreedySeg &  Pk & wd & f1 & Pk & wd & f1 \\ 
-        TeT & Pk & wd & f1 & Pk & wd & f1 \\ 
-        TeT + NSP &  Pk & wd & f1 & Pk & wd & f1 \\  
+        GreedySeg & 0.885 & 0.792 & 0.0641 & 0.762 & 0.730 & 0.109 \\
+        GraphSeg & Pk & wd & f1 & Pk & wd & f1 \\
+        TextTilling & Pk & wd & f1 & Pk & wd & f1 \\
+        TeT + BERT & Pk & wd & f1 & Pk & wd & f1 \\ 
         \hline
-        bert-base NSP &  Pk & wd & f1 & Pk & wd & f1 \\ 
-        bert-base CLS &  Pk & wd & f1 & Pk & wd & f1 \\ 
-        bert-large NSP & Pk & wd & f1 & Pk & wd & f1 \\ 
-        opt-128 &  Pk & wd & f1 & Pk & wd & f1 \\ 
-        dialoGPT-medium &  Pk & wd & f1 & Pk & wd & f1 \\ 
+        Our solution BERT-NSP & Pk & wd & f1 & Pk & wd & f1 \\ 
+        Our solution BERT-CLS & Pk & wd & f1 & Pk & wd & f1 \\  
         \hline
@@
 
-Speaking about results on audio corpora ...
+The selected solution shows the best metrics on the datasets we selected. It is important to understand that we have researched open source solutions.
 
@@
         \hline
-        model & TC & TD & TC & TD \\ 
+        model & TD & TC & TD & TC  \\ 
         \hline
-        LSA  & 0.00951 & 0.25824 & -0.01929 & 0.420026 \\
-        BertTopic & \cellcolor{green!60} 0.02751 & 0.34823 & \cellcolor{green!60} 0.01929 & 0.420026 \\
+        LSA  & 0.25824 & 0.00951 & 0.420026 & -0.01929 \\
+        BertTopic & 0.34823 & \cellcolor{green!60} 0.02751 & 0.420026 & \cellcolor{green!60} 0.01929 \\
         \hline
-        LLaMA3 8B & \cellcolor{green!20} 0.01621 & \cellcolor{green!60} 0.80707 & \cellcolor{green!20} 0.005887 & \cellcolor{green!60} 0.49525 \\ 
-        Qwen1.5 7B & 0.015453 & \cellcolor{green!20} 0.72169 & -0.07682 & \cellcolor{green!20} 0.47777 \\ 
-        Mistral 7B & 0.002248 & 0.63344 & -0.01689 & 0.034215 \\ 
+        LLaMA3 8B & \cellcolor{green!60} 0.80707 & \cellcolor{green!20} 0.01621 & \cellcolor{green!60} 0.49525 & \cellcolor{green!20} 0.005887 \\ 
+        Qwen1.5 7B & \cellcolor{green!20} 0.72169  & 0.015453 & \cellcolor{green!20} 0.47777  & -0.07682\\ 
+        Mistral 7B & 0.63344 & 0.002248 & 0.34215 & -0.01689 \\ 
         \hline
@@
         \hline
-        model  & TC & TD & TC & TD  \\ 
+
         \hline
-        LSA & 0.001375 & 0.009648 & -0.00172 & 0.013490 \\ 
-        BertTopic & \cellcolor{green!60} 0.02485 & 0.013996 & \cellcolor{green!60} 0.27210 & 0.029090 \\ 
+        LSA & 0.009648 & 0.001375 & 0.013490 & -0.00172 \\ 
+        BertTopic & 0.013996 & \cellcolor{green!60} 0.02485 & 0.029090 & \cellcolor{green!60} 0.27210 \\ 
         \hline
-        LLaMA3 8B & \cellcolor{green!20} 0.11409 & \cellcolor{green!60} 0.50863 & \cellcolor{green!20} 0.14734 & \cellcolor{green!60} 0.76137\\ 
-        Qwen1.5 & 0.10328 & \cellcolor{green!20}0.38641 & 0.00237 & \cellcolor{green!20}0.7185 \\ 
-        Mistral 7B & 0.084103 & 0.29852 & -0.11318 & 0.5761 \\ 
+        LLaMA3 8B & \cellcolor{green!60} 0.50863 & \cellcolor{green!20} 0.11409 & \cellcolor{green!60} 0.76137 & \cellcolor{green!20} 0.14734 \\ 
+        Qwen1.5  & \cellcolor{green!20} 0.38641 & 0.10328 & \cellcolor{green!20} 0.7185  & 0.00237\\ 
+        Mistral 7B & 0.29852 & 0.084103 & 0.5761 & -0.11318 \\ 
         \hline
@@
         \hline
-        LLaMA3 8B zero-shot & 0.11409 & 0.50863 & 0.14734 & 0.76137\\ 
-        LLaMA3 8B few-shot & \cellcolor{green!20} 0.12129 & \cellcolor{green!20} 0.51291 & \cellcolor{red!10} 0.13985 & \cellcolor{green!60} 0.78201\\ 
+        model & TD & TC & TD & TC  \\ 
         \hline
-        Qwen1.5 zero-shot & 0.10328 & 0.38641 & 0.00237 & 0.7185 \\ 
-        Qwen1.5 few-shot & 0.10190 & 0.38875 &  0.00239 & \cellcolor{green!20}0.7196 \\ 
+        LLaMA3 zero-shot & 0.50863 & 0.11409 & 0.76137 & 0.14734 \\ 
+        LLaMA3 few-shot & \cellcolor{green!20} 0.51291 & \cellcolor{green!20} 0.12129 & \cellcolor{green!60} 0.78201 & \cellcolor{red!10} 0.13985 \\ 
         \hline
-        Mistral zero-shot & 0.084103 & 0.29852 & -0.11318 & 0.5761 \\ 
-        Mistral few-shot & \cellcolor{red!10} 0.065381 & \cellcolor{green!20} 0.31920 & \cellcolor{green!20} -0.08673 & 0.5761 \\ 
+        Qwen1.5 zero-shot & 0.38641 & 0.10328 & 0.7185 & 0.00237 \\ 
+        Qwen1.5 few-shot & 0.38875 & 0.10190 & \cellcolor{green!20}0.7196 & 0.00239 \\ 
+        \hline
+        Mistral zero-shot & 0.29852 & 0.084103 & 0.5761 & -0.11318 \\ 
+        Mistral few-shot & \cellcolor{green!20} 0.31920 & \cellcolor{red!10} 0.065381 & 0.5761 & \cellcolor{green!20} -0.08673 \\ 
         \hline
@@
 Additionally, because of the absence of financial opportunities, open-source models were employed in the work.
-Should it be possible to reproduce all stages on more advanced models (such as GPT4 or LlaMA2 80b), the experimental results would be considerably more favourable.
+Should it be possible to reproduce all stages on more advanced models (such as GPT4 or LlaMA3 80b), the experimental results would be considerably more favourable.
 
@@
 \begin{itemize}
-    \item Various approaches for topic diarization, dialogue segmentation, topic extraction, and topic evolution have been explored and the best of them were adopted to DTM pipeline;
-    \item The search for suitable text and audio datasets was conducted, and the selected datasets were analysed and preprocessed for experiments at all stages and for the entire pipeline.
-    \item In the context of the dialog segmentation stage, the optimal open-source solution was identified and the BERT model was subjected to detailed analysis.
-This approach has been prepared and integrated into the general pipeline.
-    \item At the topic extraction stage, new approaches utilising generative models were investigated.
-Furthermore, a range of large language models, including LlaMA and Mistral, were examined, along with novel features that enhance the default architecture of the Transformer.
-    \item Topic Evolution
+    \item Suitable datasets were selected, experiments were carried out on them at all stages and quality metrics of the selected algorithms were obtained;
+    \item Among the available solutions, the best available metrics were selected.
+Experiments were carried out with each of the algorithms at each stage on ready-made models, pre-trained and trained from scratch.
+The best models were added to the final pipeline;;
+    \item At the topic extraction stage, a study was conducted on the use of LLM.
+The best open-source LLMs were selected for experiments with zero-shot, feat-shot and fine-tuning.
+The generative approach using LLM shows the results of the best available extractive solutions for topic modeling;
     \item A suitable set of metrics was identified and selected to evaluate the quality of each stage.

### 2024-05-20 paper/main.tex
udapte experiments

@@
 
+In the pipeline, topic evolution can be used in two formats. 
+The first format allows you to analyze the change of topics in a single dialog. In this format, graphs of changes in the set of topics in the order of the segments will be plotted.
+
+The second format analyzes the entire dataset. 
+In this case, either a special mark will be used for timestamp analysis along with each dialog, or the analysis will be performed in the order in which the samples are located in the dataset.
+
+With a large number of topic words, the N most common ones will be selected to visualize their changes.
+
 \chapter{Experiments}
@@
         \hline
-        pyannote-speaker-diarization-3.1 & 18.5 & 17.38 \\  
+        pyannote-speaker-diarization-2.1 & \cellcolor{green!20} 18.9 & 16.4 \\  
+        pyannote-speaker-diarization-3.1 & 18.8 & \cellcolor{green!20} 16.7 \\
+        whisper + pretrain embeds & \multirow{2}{*}{14.7} & \multirow{2}{*}{13.9} \\
+        + agglomerative cluter & & \\
         \hline
@@
 \end{center}
+
+The finished models from pyannote showed better results than the assembled pipeline from whisper and pretraining embeddings. 
+Therefore, it was decided to include pyannote models in the final pipeline.
 
@@
         \hline
-        \multicolumn{1}{| c |}{} & \multicolumn{3}{| c |}{Tiage} &  \multicolumn{3}{| c |}{SuperDialseg}  \\
-        \hline
-        model & Pk $\downarrow$ & wd $\downarrow$ & f1 & Pk $\downarrow$ & wd $\downarrow$ & f1 \\ 
+        \multirow{2}{*}{models} & \multicolumn{3}{| c |}{Tiage} &  \multicolumn{3}{| c |}{SuperDialseg}  \\
+        \cline{2-7}
+         & Pk $\downarrow$ & wd $\downarrow$ & f1 & Pk $\downarrow$ & wd $\downarrow$ & f1 \\ 
         \hline
@@
         \hline\ 
-        Our solution BERT-NSP & Pk & wd & f1 & Pk & wd & f1 \\ 
-        Our solution BERT-CLS & Pk & wd & f1 & Pk & wd & f1 \\ 
+        Our solution+BERT-NSP & Pk & wd & f1 & Pk & wd & f1 \\ 
+        Our solution+BERT-CLS & Pk & wd & f1 & Pk & wd & f1 \\ 
         \hline
-    \end{tabular}
-    \captionof{table}{Text datasets segmentation}
-\end{center}
-
-\begin{center}
-    \begin{tabular}{ | c | c | c | c | c | c | c | } 
-        \hline
-        \multicolumn{1}{| c |}{} & \multicolumn{3}{| c |}{AMI} &  \multicolumn{3}{| c |}{ICSI} \\
-        \hline
-        model  & Pk $\downarrow$ & wd $\downarrow$ & f1 & Pk $\downarrow$ & wd $\downarrow$ & f1 \\ 
+        \multirow{2}{*}{models} & \multicolumn{3}{| c |}{AMI} & \multicolumn{3}{| c |}{ICSI} \\
+        \cline{2-7}
+        & Pk $\downarrow$ & wd $\downarrow$ & f1 & Pk $\downarrow$ & wd $\downarrow$ & f1 \\ 
         \hline
@@
         \hline
-        Our solution BERT-NSP & Pk & wd & f1 & Pk & wd & f1 \\ 
-        Our solution BERT-CLS & Pk & wd & f1 & Pk & wd & f1 \\  
+        Our solution+BERT-NSP & Pk & wd & f1 & Pk & wd & f1 \\ 
+        Our solution+BERT-CLS & Pk & wd & f1 & Pk & wd & f1 \\  
         \hline
     \end{tabular}
-    \captionof{table}{Audio datasets segmentation}
+    \captionof{table}{Dialogue segmentation experiments}
 \end{center}
@@
         \hline
-        \multicolumn{1}{| c |}{} & \multicolumn{2}{| c |}{Tiage} &  \multicolumn{2}{| c |}{SuperDialseg}  \\
-        \hline
-        model & TD & TC & TD & TC  \\ 
+        \multirow{2}{*}{models} & \multicolumn{2}{| c |}{Tiage} &  \multicolumn{2}{| c |}{SuperDialseg}  \\
+        \cline{2-5}
+         & TD & TC & TD & TC  \\ 
         \hline
@@
         \hline
-    \end{tabular}
-    \captionof{table}{Text datasets topics extraction}
-\end{center}
-
-\begin{center}
-    \begin{tabular}{ | c | c | c | c | c | c | c |} 
-        \hline
-        \multicolumn{1}{| c |}{} & \multicolumn{2}{| c |}{Product} &  \multicolumn{2}{| c |}{Academic} \\
-        \hline
-
+        \multirow{2}{*}{models} & \multicolumn{2}{| c |}{AMI} &  \multicolumn{2}{| c |}{ICSI} \\
+        \cline{2-5}
+         & TD & TC & TD & TC  \\ 
         \hline
@@
     \end{tabular}
-    \captionof{table}{Audio datasets topics extraction}
+    \captionof{table}{Dialogue topics extraction experiments}
 \end{center}
@@
 The result of experiments in zero-shot that LlaMA 3 8B demonstrate the best performance, it is far ahead of everyone in Topic Diversity and competes in Topic Coherence with BertTopic. 
-A strong drawdown in quality relative to extractive model was also noticed in the article PromptTopic \cite{prompttopic:2023}, where the models also lost to BertTopic in Topic Coherence.
+A strong drawdown in quality relative to extractive model was also noticed in the article PromptTopic \cite{prompttopic:2023}, where the models also lost to BertTopic in Topic Coherence. \\
+
+The main problem at this stage is the problems with the format of the topic words themselves and the formats of the generated output of the model. 
+Firstly, the model sometimes gives out not only the word itself but also an explanation of why, there are also options when the model outputs whole sentences or phrases of 3 or more words, which does not suit our task.
+\\
+For example:
+
+\begin{lstlisting}
+    good answer for dialogue: 
+    car, summer, camp.
+
+    possible bad variants:
+    -------------------
+    car (because they travel by car)
+    summer (because all dialogue about summer)
+    camp (because the spent a lot of time in camp)
+    -------------------
+    we take a car
+    spent the whole summer together
+    camp was an important place
+    -------------------
+\end{lstlisting}
+
+Moreover, the zero-shot format has problems with the format of the output generated sequence. 
+The model with the same prompt sometimes produces a different format and parsing topic words becomes much more difficult.
+
+\begin{lstlisting}
+    possible bad variants:
+    -------------------    
+    1. car
+    2. summer
+    3. camp
+    -------------------
+    * car
+    * summer
+    * camp
+    -------------------
+\end{lstlisting}
+
+Consequently, it was important to use different techniques "to embrace" model the correct formats for convenient operation in the whole pipeline.
 
 \subsection{Few-shot prompting}
+
+The few-short approach was explained in the chapter of Methodology.
 
@@
         Mistral zero-shot & 0.29852 & 0.084103 & 0.5761 & -0.11318 \\ 
-        Mistral few-shot & \cellcolor{green!20} 0.31920 & \cellcolor{red!10} 0.065381 & 0.5761 & \cellcolor{green!20} -0.08673 \\ 
+        Mistral few-shot & \cellcolor{green!20} 0.31920 & \cellcolor{red!10} 0.065381 & 0.5761 & \cellcolor{green!20} -0.008673 \\ 
         \hline
     \end{tabular}
-    \captionof{table}{Comparisons of zero-shot and few-shot results}
+    \captionof{table}{Comparisons of models in zero-shot and few-shot}
 \end{center}
@@
 
+It was also decided to carry out fine-tuning of the models on the selected train part of the datasets (tiage, superseg). 
+The main point of this stage is to train the model for a given topic words format and their output format, as in few-shot. 
+In addition, it is possible to improve the quality metrics of the models.
+
+\begin{center}
+    \begin{tabular}{ | c | c | c | c | c | c | c |} 
+        \hline
+        \multicolumn{1}{| c |}{} & \multicolumn{2}{| c |}{AMI} &  \multicolumn{2}{| c |}{ICSI} \\
+        \hline
+        model & TD & TC & TD & TC  \\ 
+        \hline
+        LLaMA3 few-shot & 0.51291 & 0.12129 & 0.78201 & 0.13985 \\ 
+        LLaMA3 fine-tuning & \cellcolor{green!30} 0.53812 & \cellcolor{green!20} 0.13329 & \cellcolor{green!20} 0.79854 & \cellcolor{green!30} 0.15636 \\
+        \hline
+        Qwen1.5 few-shot & 0.38875 & 0.10190 & 0.71960 & 0.00239 \\ 
+        Qwen1.5 fine-tuning & \cellcolor{green!30} 0.39471 & \cellcolor{green!20} 0.11004 & \cellcolor{green!20} 0.72041 & 0.00241 \\ 
+        \hline
+        Mistral few-shot & 0.31920 & 0.065381 & 0.57617 & -0.008673 \\ 
+        Mistral fine-tuning & \cellcolor{green!10} 0.31998 & 0.065421 & 0.57531 & \cellcolor{green!50} 0.00031 \\ 
+        \hline
+    \end{tabular}
+    \captionof{table}{Comparisons of models in zero-shot and after fine-tuning}
+\end{center}
+
+As with few-shot, the metrics have increased, but now it is convenient to work with the model in the whole pipeline, because now it is easy to understand the "behavior" of the model during response generation.
+In addition, the increase in the metric comes from the fact that it was possible to parse all samples, unlike some in zero-shot and few-shot.
+
 \section{Topic Evolution}
+
+In this section, we will show examples of the results of analyzing changes in topics over time.
+All examples will be presented on the AMI dataset.
+
+Topic evolution over time on one dialogue:
+
+Topic evolution over time on dataset:
 

### 2024-05-21 paper/main.tex
update segmetation exps

@@
 
-\section{Text clustering}
+\section{Dialogue Segmenentation}
 
@@
 At this stage, experiments will be conducted using the encoder model for the Utterance-Coherence Scoring algorithm.
-In addition, algorithms for the time of model inference will be performed throughout the pipeline to select the fastest solution.
 
@@
         GraphSeg & 0.462 & 0.466 & 0.239 & 0.496 & 0.515 & 0.238 \\
-        TextTilling & 0.469 & 0.488 & 0.204 & 0.446 & 0.458 & 0.378 \\
-        TeT + BERT & 0.425 & 0.439 & 0.285 & 0.522 & 0.531 & 0.201 \\
+        TextTilling & 0.469 & 0.488 & 0.204 & \cellcolor{green!20} 0.446 & \cellcolor{green!20} 0.458 & \cellcolor{green!20} 0.378 \\
+        TeT + BERT & \cellcolor{green!20} 0.425 & \cellcolor{green!20} 0.439 & \cellcolor{green!20} 0.285 & 0.522 & 0.531 & 0.201 \\
         \hline\ 
-        Our solution+BERT-NSP & Pk & wd & f1 & Pk & wd & f1 \\ 
-        Our solution+BERT-CLS & Pk & wd & f1 & Pk & wd & f1 \\ 
+        CohScoring+BERT-NSP & 0.4465 & 0.4575 & 0.267 & 0.5071 & 0.5219 & 0.205 \\ 
+        CohScoring+BERT-NSP fine-tune & \cellcolor{green!40} 0.3981 & \cellcolor{green!40} 0.4018 & \cellcolor{green!40} 0.32 & \cellcolor{green!40} 0.4361 & \cellcolor{green!40} 0.44510 & \cellcolor{green!40} 0.285 \\
         \hline
@@
         \hline
-        GreedySeg & 0.885 & 0.792 & 0.0641 & 0.762 & 0.730 & 0.109 \\
-        GraphSeg & Pk & wd & f1 & Pk & wd & f1 \\
+        GreedySeg & 0.885 & 0.792 & 0.3641 & 0.762 & 0.730 & 0.309 \\
+        GraphSeg & 0.875 & 0.750 & 0.422 & 0.810 & 0.756 & 0.275 \\
         TextTilling & Pk & wd & f1 & Pk & wd & f1 \\
-        TeT + BERT & Pk & wd & f1 & Pk & wd & f1 \\ 
+        TeT + BERT & Pk & wd & 0.443 & Pk & wd & 0.158 \\ 
         \hline
-        Our solution+BERT-NSP & Pk & wd & f1 & Pk & wd & f1 \\ 
-        Our solution+BERT-CLS & Pk & wd & f1 & Pk & wd & f1 \\  
+        CohScoring+BERT-NSP & \cellcolor{green!20} 0.6447 & \cellcolor{green!20} 0.7394 & \cellcolor{green!20} 0.551 & \cellcolor{green!20} 0.6148 & \cellcolor{green!20} 0.6731 & \cellcolor{green!20} 0.779 \\
+        CohScoring+BERT-NSP fine-tune & \cellcolor{green!40} 0.5854 & \cellcolor{green!40} 0.6032 & \cellcolor{green!40} 0.587 & \cellcolor{green!40} 0.5842 & \cellcolor{green!40} 0.5932 & \cellcolor{green!40} 0.815 \\
         \hline
@@
 
-The selected solution shows the best metrics on the datasets we selected. It is important to understand that we have researched open source solutions.
+The selected solution shows the best metrics on the datasets we selected. 
+It is important to understand that we have researched open source solutions.
 

### 2024-05-21 paper/main.tex
update topic evolution and paper

@@
 
+\begin{center}
+    \centering
     Topic evolution over time on one dialogue:
+\end{center}
 
+\begin{lstlisting}
+    dialogue_topics_analyze(dialogue, 
+                            n_topic_words)
+                        \end{lstlisting}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.5]{img/output.png}
+    \caption{Topic evolution over time on one dialogue}
+\end{figure}
+
+The picture shows the change of themes over time. 
+There were 5 segments in this dialog, so 5 timestamps. \\
+
+\begin{center}
+    \centering
     Topic evolution over time on dataset:
+\end{center}
+
+\begin{lstlisting}
+    corpus_topics_analyze(dataset, n_topic_words,
+                          top_popular=8, time_range=50)
+\end{lstlisting}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.7]{img/ami_topics_evol.png}
+    \caption{Topic evolution in AMI dataset}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.7]{img/topic_evol_icsi.png}
+    \caption{Topic evolution in ICSI dataset}
+\end{figure}
+
+In analyzing the evolution of topics on the dataset, top\_popular is also selected, because there are many more topics on one dialog and it is impossible to analyze everything at once. 
+In addition, the time\_range parameter is responsible for reducing time intervals for more convenient analysis.
 

### 2024-05-23 paper/main.tex
update lib

@@
 The dialogue segmentation stage represents the initial stage for textual type of data. 
-The result is a set of utterance indices that indicate the beginning of a thematically related group.
+The result is a set of utterance indices that indicate the beginning of a topicaly related group.
 
@@
 
-In order to indicate the thematic similarity of two statements in a dialogue, the $[CLS]$ token is employed, which performs a similar role in the usual BERT. 
+In order to indicate the topic similarity of two statements in a dialogue, the $[CLS]$ token is employed, which performs a similar role in the usual BERT. 
 The resulting value is subjected to a linear dictionary and a final similarity value is obtained, which is then utilised to calculate Coherence scores.
@@
 
-The greater the depth score between two statements, the less likely it is that they share a thematic relationship. 
-To identify the pairs along which the boundaries of the thematic segments lie, the threshold value is calculated as follows: $threshold = \mu - \frac{\sigma}{2}$, where $\mu$ is the average value and $\sigma$ is the standard deviation.
+The greater the depth score between two statements, the less likely it is that they share a topic relationship. 
+To identify the pairs along which the boundaries of the topic segments lie, the threshold value is calculated as follows: $threshold = \mu - \frac{\sigma}{2}$, where $\mu$ is the average value and $\sigma$ is the standard deviation.
 Pairs with a depth score above the $threshold$ value will be choosen as boundaries.
@@
     \item \textbf{Topic coherence} \cite{topiccoh:2009} is a measure of the degree of interconnection of words within a topic.
-In our experiments, we employed normalized point-to-point mutual information (PMI) as a metric for thematic consistency.
+In our experiments, we employed normalized point-to-point mutual information (PMI) as a metric for topic consistency.
 A higher score for a given name indicates greater consistency, with an absolute correlation estimated at 1 point.
@@
 
-The pipeline of dynamic thematic modelling of audio dialogues, which will be considered in this paper, is divided into several main stages. 
+The pipeline of dynamic topic modelling of audio dialogues, which will be considered in this paper, is divided into several main stages. 
 These include ASR, dialogue segmentation, topic extraction and topic evolution.
@@
 will be used to create the pipeline.
+
+The main stage to be investigated is topic extraction. 
+The use of LLM (generative approach) for topic modeling began to be investigated quite recently. 
+For audio dialogues, which are quite different from text ones, it was not used at all.
 
@@
         \hline\ 
-        CohScoring+BERT-NSP & 0.4465 & 0.4575 & 0.267 & 0.5071 & 0.5219 & 0.205 \\ 
-        CohScoring+BERT-NSP fine-tune & \cellcolor{green!40} 0.3981 & \cellcolor{green!40} 0.4018 & \cellcolor{green!40} 0.32 & \cellcolor{green!40} 0.4361 & \cellcolor{green!40} 0.44510 & \cellcolor{green!40} 0.285 \\
+        Our method & 0.4465 & 0.4575 & 0.367 & 0.5071 & 0.5219 & 0.430 \\ 
+        Our method + fine-tune & \cellcolor{green!40} 0.4094 & \cellcolor{green!40} 0.4048 & \cellcolor{green!40} 0.7374 & \cellcolor{green!40} 0.4361 & \cellcolor{green!40} 0.44510 & \cellcolor{green!40} 0.7421 \\
         \hline
@@
         \hline
-        CohScoring+BERT-NSP & \cellcolor{green!20} 0.6447 & \cellcolor{green!20} 0.7394 & \cellcolor{green!20} 0.551 & \cellcolor{green!20} 0.6148 & \cellcolor{green!20} 0.6731 & \cellcolor{green!20} 0.779 \\
-        CohScoring+BERT-NSP fine-tune & \cellcolor{green!40} 0.5854 & \cellcolor{green!40} 0.6032 & \cellcolor{green!40} 0.587 & \cellcolor{green!40} 0.5842 & \cellcolor{green!40} 0.5932 & \cellcolor{green!40} 0.815 \\
+        Our method & \cellcolor{green!20} 0.6447 & \cellcolor{green!20} 0.7394 & \cellcolor{green!20} 0.551 & \cellcolor{green!20} 0.6148 & \cellcolor{green!20} 0.6731 & \cellcolor{green!20} 0.779 \\ 
+        Our method + fine-tune & \cellcolor{green!40} 0.5279 & \cellcolor{green!40} 0.6932 & \cellcolor{green!40} 0.610 & \cellcolor{green!40} 0.5842 & \cellcolor{green!40} 0.5932 & \cellcolor{green!40} 0.815 \\
         \hline
@@
 \section{Topic extraction}
+
+At this stage, we conducted the most experiments, ready-made solutions were used or adapted for all stages, at the topic extraction stage we used LLM without relying on ready-made solutions.
 
@@
         \hline
-        LLaMA3 8B & \cellcolor{green!60} 0.50863 & \cellcolor{green!20} 0.11409 & \cellcolor{green!60} 0.76137 & \cellcolor{green!20} 0.14734 \\ 
-        Qwen1.5  & \cellcolor{green!20} 0.38641 & 0.10328 & \cellcolor{green!20} 0.7185  & 0.00237\\ 
-        Mistral 7B & 0.29852 & 0.084103 & 0.5761 & -0.11318 \\ 
+        LLaMA3 8B & \cellcolor{green!60} 0.50863 & \cellcolor{green!20} 0.01140 & \cellcolor{green!60} 0.76137 & \cellcolor{green!20} 0.14734 \\ 
+        Qwen1.5  & \cellcolor{green!20} 0.38641 & 0.01032 & \cellcolor{green!20} 0.7185  & 0.00237\\ 
+        Mistral 7B & 0.29852 & 0.00841 & 0.5761 & -0.11318 \\ 
         \hline
     \end{tabular}
+    \label{table:topicextr}
     \captionof{table}{Dialogue topics extraction experiments}
@@
 Experiments were carried out with each of the algorithms at each stage on ready-made models, pre-trained and trained from scratch.
-The best models were added to the final pipeline;;
+The best models were added to the final pipeline;
     \item At the topic extraction stage, a study was conducted on the use of LLM.
 The best open-source LLMs were selected for experiments with zero-shot, feat-shot and fine-tuning.
-The generative approach using LLM shows the results of the best available extractive solutions for topic modeling;
-    \item A suitable set of metrics was identified and selected to evaluate the quality of each stage.
-In order to utilise the metrics, a library with an implementation was Investigated and integrated into pipeline.
+Using LLM at this stage, it was possible to increase the metrics for audio dialogues.
+\\
+     - Topic Diversity was increased around 10 times, and Topic Coherence is about the same value compared to BertTopic and LSA as shown in \ref{table:topicextr}; \\
+     - As for the textual dialogues worth noting that the increase ща Topic Deversity is less and on Topic Coherence BertTopic sometimes shown higher results.
 \end{itemize}
@@
 \putbibliography
+\nocite{hydro:2023}
+\nocite{sesseg:2016}
+\nocite{todbert:2020}
+\nocite{topaware:2020}
+\nocite{Plaquet23:2023}
+\nocite{dialsystem:2018}
+\nocite{jointmdl:2020}
+\nocite{clusterllm:2023}
+\nocite{atnm:2023}
+\nocite{plsa:2013}
+
 \end{document}

### 2024-05-24 paper/main.tex
paper fin

@@
     \item The identification of suitable datasets for experiments;
-    \item The investigation of the best suitable open-source solutions for speaker diarisation, dialogue clustering, topic extraction and topic evolution;
+    \item The investigation of the best suitable open-source solutions for speaker diarization, dialogue clustering, topic extraction and topic evolution;
     \item The exploration of the possibilities of using Large Language Models for some stages and the selection of models suitable for training in limited resources;
@@
 
-Speaker diarisation is the process of automatically segmenting and identifying different speakers in an audio recording. 
-The goal of speaker diarisation is to divide the audio stream into non-homogeneous segments, each segment corresponding to a specific speaker or speaker turn.
+Speaker diarization is the process of automatically segmenting and identifying different speakers in an audio recording. 
+The goal of speaker diarization is to divide the audio stream into non-homogeneous segments, each segment corresponding to a specific speaker or speaker turn.
 
@@
 The pipeline of dynamic topic modelling of audio dialogues, which will be considered in this paper, is divided into several main stages. 
-These include ASR, dialogue segmentation, topic extraction and topic evolution.
+These include speaker diarization, dialogue segmentation, topic extraction and topic evolution.
 
@@
             \hline
-            dataset & train/val/test & Avg Turns & Avg tokens \\ 
+            dataset & train/val/test & Avg Turns \\ 
             \hline
-            SuperDialseg & 6984/1322/1322 & 13 & 239 \\ 
-            Tiage &  7939/-/1000 & 14.7  & - \\ 
+            SuperDialseg & 6984/1322/1322 & 13  \\ 
+            Tiage &  7939/-/1000 & 14.7 \\ 
             \hline
-            Product Meetings (AMI) & 137 & 535.6 & - \\ 
-            Academic Meetings (ICSI) & 59 & 819.0 & - \\ 
+            Product Meetings (AMI) & 137 & 535.6 \\ 
+            Academic Meetings (ICSI) & 59 & 819.0 \\
             \hline
@@
         \hline\ 
-        Our method & 0.4465 & 0.4575 & 0.367 & 0.5071 & 0.5219 & 0.430 \\ 
-        Our method + fine-tune & \cellcolor{green!40} 0.4094 & \cellcolor{green!40} 0.4048 & \cellcolor{green!40} 0.7374 & \cellcolor{green!40} 0.4361 & \cellcolor{green!40} 0.44510 & \cellcolor{green!40} 0.7421 \\
+        Our method & 0.447 & 0.457 & 0.367 & 0.507 & 0.522 & 0.430 \\ 
+        Our method + fine-tune & \cellcolor{green!40} 0.409 & \cellcolor{green!40} 0.405 & \cellcolor{green!40} 0.737 & \cellcolor{green!40} 0.436 & \cellcolor{green!40} 0.445 & \cellcolor{green!40} 0.742 \\
         \hline
@@
         \hline
-        GreedySeg & 0.885 & 0.792 & 0.3641 & 0.762 & 0.730 & 0.309 \\
+        GreedySeg & 0.885 & 0.792 & 0.364 & 0.762 & 0.730 & 0.309 \\
         GraphSeg & 0.875 & 0.750 & 0.422 & 0.810 & 0.756 & 0.275 \\
-        TextTilling & Pk & wd & f1 & Pk & wd & f1 \\
-        TeT + BERT & Pk & wd & 0.443 & Pk & wd & 0.158 \\ 
+        TextTilling & 0.867 & 0.730 & 0.429 & 0.805 & 0.753 & 0.295 \\
+        TeT + BERT & 0.794 & 0.705 & 0.510 & 0.798 & 0.720 & 0.158 \\ 
         \hline
-        Our method & \cellcolor{green!20} 0.6447 & \cellcolor{green!20} 0.7394 & \cellcolor{green!20} 0.551 & \cellcolor{green!20} 0.6148 & \cellcolor{green!20} 0.6731 & \cellcolor{green!20} 0.779 \\ 
-        Our method + fine-tune & \cellcolor{green!40} 0.5279 & \cellcolor{green!40} 0.6932 & \cellcolor{green!40} 0.610 & \cellcolor{green!40} 0.5842 & \cellcolor{green!40} 0.5932 & \cellcolor{green!40} 0.815 \\
+        Our method & \cellcolor{green!20} 0.645 & \cellcolor{green!20} 0.739 & \cellcolor{green!20} 0.551 & \cellcolor{green!20} 0.615 & \cellcolor{green!20} 0.673 & \cellcolor{green!20} 0.779 \\ 
+        Our method + fine-tune & \cellcolor{green!40} 0.528 & \cellcolor{green!40} 0.693 & \cellcolor{green!40} 0.610 & \cellcolor{green!40} 0.584 & \cellcolor{green!40} 0.593 & \cellcolor{green!40} 0.815 \\
         \hline
@@
 The best models were added to the final pipeline;
+    \item In terms of dialogues segmentation, BERT was trained specifically for audio dataset to calculate depth scores and gave better results than TextTililng by 0.1 for f1 metrics and 0.15 less for pk and windiff.
     \item At the topic extraction stage, a study was conducted on the use of LLM.
@@
      - Topic Diversity was increased around 10 times, and Topic Coherence is about the same value compared to BertTopic and LSA as shown in \ref{table:topicextr}; \\
-     - As for the textual dialogues worth noting that the increase ща Topic Deversity is less and on Topic Coherence BertTopic sometimes shown higher results.
+     - As for the textual dialogues worth noting that the increase of Topic Deversity is less and on Topic Coherence BertTopic sometimes shown higher results.
 \end{itemize}
